---
title:  "Plasticity for the win: Flexible transcriptional response to host plant switches
  in the comma butterfly (Polygonia c-album)"
subtitle: "Performance assay"
author: "Katharina Schneider, Rachel A. Steward, Maria de la Paz Celorio-Mancera, Niklas Janz, Dick Moberg, Christopher W. Wheat, Sören Nylin "
date: "2024-06-25"
output: html_document
---

## Background

Larvae of Pca (Polygonia c-album) were reared on either of 3 different host plants (U=Urtica dioica, S=Salix caprea, R=Ribes uva-crispa).
Larvae were reared under short-day conditions at 17C with a 12:12h light cycle.
When reaching the 5th instar (earlier stages were hard to dissect) larvae were moved to a new plant. This was either the same species (e.g. U-U; "control") or an alternative host (e.g. U-S, "switch").


# Load packages
```{r load packages, "message = F"}
library(readxl) # version "1.4.1"
library(ggplot2) #version "3.5.0"
library(DHARMa) #version "0.4.6"
library(glmmTMB) #version "1.1.9"
library(tidyverse) #version "1.3.2"
library(car) #version "3.1-2"
``` 

# set directory
```{r set directory}
setwd=("/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Performance/")
```

# Load and prepare data
```{r load and prepare data}
#SE_MC=read_excel("/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Switch_experiment_MC.xlsx")
SE_MC=read_excel("Switch_experiment_MC.xlsx")
str(SE_MC)

#developmental stages are categorized in column "instar_edit". i corresponds to stages before the switch, i5 marks the measurement when a larva reached the 5th instar, as corresponds to measurements after the switch, p corresponds to pupa

SE_MC$Family=as.factor(SE_MC$Family)
SE_MC$Individual=as.factor(SE_MC$Individual)
SE_MC$Weight=as.numeric(SE_MC$Weight)
SE_MC$Weight_µg=1000*(SE_MC$Weight) #change unit to avoid negative logs

``` 

#calculate performance
performance was m easured using growth rate
Growth rate was calculated as = log(mass)/developmental time
(since we want to compare the daily weight gain developmental time was 1)

```{r calculate growth rate}
SE_MC$logWeight=log(SE_MC$Weight_µg)
```


#add Treatment as a category
```{r make colmun with treatment categories}
SE_MC$treatment_lat=paste(SE_MC$Start_lat, SE_MC$End_lat, sep="-") #_lat=refers to column containing Satin names of the host
```


#exclude pupal weights
since we wanted to measure the effect on daily weight gain during larval development, pupal weights were excluded
```{r remove pupal weights}
SE_larv=SE_MC[SE_MC$Age!="P",]
SE_larv$Day=as.numeric(SE_larv$Age)
```


#highlight time of host switch
to be able to mark the day of the host switch, we need to find the days of the switch
these are named "i5" in the data sheet, as these are the days/the age when the larvae reached the 5th instar
to get the host switch timepoint per treatment, the avereage was calculated per group. 

```{r host switch timing}

SE_MC_i5=SE_MC[SE_MC$Instar=="i5",] #this will collect all the lines with "i5"
SE_MC_i5$Day=as.numeric(SE_MC_i5$Age)

#make averages to get average age at host switch per treatment group
ave_age=SE_larv[SE_larv$Instar_edit=="i5",] %>%
  group_by(treatment_lat) %>%
  summarize(average_value=mean(as.numeric(Age)))

print(ave_age)

```


#plot growth rate
```{r plot growth rate}

#prepare vector to include vertical lines when larvae were switched (average age!)

vline_data <- data.frame(
  treatment_lat = c("U-U", "S-U", "R-U", "U-S", "S-S", "R-S", "U-R", "S-R", "R-R"),
  vline_x = c(15.8, 23.4, 20.8,17.7,  19.7, 20.3,  18 ,  24,   21.1 )
)

ggplot(SE_larv, aes(x = Day, y = logWeight, color = treatment_lat)) +
  geom_point() +
  geom_smooth(method=loess, se = TRUE, aes(fill = treatment_lat), alpha = 0.2) + #For method = NULL the smoothing method is chosen based on the size of the largest group (across all panels). stats::loess() is used for less than 1,000 observations; otherwise mgcv::gam() is used with formula = y ~ s(x, bs = "cs") with method = "REML". Somewhat anecdotally, loess gives a better appearance, but is O(N2) in memory, so does not work for larger datasets.
  scale_color_manual(values = c("#888888", "#CC6677", "#DDCC77", "#6699CC", "#661100", "#117733", "#332288", "#AA4499", "#999933")) +
  scale_fill_manual(values = c("#888888", "#CC6677", "#DDCC77", "#6699CC", "#661100", "#117733", "#332288", "#AA4499", "#999933")) +
  facet_wrap(~factor(treatment_lat, c("U-U", "S-U", "R-U", "U-S", "S-S", "R-S", "U-R", "S-R", "R-R"))) +
  #facet_wrap(~factor(treatment, c("U-U", "U-S", "U-R", "S-U", "S-S", "S-R", "R-U", "R-S", "R-R"))) +
  geom_vline(data = vline_data, aes(xintercept = vline_x)) + #add lines to plots for day of switch
  theme_bw() +
  scale_x_continuous(breaks = seq(from = 0, to = 36, by = 4)) +
  theme(axis.title = element_text(size = 16)) +
  theme(axis.text = element_text(size = 14)) +
  theme(strip.text = element_text(size = 15, face = "bold")) +
  ylab("Growth rate") +
  theme(legend.position = "none")

```

# tedt for an effect of host switch on performance
```{r linear model}

model = glmmTMB(logWeight ~ treatment_lat + (1|Family/Individual), data = SE_larv, family=gaussian) #glmmTMB was necessary to run model with nested random effect

res_model=simulateResiduals(model)
plot(res_model)

Anova(model)#--> no sign difference --> no effect on performance
    # Chisq Df Pr(>Chisq)
    # treatment_lat 13.604 8    0.0927
```


# testing the effect of host switch on performance after the switch
This is included in the supplementary

for this we first need to subset the dataset and remove every measurement that was taken before the larvae were switched to a new host plant (i.e. after switch until start of pupation; pupal weight NOT included!)

```{r subsetting data}
After_switch=SE_larv[SE_larv$Instar_edit!="i",]
After_switch$Day=as.numeric(After_switch$Day)
```

#plot growth rate after switch 

```{r plot growth rate only for measurements after switch}
ggplot(After_switch, aes(x=Day, y=logWeight, color=treatment_lat))+
  geom_point()+ #this will plot the individual data points
  geom_smooth(se=TRUE,  aes(fill=treatment_lat), alpha=0.2)+
  scale_color_manual(values = c("#888888", "#CC6677", "#DDCC77", "#6699CC", "#661100", "#117733", "#332288", "#AA4499", "#999933")) +
  scale_fill_manual(values = c("#888888", "#CC6677", "#DDCC77", "#6699CC", "#661100", "#117733", "#332288", "#AA4499", "#999933")) +
  #facet_wrap(~factor(After_switch$treatment_lat, levels=c("U-U", "U-S", "U-R", "S-S", "S-U", "S-R", "R-R", "R-U", "R-S")))+
  facet_wrap(~factor(treatment_lat, c("U-U", "S-U", "R-U", "U-S", "S-S", "R-S", "U-R", "S-R", "R-R"))) +
  theme_bw()+
  scale_x_continuous(breaks=seq(from=0, to=36, by=2))+
  theme(axis.title=element_text(size=15))+
  theme(axis.text=element_text(size=12))+
  theme(strip.text = element_text(size=13, face="bold"))+
  ggtitle("After Switch")+
  labs(x="Growth rate")+
  theme(legend.position = "none")
```

#test effect of host switch on performance only using measurements after the switch

```{r effect of host switch - after switch measurements}

lm_SP=glmmTMB(logWeight ~ treatment_lat + (1|Family/Individual), data = After_switch, family=gaussian)

sim_lm_SP=simulateResiduals(lm_SP)
plot(sim_lm_SP)

Anova(lm_SP) # --> no sign difference
  #Chisq Df Pr(>Chisq)
  #treatment_lat 3.3205  8     0.9127

``` 



