---
title: "Plasticity for the win: Flexible transcriptional response to host plant switches
  in the comma butterfly (Polygonia c-album)"
author: "Katharina Schneider, Rachel A. Steward, Maria de la Paz Celorio-Mancera,
  Niklas Janz, Dick Moberg, Christopher W. Wheat, Sören Nylin"
date: "2024-06-25"
output: html_document
---

#1. Background

This experiment is based on gene expression data from 2 different experiments, that differ in the used host plants.
In the first experiment (Switch 1), larvae were switched between Urtica dioica (N resp. U for later analysis), Salix caprea (S) and Ulmus (E). The second switch experiment was done between Urtica dioica, Salix caprea and Ribes uva-crispa (G resp. R).

Larvae of Pca (Polygonia c-album) were reared on either of 3 different host plants (N, S, E resp. U, S, G).
Larvae were reared under short-day conditions at 17C with a 12:12h light cycle.
When reaching the 5th instar (earlier stages were hard to dissect) larvae were moved to a new plant. This was either the same species (e.g. N-N; "control") or an alternative host (e.g. N-S, "switch").
To check for potential transcriptional changes after the switch, gut tissue was dissected 2hrs (the time that a food pellet roughly needs to pass through the larva) and 17hrs after larvae were switched/initiated feeding on the new host.
3 families with 2 individuals per family were used per treatment to include potential variation. However: family 3 had some peculiarities --> they showed more progressed developmental characters (shrunk gut, elongated body)--> family 3 was thus excluded from extraction labelling of the samples: "Host1Host2Family.Individual.Hours", e.g. NN2.1.2hrs, NS1.2.17hrs

##GOAL 
the overall expression profile of the larvae will be compared to check for host specific adaptations/expression patterns/modules

##RNA extraction and sequencing
Extraction was done using: TRI-reagent/BCP RNA isolation protocol with DNA depletion
quality check was done using Bioanalyzer, concentrations were measured using Qubit
Extracts were send to SciLifeLab for library preparation and sequencing (according to them they measured a concentration 10 times higher,
than what we measured with Qubit, so we agreed, that they will dilute the samples again and proceed with the library prep and sequencing)
data processing (trimming, mapping, abundance estimation) was done on Uppmax

##Reference genome
Celorio-Mancera et al. 2021: Chromosome level assembly of the comma butterly (Polygonia c-album)

#2. load packages
```{r load packages, "message = F"}

library(limma) #version 3.54.0
library(edgeR) #version 3.40.0
library(tidyverse) #version 1.3.2 
library(dplyr) #version 1.1.4
library(ggplot2) #version 3.4.0
library(ggpubr) #version 0.5.0
library(stringr) #version 1.5.1
library(gplots) #version 3.1.3.1
library(ggtext) #version 0.1.2
library(ComplexHeatmap) #version 2.14.0
library(circlize) #version 0.4.16
library(scales) #version 1.3.0
library(ggVennDiagram) #version 1.5.2
library(UpSetR) #version 1.4.0
library(cowplot) #version 1.1.3
library(gridExtra) #version 2.3
library(topGO) #version 2.50.0
library(ggplotify) #version 0.1.2
library(RColorBrewer) #version 1.1-3
library(xlsx) #version 0.6.5

```

#3. Data preparation

## load count data

Switch 1 (i.e. Switch A)
...........................
The first Switch experiment was done using: Ulmus glabra, Salix caprea and Urtica dioica
experimental design: 4 families, 1 individual per family, sampling at two timepoints (2hrs and 17hrs), technical replicates !
36 samples per timepoint (some needed to be removed, see below)

```{r read in Switch 1}

Switch_1_all=read.delim("/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Gene_counts/Switch1/Count_S1", row.names="Geneid")

# remove samples with low mapping success (ca 40%): ES_tworep8, ES_tworep7
#additionally 2 samples were identified as outliers in the PCA and where thus removed (EN_tworep1, NS_seventeenrep1)
Switch_1=Switch_1_all[ , -which(names(Switch_1_all) %in% c("ES_tworep8", "ES_tworep7", "EN_tworep1", "EN_tworep2", "NS_seventeenrep1", "NS_seventeenrep2", "Chr", "Start", "End", "Strand", "Length"))]


#4 families, 1 larva per family
#BioRep=Family
Meta_1_all=read.csv("/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Switch_1_Meta.csv", header=TRUE)
Meta_1=Meta_1_all[ !(Meta_1_all$Sample) %in% c("EN_tworep1", "EN_tworep2", "NS_seventeenrep1", "NS_seventeenrep2"),]

dim(Switch_1) # 21761   138
dim(Meta_1) # 138  10
```


Switch 2 (i.e. Switch B)
..........................
The second Switch experiment was done using: Gooseberry, Salix and Nettle
experimental design: 2 families, 2 individuals per families, sampling at two timepoints (2hrs and 17hrs)
37 samples per timepoint

```{r read in Switch 2}

Switch_2_all=read.delim("/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Gene_counts/Switch2/Count_S2.txt", row.names="Geneid")

#two samples from family 3 were used to test the extraction protocol. Remove these samples!
#additionally SN_tworep_1.2 and SN_tworep_1.2B and NG_tworep_2.2 were remmoved due to uncertainties of the sampling time. 
Switch_2=Switch_2_all[, -which(names(Switch_2_all) %in% c("NS_seventeenrep_3.1", "NS_seventeenrep_3.2", "SN_tworep_1.2", "SN_tworep_1.2B", "NG_tworep_2.2", "Chr", "Start", "End", "Strand", "Length"))]


#2 families, 2 larvae per family
Meta_2_all=read.csv("/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Switch_2_Meta.csv", header=TRUE)
Meta_2=Meta_2_all[ !(Meta_2_all$Sample) %in% c("NS_seventeenrep_3.1", "NS_seventeenrep_3.2", "SN_tworep_1.2", "SN_tworep_1.2B", "NG_tworep_2.2"),]

dim(Switch_2) #21761    69
dim(Meta_2) # 69 16
```

##split data set by sampling time point (i.e. 2hrs and 17 hrs)

```{r split by time}
#Switch 1 2hrs
#....................

Switch_1_2=Switch_1[grep("two", names(Switch_1))]  #since I am selecting for columns with "two" in the title, Length, Chrom, etc. already removed
dim(Switch_1_2)
#68 samples, 21761 genes --> for this dataset samples were run twice --> need to be combined for analysis
Meta_1_2=subset(Meta_1, Meta_1$Time=="2")
nrow(Meta_1_2) #68

Groups_1_2=factor(paste(Meta_1_2$Host1, Meta_1_2$Host2, Meta_1_2$Time, sep="."))
Meta_1_2=cbind(Meta_1_2, Group=Groups_1_2)

#Switch 1 17hrs
#................
Switch_1_17=Switch_1[grep("seventeen", names(Switch_1))]
dim(Switch_1_17)
#70 samples, 21761 genes (dataset 1 needs to be combined --> sum tech reps!)
Meta_1_17=subset(Meta_1, Meta_1$Time=="17")
nrow(Meta_1_17) #70
Groups_1_17=factor(paste(Meta_1_17$Host1, Meta_1_17$Host2, Meta_1_17$Time, sep="."))
Meta_1_17=cbind(Meta_1_17, Group=Groups_1_17)

#Switch 2 2hrs
#................
Switch_2_2=Switch_2[grep("two", names(Switch_2))] #since I am selecting for columns with "two" in the title, Length, Chrom, etc. already removed
ncol(Switch_2_2) #34
dim(Switch_2_2)# 21761    34
Meta_2_2=subset(Meta_2, Meta_2$Time=="2")
nrow(Meta_2_2)#34
Groups_2_2=factor(paste(Meta_2_2$Host1, Meta_2_2$Host2, Meta_2_2$Time, sep="."))
table(Groups_2_2)
Meta_2_2=cbind(Meta_2_2, Group=Groups_2_2)

#Switch 2 17hrs
#..................
Switch_2_17=Switch_2[grep("seventeen", names(Switch_2))]
ncol(Switch_2_17) #35
dim(Switch_2_17) #21761    35
Meta_2_17=subset(Meta_2, Meta_2$Time=="17")
nrow(Meta_2_17)#35
Groups_2_17=factor(paste(Meta_2_17$Host1, Meta_2_17$Host2, Meta_2_17$Time, sep="."))
table(Groups_2_17)
Meta_2_17=cbind(Meta_2_17, Group=Groups_2_17)

```

#4. Make differential gene expression list (DGE-list)

edgeR stores data in a list; main components: counts (=matrix that contains integer counts), sample (containing info about samples or libraries), genes (annotation)

```{r make DGE list}

#Switch 1
#............
#for the first Switch experiment samples were sequenced twice, so we need to combine the samples. For this a separate Sample name was created in the meta-list. 
#After creating the list, these samples can be combined based on common ID using sumTechReps 

#2hrs
S1_list_2=DGEList(counts=Switch_1_2, samples=Meta_1_2, group=Groups_1_2, genes=rownames(Switch_1_2))
list_1_2=sumTechReps(S1_list_2, ID=Meta_1_2$SampleID)
nrow(list_1_2) #21761

#17hrs
S1_list_17=DGEList(counts=Switch_1_17, samples=Meta_1_17, group=Groups_1_17, genes=rownames(Switch_1_17))
list_1_17=sumTechReps(S1_list_17, ID=Meta_1_17$SampleID)
nrow(list_1_17)

#plot library size Switch 1

par(mfrow=c(2,2))

libs12=barplot(list_1_2$samples$lib.size*1e-6, names=1:ncol(list_1_2), ylim=c(0,30), ylab="library size in mio", xlab="sample", main="library size Switch A - 2hrs", cex.lab=1.3, cex.main=1.5 )
mean(list_1_2$samples$lib.size*1e-6) #18.62058

libs117=barplot(list_1_17$samples$lib.size*1e-6, names=1:ncol(list_1_17), ylim=c(0,30), ylab="library size in mio", xlab="sample", main="library size Switch A - 17hrs", cex.lab=1.3, cex.main=1.5 )
mean(list_1_17$samples$lib.size*1e-6) #15.65437



##Switch 2
#...........

#2hrs
list_2_2=DGEList(counts=Switch_2_2, samples=Meta_2_2, group=Groups_2_2, genes=rownames(Switch_2_2))
list_2_2$samples
colnames(list_2_2)=Meta_2_2$Sample
nrow(list_2_2) #21761
list_2_2$samples$Host1=as.factor(list_2_2$samples$Host1)
#list_2_2$samples$Host1=relevel(list_2_2$samples$Host1, ref="N")

#17hrs
list_2_17=DGEList(counts=Switch_2_17, samples=Meta_2_17, group=Groups_2_17, genes=rownames(Switch_2_17))
list_2_17$samples
colnames(list_2_17)=Meta_2_17$Sample
nrow(list_2_17)

#plot librarysize
libs22=barplot(list_2_2$samples$lib.size*1e-6, names=1:ncol(list_2_2), ylim=c(0,30), ylab="library size in mio", xlab="sample", main="library size Switch B - 2hrs", cex.lab=1.3, cex.main=1.5 )
mean(list_2_2$samples$lib.size*1e-6) #20.27938
libs217=barplot(list_2_17$samples$lib.size*1e-6, names=1:ncol(list_2_17), ylim=c(0,30), ylab="library size in mio", xlab="sample", main="library size Switch B - 17hrs", cex.lab=1.3, cex.main=1.5 )
mean(list_2_17$samples$lib.size*1e-6) #22.34895
```


#5. filtering and normalization 

```{r filter and normalization}
# https://www.biostars.org/p/430936/

Group_1_2=list_1_2$samples$Group

Group_1_17=list_1_17$samples$Group
Group_2_2=list_2_2$samples$Group
Group_2_17=list_2_17$samples$Group

#Switch 1
#..........

#2hrs
keep_1_2=filterByExpr(list_1_2, group=Group_1_2)
table(keep_1_2)
# FALSE  TRUE
# 8635 13101
list_1_2_filt=list_1_2[keep_1_2, keep.lib.sizes=FALSE]  #FALSE because library size is different for different samples

list_1_2_norm = calcNormFactors(list_1_2_filt)
list_1_2_filt$samples
nrow(list_1_2_norm) #13101

#17hrs
keep_1_17=filterByExpr(list_1_17, group=Group_1_17)
table(keep_1_17)
# FALSE  TRUE
# 8950 13075
list_1_17_filt=list_1_17[keep_1_17, keep.lib.sizes=FALSE]  #FALSE because library size is different for different samples
nrow(list_1_17_filt) #13075

list_1_17_norm = calcNormFactors(list_1_17_filt)
list_1_17_filt$samples
nrow(list_1_17_norm) #13075


#Switch 2
#..........

#2hrs
keep_2_2=filterByExpr(list_2_2, group=Group_2_2)
table(keep_2_2)
# FALSE  TRUE
#  8905 12856
list_2_2_filt=list_2_2[keep_2_2, keep.lib.sizes=FALSE]  #FALSE because library size is different for different samples
nrow(list_2_2_filt) #12856

list_2_2_norm = calcNormFactors(list_2_2_filt)
list_2_2_filt$samples
nrow(list_2_2_norm) #12856


#17hrs
keep_2_17=filterByExpr(list_2_17, group=Group_2_17)
table(keep_2_17)
# FALSE  TRUE
# 8851 12910
list_2_17_filt=list_2_17[keep_2_17, keep.lib.sizes=FALSE]  #FALSE because library size is different for different samples
nrow(list_2_17_filt) #12910

list_2_17_norm = calcNormFactors(list_2_17_filt)
list_2_17_filt$samples
nrow(list_2_17_norm)#12910
```


##look at filtered and normalized data

```{r preparation for data visualization}

#color palette
#Switch 1
cb1 <- c("#4b1f6f", "#E69F00" ,"#999999" ) #for host plants;  E=purple , N=yellow, S=grey
cb2<- c("#000000", "#117733" ,"#DDCC77", "#88CCEE") #family for Switch 1
#Switch 2
cbfam <- c("#007C7C", "#D55E00" ) # family
cb3 <- c("#0072B2" ,"#999999", "#E69F00" ) #for host plants;  G=blue, N=yellow, S=grey
```


### data preparation for visualization
Data were pseudo-log transformed

```{r data transformation}

#Switch 1
pseudo_1_2=log2(cpm(list_1_2_norm)+1)
pseudo_1_17=log2(cpm(list_1_17_norm)+1)

#Switch 2
pseudo_2_2=log2(cpm(list_2_2_norm)+1)
pseudo_2_17=log2(cpm(list_2_17_norm)+1)


```


###Run and plot PCAs - Switch 1

```{r PCAs}

#Switch 1_2

PCA_1_2_norm=prcomp(t(pseudo_1_2))
PCA_1_2_norm_summary <- as.data.frame(t((summary(PCA_1_2_norm))$importance)) %>% rownames_to_column(var = "Principal Component")
Scree_1_2_norm=
  ggplot(PCA_1_2_norm_summary[1:6,], aes( x= `Principal Component`, y = `Proportion of Variance`)) +
  geom_bar(stat = "identity", linewidth=2, fill="grey") +
  labs(x = "Principal component", "Prop. of variance") +
  geom_label(stat = "identity", aes(label = round(`Proportion of Variance`, 3)),position = position_stack(vjust = 1.0),size = 3) +
  theme_classic(base_size = 10)+
  ylim(0,0.45)

summary(PCA_1_2_norm)
#--> PC 1: 32.3%, PC 2: 12.6%, PC3=6.3%


PCA_1_2_norm_df=data.frame(PCA_1_2_norm$x) #PCs are columns, samples are rows
PCA_1_2_norm_df$plotx=PCA_1_2_norm_df[,1]
PCA_1_2_norm_df$ploty=PCA_1_2_norm_df[,2]
PCA_1_2_norm_df$plotz=PCA_1_2_norm_df[,3]

#add Hostinformation
#Switch 1 had technical replicates, these had to be combined in the DGE list, for the PCA, a separate 
#Sheet with Meta_information, in which technical replicates were combined, was used ()
Meta_1_2_sumRep=read.delim("/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Meta_2_notechrep.txt")
Meta_1_2_sumRep$Name=factor(paste(Meta_1_2_sumRep$Host1, Meta_1_2_sumRep$Host2, sep="."))
Host=as.character(Meta_1_2_sumRep$Name)
Host1_12=as.character(Meta_1_2_sumRep$Host1)
Host2_12=as.character(Meta_1_2_sumRep$Host2)
Fam=as.character(Meta_1_2_sumRep$BioRep)
PCA_1_2_norm_df$Family=Fam
#PCA_1_2_norm_df$Host1=Host1
#PCA_1_2_norm_df$Host2=Host2


Norm_1_2_12=ggplot(PCA_1_2_norm_df, aes(x=plotx,y=ploty))+
  #geom_label()+
  geom_point(size=4, aes(shape=Host1_12, color=Host2_12), alpha=0.5)+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch A - 2hrs -filt")+
 # xlab ("PC1 [40.6%]") + ylab("PC2 [14.7%]")+
  xlab ("PC1 [32.3%]") + ylab("PC2 [12.6%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cb1)+
  #scale_shape_manual(values=c(17, 19, 15))
  scale_shape_manual(values=c(15, 16, 17))+
  #stat_ellipse()
  theme(legend.position = "none")
  #labs(color="second Host")

Norm_1_2_13=ggplot(PCA_1_2_norm_df, aes(x=plotx,y=plotz))+
  geom_point(size=4, aes(shape=Host1_12, color=Host2_12), alpha=0.5)+
 # geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch A - 2hrs -filt ")+
  xlab ("PC1 [32.3%]") + ylab("PC3 [6.3%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cb1)+
  scale_shape_manual(values=c(15, 16, 17))+
theme(legend.position = "none")

Norm_1_2_12_fam=ggplot(PCA_1_2_norm_df, aes(x=plotx,y=ploty))+
  geom_point(size=4, aes(color=Family), alpha=0.5)+
  labs(color="Family")+
 # geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch A - 2hrs -filt")+
  #xlab ("PC1 [40.6%]") + ylab("PC2 [14.7%]")+
  xlab ("PC1 [32.3%]") + ylab("PC2 [12.6%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cb2)+
  theme(legend.position = "none")


Norm_1_2_13_fam=ggplot(PCA_1_2_norm_df, aes(x=plotx,y=plotz))+
  geom_point(size=4, aes(color=Family), alpha=0.5)+
  labs(color="Family")+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ( "Switch A - 2hrs -filt")+
  xlab ("PC1 [32.3%]") + ylab("PC3 [6.3%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cb2)+
  labs(color="Family")+
  theme(legend.position = "none")


#Switch 1_17
PCA_1_17_norm=prcomp(t(pseudo_1_17))

PCA_1_17_norm_summary <- as.data.frame(t((summary(PCA_1_17_norm))$importance)) %>% rownames_to_column(var = "Principal Component")
Scree_1_17_norm=
  ggplot(PCA_1_17_norm_summary[1:6,], aes( x= `Principal Component`, y = `Proportion of Variance`)) +
  geom_bar(stat = "identity", size=2, fill="grey") +
  labs(x = "Principal component", "Prop. of variance") +
  geom_label(stat = "identity", aes(label = round(`Proportion of Variance`, 3)),position = position_stack(vjust = 1.0),size = 3) +
  theme_classic(base_size = 10)+
  ylim(0,0.45)

  summary(PCA_1_17_norm)
  #--> PC 1: 39.8%, PC 2: 8.1%, PC3=7.4%

PCA_1_17_norm_df=data.frame(PCA_1_17_norm$x) #PCs are columns, samples are rows
PCA_1_17_norm_df$plotx=PCA_1_17_norm_df[,1]
PCA_1_17_norm_df$ploty=PCA_1_17_norm_df[,2]
PCA_1_17_norm_df$plotz=PCA_1_17_norm_df[,3]


#add Hostinformation
#Switch 1 had technical replicates, these had to be combined in the DGE list, for the PCA, a separate 
#Sheet with Meta_information, in which technical replicates were combined, was used ()
Meta_1_17_sumRep=read.delim("/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Meta_1_17_notechrep.txt")
Meta_1_17_sumRep$Name=factor(paste(Meta_1_17_sumRep$Host1, Meta_1_17_sumRep$Host2, sep="."))
Host=as.character(Meta_1_17_sumRep$Name)
Host1_17=as.character(Meta_1_17_sumRep$Host1)
Host2_17=as.character(Meta_1_17_sumRep$Host2)
PCA_1_17_norm_df$Name=Host
Fam=as.character(Meta_1_17_sumRep$BioRep)
PCA_1_17_norm_df$Family=Fam
# PCA_1_17_norm_df$Host1=Host1
# PCA_1_17_norm_df$Host2=Host2


Norm_1_17_12=ggplot(PCA_1_17_norm_df, aes(x=plotx,y=ploty))+
  #geom_label()
  geom_point(size=4, aes(shape=Host1_17, color=Host2_17), alpha=0.5)+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch A - 17hrs -filt")+
 # xlab ("PC1 [46.2%]") + ylab("PC2 [11.1%]")+
  xlab ("PC1 [39.8%]") + ylab("PC2 [8.1%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cb1)+
  scale_shape_manual(values=c(15, 16, 17))+
  theme(legend.position = "none")
  #labs(color="second Host")

Norm_1_17_13=ggplot(PCA_1_17_norm_df, aes(x=plotx,y=plotz))+
  geom_point(size=4, aes(shape=Host1_17, color=Host2_17), alpha=0.5)+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch A - 17hrs -filt")+
  xlab ("PC1 [39.8%]") + ylab("PC3 [7.4%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cb1)+
  scale_shape_manual(values=c(15, 16, 17))+
    theme(legend.position = "none")

Norm_1_17_12_fam=ggplot(PCA_1_17_norm_df, aes(x=plotx,y=ploty))+
  geom_point(size=4, aes(color=Family), alpha=0.5)+
  labs(color="Family")+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch A - 17 hrs -filt")+
  #xlab ("PC1 [46.2%]") + ylab("PC2 [11.1%]")+
  xlab ("PC1 [39.8%]") + ylab("PC2 [8.1%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cb2)+
  theme(legend.position = "none")+
  labs(color="Family")

Norm_1_17_13_fam=ggplot(PCA_1_17_norm_df, aes(x=plotx,y=plotz))+
  geom_point(size=4, aes(color=Family), alpha=0.5)+
  labs(color="Family")+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ( "Switch A - 17hrs -filt")+
  xlab ("PC1 [39.8%]") + ylab("PC3 [7.4%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cb2)+
  labs(color="Family")+
  theme(legend.position = "none")

```

###Run and plot PCAs - Switch 2

```{r PCAs}

#Switch 2_2

PCA_2_2_norm=prcomp(t(pseudo_2_2))

PCA_2_2_norm_summary <- as.data.frame(t((summary(PCA_2_2_norm))$importance)) %>% rownames_to_column(var = "Principal Component")
Scree_2_2_norm=
  ggplot(PCA_2_2_norm_summary[1:6,], aes( x= `Principal Component`, y = `Proportion of Variance`)) +
  geom_bar(stat = "identity", size=2, fill="grey") +
  labs(x = "Principal component", "Prop. of variance") +
  geom_label(stat = "identity", aes(label = round(`Proportion of Variance`, 3)),position = position_stack(vjust = 1.0),size = 3) +
  theme_classic(base_size = 10)+
  ylim(0,0.45)

  summary(PCA_2_2_norm)
  #--> PC 1: 17.6%, PC 2: 13.1%, PC3=9.2%

PCA_2_2_norm_df=data.frame(PCA_2_2_norm$x) #PCs are columns, samples are rows
PCA_2_2_norm_df$plotx=PCA_2_2_norm_df[,1]
PCA_2_2_norm_df$ploty=PCA_2_2_norm_df[,2]
PCA_2_2_norm_df$plotz=PCA_2_2_norm_df[,3]


#add Hostinformation
Meta_2_2$Name=factor(paste(Meta_2_2$Host1_lat, Meta_2_2$Host2_lat, sep="."))
Host=as.character(Meta_2_2$Name)
Host1_2_2=as.character(Meta_2_2$Host1_lat)
Host2_2_2=as.character(Meta_2_2$Host2_lat)
PCA_2_2_norm_df$Name=Host
Fam=as.character(Meta_2_2$Family)
PCA_2_2_norm_df$Family=Fam
# PCA_2_2_norm_df$Host1=Host1
# PCA_2_2_norm_df$Host2=Host2

Norm_2_2_12=ggplot(PCA_2_2_norm_df, aes(x=plotx,y=ploty))+
  geom_point(size=4, aes(shape=Host1_2_2, color=Host2_2_2), alpha=0.5)+
  labs(color="Host2")+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch B - 2hrs -filt")+
  #xlab ("PC1 [25.5%]") + ylab("PC2 [12.4%]")+
  xlab ("PC1 [17.6%]") + ylab("PC2 [13.1%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cb3)+
 # theme(legend.position = "none")+
  labs(color="second Host")+
  scale_shape_manual(values=c(18, 17,16))


Norm_2_2_13=ggplot(PCA_2_2_norm_df, aes(x=plotx,y=plotz))+
  geom_point(size=4, aes(shape=Host1_2_2, color=Host2_2_2),alpha=0.5)+
  labs(color="Host2")+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch B -  2hrs -filt")+
  xlab ("PC1 [17.6%]") + ylab("PC3 [9.2%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cb3)+
  labs(color="second Host")+
  scale_shape_manual(values=c(18, 17,16))+
  theme(legend.position = "none")

Norm_2_2_12_fam=ggplot(PCA_2_2_norm_df, aes(x=plotx,y=ploty))+
  geom_point(size=4, aes(color=Family), alpha=0.5)+
  labs(color="Family")+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch B - 2hrs -filt")+
  #xlab ("PC1 [25.5%]") + ylab("PC2 [12.4%]")+
  xlab ("PC1 [17.6%]") + ylab("PC2 [13.1%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cbfam)+
  theme(legend.position = "none")+
  labs(color="Family")

Norm_2_2_13_fam=ggplot(PCA_2_2_norm_df, aes(x=plotx,y=plotz))+
  geom_point(size=4, aes(color=Family), alpha=0.5)+
  labs(color="Family")+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ( "Switch B - 2hrs -filt")+
  xlab ("PC1 [17.6%]") + ylab("PC3 [9.2%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cbfam)+
  labs(color="Family")+
  theme(legend.position = "none")


#Switch 2_17

PCA_2_17_norm=prcomp(t(pseudo_2_17))

PCA_2_17_norm_summary <- as.data.frame(t((summary(PCA_2_17_norm))$importance)) %>% rownames_to_column(var = "Principal Component")
Scree_2_17_norm=
  ggplot(PCA_2_17_norm_summary[1:6,], aes( x= `Principal Component`, y = `Proportion of Variance`)) +
  geom_bar(stat = "identity", size=2, fill="grey") +
  labs(x = "Principal component", "Prop. of variance") +
  geom_label(stat = "identity", aes(label = round(`Proportion of Variance`, 3)),position = position_stack(vjust = 1.0),size = 3) +
  theme_classic(base_size = 10)+
  ylim(0,0.45)

  summary(PCA_2_17_norm)
  #--> PC 1: 16.3%, PC 2: 14.1%, PC3=10.4%

PCA_2_17_norm_df=data.frame(PCA_2_17_norm$x) #PCs are columns, samples are rows
PCA_2_17_norm_df$plotx=PCA_2_17_norm_df[,1]
PCA_2_17_norm_df$ploty=PCA_2_17_norm_df[,2]
PCA_2_17_norm_df$plotz=PCA_2_17_norm_df[,3]


#add Hostinformation
Meta_2_17$Name=factor(paste(Meta_2_17$Host1_lat, Meta_2_17$Host2_lat, sep="."))
Host=as.character(Meta_2_17$Name)
Host1_2_17=as.character(Meta_2_17$Host1_lat)
Host2_2_17=as.character(Meta_2_17$Host2_lat)
PCA_2_17_norm_df$Name=Host
Fam=as.character(Meta_2_17$Family)
PCA_2_17_norm_df$Family=Fam
# PCA_2_17_norm_df$Host1=Host1
# PCA_2_17_norm_df$Host2=Host2

Norm_2_17_12=ggplot(PCA_2_17_norm_df, aes(x=plotx,y=ploty))+
  geom_point(size=4, aes(shape=Host1_2_17, color=Host2_2_17), alpha=0.5)+
  labs(color="Host2")+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch B - 17hrs -filt")+
  #xlab ("PC1 [25.1%]") + ylab("PC2 [12.5%]")+
  xlab ("PC1 [16.3%]") + ylab("PC2 [14.1%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cb3)+
  theme(legend.position = "none")+
  labs(color="second Host")+
  scale_shape_manual(values=c(18, 17,16))

Norm_2_17_13=ggplot(PCA_2_17_norm_df, aes(x=plotx,y=plotz))+
  geom_point(size=4, aes(shape=Host1_2_17, color=Host2_2_17), alpha=0.5)+
  labs(color="Host2")+
 # geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch B - 17hrs -filt")+
  xlab ("PC1 [16.3%]") + ylab("PC3 [10.4%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cb3)+
  labs(color="second Host")+
  scale_shape_manual(values=c(18, 17,16))+
  theme(legend.position = "none")

Norm_2_17_12_fam=ggplot(PCA_2_17_norm_df, aes(x=plotx,y=ploty))+
  geom_point(size=4, aes(color=Family), alpha=0.5)+
  #labs(color="Family")+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch B - 17hrs -filt")+
 # xlab ("PC1 [25.1%]") + ylab("PC2 [12.5%]")+
  xlab ("PC1 [16.3%]") + ylab("PC2 [14.1%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cbfam)+
  theme(legend.position = "none")+
  labs(color="Family")

Norm_2_17_13_fam=ggplot(PCA_2_17_norm_df, aes(x=plotx,y=plotz))+
  geom_point(size=4, aes(color=Family), alpha=0.5)+
  labs(color="Family")+
  #geom_text(hjust=0, vjust=0)+
  ggtitle ("Switch B - 17hrs -filt")+
  xlab ("PC1 [16.3%]") + ylab("PC3 [10.4%]")+
  coord_fixed(ratio=1)+
  theme_bw()+theme(aspect.ratio = 1)+theme(panel.grid=element_blank())+
  scale_colour_manual(values=cbfam)+
  labs(color="Family")+
  theme(legend.position = "none")
```


### plot PCAs together
```{r produce plots}
#plot filtered PCA
#PC1-2
#------
g1=ggarrange(Scree_1_2_norm, Scree_1_17_norm,
          Norm_1_2_12, Norm_1_17_12,
          Norm_1_2_12_fam, Norm_1_17_12_fam,
          ncol=2, nrow=3)

g2=ggarrange( Scree_2_2_norm, Scree_2_17_norm,
              Norm_2_2_12, Norm_2_17_12,
              Norm_2_2_12_fam, Norm_2_17_12_fam,
             ncol=2, nrow=3)

PC12_legend=ggarrange(g1, g2,ncol=2, nrow=1)

PC12=ggarrange(g1, g2,ncol=2, nrow=1)
plot(PC12)

#ggsave("/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/Figures_edited/PC12.pdf", PC12, device="pdf", height=25, width=32, units="cm", dpi=700)

#PC1-3
#---------

g3=ggarrange(Scree_1_2_norm, Scree_1_17_norm,
             Norm_1_2_13, Norm_1_17_13,
             Norm_1_2_13_fam, Norm_1_17_13_fam,
             ncol=2, nrow=3)

g4=ggarrange( Scree_2_2_norm, Scree_2_17_norm,
              Norm_2_2_13, Norm_2_17_13,
              Norm_2_2_13_fam, Norm_2_17_13_fam,
              ncol=2, nrow=3)

PC13=ggarrange(g3, g4)
plot(PC13)

#ggsave("/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/Figures_edited/PC13.pdf", PC13, device="pdf", height=25, width=32, units="cm", dpi=700)
```


###look at filtered and normalized data with ComplexHeatmap

for this we will also use the pseudo-log transformed dataset
we will look at the top 5000 most variant genes

```{r Complex heatmap}

top_var.12 <-names(sort(apply(pseudo_1_2, 1, var), decreasing = T))[1:5000]
top_var.12 <- (pseudo_1_2[top_var.12,])

top_var.117 <-names(sort(apply(pseudo_1_17, 1, var), decreasing = T))[1:5000]
top_var.117 <- (pseudo_1_17[top_var.117,]) 

top_var.22 <-names(sort(apply(pseudo_2_2, 1, var), decreasing = T))[1:5000]
top_var.22 <- (pseudo_2_2[top_var.22,])

top_var.217 <-names(sort(apply(pseudo_2_17, 1, var), decreasing = T))[1:5000]
top_var.217 <- (pseudo_2_17[top_var.217,])

cor_matrix_1_2=cor((top_var.12), method=c("pearson"))
cor_matrix_1_17=cor((top_var.117), method=c("pearson"))
cor_matrix_2_2=cor((top_var.22),  method=c("pearson"))
cor_matrix_2_17=cor((top_var.217),  method=c("pearson"))

``` 

####prepare color palette
```{r color palette for heatmap}
nb.cols <- 18
mycolorsS1 <- colorRampPalette(brewer.pal(8, "Purples"))(nb.cols)
nb.cols <- 18
mycolorsS2 <- colorRampPalette(brewer.pal(8, "Blues"))(nb.cols)
```

####Create Heatmaps

```{r Switch 1 - annotation and heatmap}

#Switch 1
# 2 hours

#add annotation

# Ensure all sample names in the correlation matrix are present in the metadata
sample_names_matrix_1_2 <- colnames(cor_matrix_1_2)
sample_names_metadata_1_2 <- Meta_1_2_sumRep$SampleID

if (!all(sample_names_matrix_1_2 %in% sample_names_metadata_1_2)) {
  stop("Some sample names in the correlation matrix are not found in the metadata.")
}

# Ensure the order of samples in the metadata matches the order in the correlation matrix
metadata_1_2 <- Meta_1_2_sumRep[match(sample_names_matrix_1_2, Meta_1_2_sumRep$SampleID), ]

#make Host1 annotation
Host1_names_1_2 <-Meta_1_2_sumRep$Host1
unique_Host1_1_2=unique(Host1_names_1_2)
Host1_colors_1_2=setNames(setNames(c("#4b1f6f", "#E69F00", "#999999"), 0.5), unique_Host1_1_2)

# Verify the groups and create color mapping
group_names_1_2 <- Meta_1_2_sumRep$Host2
unique_groups_1_2 <- unique(group_names_1_2)
group_colors_1_2 <- setNames(alpha(c("#4b1f6f", "#E69F00", "#999999"), 0.5), unique_groups_1_2)

#make family annotation
family_names_1_2 <-Meta_1_2_sumRep$BioRep
unique_family_1_2=unique(family_names_1_2)
family_colors_1_2=setNames(alpha(c("#000000", "#117733" ,"#DDCC77", "#88CCEE"), 0.5), unique_family_1_2)

#Create the annotation object
sample_annotation_H1_1_2 <- HeatmapAnnotation(
  Group = Host1_names_1_2,
  col = list(Group = Host1_colors_1_2),
  annotation_name_side = "left",
  annotation_legend_param = list(
    Group = list(
      title = "Host 1"
    )))

#Create the annotation object
sample_annotation_1_2 <- HeatmapAnnotation(
  Group = group_names_1_2,
  col = list(Group = group_colors_1_2),
  annotation_name_side = "left",
  annotation_legend_param = list(
    Group = list(
      title = "Host 2"
)))


# Create the Family annotation
family_annotation_1_2 <- HeatmapAnnotation(
  Family = family_names_1_2,
  col = list(Family = family_colors_1_2),
  annotation_name_side = "left",
  annotation_legend_param = list(
    Family = list(
      title = "Family"
    ))
)

# Combine the annotations
combined_annotations_1_2 <- HeatmapAnnotation(
  df = data.frame(Host1=Host1_names_1_2, Group = group_names_1_2, Family = family_names_1_2),
  col = list(Host1=Host1_colors_1_2, Group = group_colors_1_2, Family = family_colors_1_2),
  which = "column",
  annotation_name_side = "left"
)

#Heatmap_1_2
H1=Heatmap(cor_matrix_1_2,
          col =  mycolorsS1,
          #col = viridis_pal,
          cluster_columns = TRUE,
          cluster_rows = TRUE,
          show_heatmap_legend = TRUE,
          column_title = "Switch A - 2 hours",
          name="Correlation coefficient",
          column_title_gp=grid::gpar(fontsize=16),
          row_dend_width = unit(2.5, "cm"),
          column_dend_height = unit(2.5, "cm"),
          top_annotation = combined_annotations_1_2)




#Switch 1
#17 hours

#add annotation

# Ensure all sample names in the correlation matrix are present in the metadata
sample_names_matrix_1_17 <- colnames(cor_matrix_1_17)
sample_names_metadata_1_17 <- Meta_1_17_sumRep$SampleID

    if (!all(sample_names_matrix_1_17 %in% sample_names_metadata_1_17)) {
        stop("Some sample names in the correlation matrix are not found in the metadata.")
        }

# Ensure the order of samples in the metadata matches the order in the correlation matrix
metadata_1_17 <- Meta_1_17_sumRep[match(sample_names_matrix_1_17, Meta_1_17_sumRep$SampleID), ]

#make Host1 annotation
Host1_names_1_17 <-Meta_1_17_sumRep$Host1
unique_Host1_1_17=unique(Host1_names_1_17)
Host1_colors_1_17=setNames(setNames(c("#4b1f6f", "#E69F00", "#999999"), 0.5), unique_Host1_1_17)

# Verify the groups and create color mapping
group_names_1_17 <- Meta_1_17_sumRep$Host2
unique_groups_1_17 <- unique(group_names_1_17)
group_colors_1_17 <- setNames(alpha(c("#4b1f6f", "#E69F00", "#999999"), 0.5), unique_groups_1_17)

family_names_1_17 <-Meta_1_17_sumRep$BioRep
unique_family_1_17=unique(family_names_1_17)
family_colors_1_17=setNames(alpha(c("#000000", "#117733" ,"#DDCC77", "#88CCEE"), 0.5), unique_family_1_17)

#Create the annotation object - Host 1
sample_annotation_H1_1_17 <- HeatmapAnnotation(
Group = Host1_names_1_17,
col = list(Group = Host1_colors_1_17),
annotation_name_side = "left",
annotation_legend_param = list(
Group = list(title = "Host 1"
              )))

#Create the annotation object - Host 2
sample_annotation_1_17 <- HeatmapAnnotation(
Group = group_names_1_17,
col = list(Group = group_colors_1_17),
annotation_name_side = "left",
annotation_legend_param = list(
              Group = list(
                title = "Host 2"
              )))

# Create the Family annotation
  family_annotation_1_17 <- HeatmapAnnotation(
  Family = family_names_1_17,
  col = list(Family = family_colors_1_17),
  annotation_name_side = "left",
  annotation_legend_param = list(
  Family = list(
  title = "Family"
        )))


# Combine the annotations
combined_annotations_1_17 <- HeatmapAnnotation(
df = data.frame(Host1=Host1_names_1_17, Group = group_names_1_17, Family = family_names_1_17),
col = list(Host1=Host1_colors_1_17, Group = group_colors_1_17, Family = family_colors_1_17), 
which = "column",annotation_name_side = "left")

#Heatmap_1_17
H2=Heatmap(cor_matrix_1_17,
          col =  mycolorsS1,
          cluster_columns = TRUE,
          cluster_rows = TRUE,
          show_heatmap_legend = TRUE,
          column_title = "Switch A - 17 hours",
          name="Correlation coefficient",
          column_title_gp=grid::gpar(fontsize=16),
          row_dend_width = unit(2.5, "cm"),
          column_dend_height = unit(2.5, "cm"),
          top_annotation = combined_annotations_1_17)

```


```{r Switch 2 - annotation and heatmap}


#Switch 2
#2 hours

#add annotation
# Ensure all sample names in the correlation matrix are present in the metadata
sample_names_matrix_2_2 <- colnames(cor_matrix_2_2)
sample_names_metadata_2_2 <- Meta_2_2$Sample
if (!all(sample_names_matrix_2_2 %in% sample_names_metadata_2_2)) {
stop("Some sample names in the correlation matrix are not found in the metadata.")
 }

# Ensure the order of samples in the metadata matches the order in the correlation matrix
metadata_2_2 <- Meta_2_2[match(sample_names_matrix_2_2, Meta_2_2$Sample), ]

#make Host1 annotation
Host1_names_2_2 <-Meta_2_2$Host1
unique_Host1_2_2=unique(Host1_names_2_2)
Host1_colors_2_2=setNames(c("#0072B2", "#E69F00", "#999999"), unique_Host1_2_2)

# Verify the groups and create color mapping
group_names_2_2 <- Meta_2_2$Host2
unique_groups_2_2 <- unique(group_names_2_2)
group_colors_2_2 <- setNames(alpha(c("#0072B2", "#E69F00", "#999999"), 0.5), unique_groups_2_2)
family_names_2_2 <-Meta_2_2$Family
unique_family_2_2=unique(family_names_2_2)
family_colors_2_2=setNames(alpha(c("#007C7C", "#D55E00"), 0.5), unique_family_2_2)

#Create the annotation object
sample_annotation_H1_2_2 <- HeatmapAnnotation(
    Group = Host1_names_2_2,
    col = list(Group = Host1_colors_2_2),
    annotation_name_side = "left",
    annotation_legend_param = list(
    Group = list(
    title = "Host 1"
     )))


sample_annotation_2_2 <- HeatmapAnnotation(
  Group = group_names_2_2,
  col = list(Group = group_colors_2_2),
  annotation_name_side = "left",
  annotation_legend_param = list(
  Group = list(
  title = "Host 2"
   )))


# Create the Family annotation
family_annotation_2_2 <- HeatmapAnnotation(
Family = family_names_2_2,
col = list(Family = family_colors_2_2),
annotation_name_side = "left",
annotation_legend_param = list(
Family = list(title = "Family")))

# Combine the annotations
combined_annotations_2_2 <- HeatmapAnnotation(
df = data.frame(Host1= Host1_names_2_2, Group = group_names_2_2, Family = family_names_2_2),
col = list(Host1=Host1_colors_2_2, Group = group_colors_2_2, Family = family_colors_2_2),
which = "column",
annotation_name_side = "left")

#Heatmap_2_2
H3=Heatmap(cor_matrix_2_2,
  col =  mycolorsS2,
  cluster_columns = TRUE,
  cluster_rows = TRUE,
  show_heatmap_legend = TRUE, 
  column_title = "Switch B - 2 hours",
  name="Correlation coefficient",
  column_title_gp=grid::gpar(fontsize=16),
  row_dend_width = unit(2.5, "cm"),
  column_dend_height = unit(2.5, "cm"),                                
  top_annotation = combined_annotations_2_2)


# Switch 2
# 17 hours


#add annotation
# Ensure all sample names in the correlation matrix are present in the metadata
sample_names_matrix_2_17 <- colnames(cor_matrix_2_17)
sample_names_metadata_2_17 <- Meta_2_17$Sample
if (!all(sample_names_matrix_2_17 %in% sample_names_metadata_2_17)) {
  stop("Some sample names in the correlation matrix are not found in the metadata.")
                                }
# Ensure the order of samples in the metadata matches the order in the correlation matrix
metadata_2_17 <- Meta_2_17[match(sample_names_matrix_2_17, Meta_2_17$Sample), ]

#make Host1 annotation
Host1_names_2_17 <-Meta_2_17$Host1
unique_Host1_2_17=unique(Host1_names_2_17)
Host1_colors_2_17=setNames(c("#0072B2", "#E69F00", "#999999"), unique_Host1_2_17)


# Verify the groups and create color mapping
group_names_2_17 <- Meta_2_17$Host2
unique_groups_2_17 <- unique(group_names_2_17)
group_colors_2_17 <- setNames(alpha(c("#0072B2", "#E69F00", "#999999"), 0.5), unique_groups_2_17)
family_names_2_17 <-Meta_2_17$Family                               
unique_family_2_17=unique(family_names_2_17)
family_colors_2_17=setNames(alpha(c("#007C7C", "#D55E00"), 0.5), unique_family_2_17)


#Create the annotation object
sample_annotation_H1_2_17 <- HeatmapAnnotation(
Group = Host1_names_2_17,
col = list(Group = Host1_colors_2_17),
annotation_name_side = "left",
annotation_legend_param = list(
  Group = list(
    title = "Host 1"
    )))

#Create the annotation object
 sample_annotation_2_17 <- HeatmapAnnotation(
   Group = group_names_2_17,
   col = list(Group = group_colors_2_2),
   annotation_name_side = "left",
   annotation_legend_param = list(
     Group = list(
       title = "Host 2"
       )))


# Create the Family annotation
 family_annotation_2_17 <- HeatmapAnnotation(
   Family = family_names_2_17,
   col = list(Family = family_colors_2_17),
   annotation_name_side = "left",
   annotation_legend_param = list(
     Family = list(
       title = "Family"
       )))

# Combine the annotations
combined_annotations_2_17 <- HeatmapAnnotation(
df = data.frame(Host1= Host1_names_2_17, Group = group_names_2_17, Family = family_names_2_17),
col = list(Host1=Host1_colors_2_17, Group = group_colors_2_17, Family = family_colors_2_17),
 which = "column",annotation_name_side = "left")

#Heatmap_2_17
H4=Heatmap(cor_matrix_2_17,
           col =  mycolorsS2,
           cluster_columns = TRUE,
           cluster_rows = TRUE,
           show_heatmap_legend = TRUE,
           column_title = "Switch B - 17 hours",
           name="Correlation coefficient",
           column_title_gp=grid::gpar(fontsize=16),
           row_dend_width = unit(2.5, "cm"),
           column_dend_height = unit(2.5, "cm"),
           top_annotation = combined_annotations_2_17)
```

###plot heatmaps together

```{r plot heatmaps}

H1_ggplot <- as.ggplot(H1)
H2_ggplot <- as.ggplot(H2)
H3_ggplot <- as.ggplot(H3)
H4_ggplot <- as.ggplot(H4)

HM=grid.arrange(H1_ggplot, H2_ggplot, H3_ggplot, H4_ggplot, ncol = 2)
#ggsave("/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/Figures_edited/Heatmaps.top5000_bothHosts.pdf", HM, device="pdf", height=20, width=20)
```

#6. Look at effect of individual hosts plants and/or their interaction on the number of differentially expressed genes

````{r design model matrix}
# https://support.bioconductor.org/p/56568 + edgeR User guide

#Switch 1
#............
#2hrs
Host1_1_2 <- list_1_2$samples$Host1
Host2_1_2 <- list_1_2$samples$Host2
Family_1_2 <-list_1_2$samples$BioRep
Group_1_2<-list_1_2$samples$group

int_eff_des_1_2 <- model.matrix(~ 1 + Family_1_2 + Host1_1_2 * Host2_1_2 ) #Family included as a random effect
main_eff_des_1_2 <- model.matrix(~ 1 + Family_1_2+ Host1_1_2 + Host2_1_2)
group_des_1_2<-model.matrix(~0+Family_1_2+Group_1_2)


#17hrs
Host1_1_17 <- list_1_17$samples$Host1
Host2_1_17 <- list_1_17$samples$Host2
Family_1_17 <-list_1_17$samples$BioRep
Group_1_17<-list_1_17$samples$group

int_eff_des_1_17 <- model.matrix(~ 1 + Family_1_17 + Host1_1_17 * Host2_1_17)
main_eff_des_1_17 <- model.matrix(~ 1 + Family_1_17 + Host1_1_17 + Host2_1_17)
group_des_1_17<-model.matrix(~0+Family_1_17+Group_1_17)

#Switch 2
#............
#2hrs
Host1_2_2 <- list_2_2$samples$Host1
Host2_2_2 <- list_2_2$samples$Host2
Family_2_2 <-list_2_2$samples$Family
Group_2_2<-list_2_2$samples$group

int_eff_des_2_2 <- model.matrix(~ 1+ Family_2_2 + Host1_2_2 * Host2_2_2 )
main_eff_des_2_2 <- model.matrix(~ 1+ Family_2_2 + Host1_2_2 + Host2_2_2)
group_des_2_2<-model.matrix(~0+Family_2_2+Group_2_2)


#17hrs
Host1_2_17 <- list_2_17$samples$Host1
Host2_2_17 <- list_2_17$samples$Host2
Family_2_17 <-list_2_17$samples$Family
Group_2_17<-list_2_17$samples$group

int_eff_des_2_17 <- model.matrix(~ 1 + Family_2_17 + Host1_2_17 * Host2_2_17)
main_eff_des_2_17 <- model.matrix(~ 1 + Family_2_17 + Host1_2_17 + Host2_2_17)
group_des_2_17<-model.matrix(~0+Family_2_17+Group_2_17)

```

## Estimate Dispersion
```{r estimate dispersion}
#Switch 1
#..........

#2hrs
y.genes.filt.H1xH2_1_2 <- estimateDisp(list_1_2_norm, int_eff_des_1_2, robust=TRUE)
y.genes.filt.H1andH2_1_2 <- estimateDisp(list_1_2_norm, main_eff_des_1_2, robust=TRUE)

#17hrs
y.genes.filt.H1xH2_1_17 <- estimateDisp(list_1_17_norm, int_eff_des_1_17, robust=TRUE)
y.genes.filt.H1andH2_1_17 <- estimateDisp(list_1_17_norm, main_eff_des_1_17, robust=TRUE)

#Switch 2
#..........

#2hrs
y.genes.filt.H1xH2_2_2 <- estimateDisp(list_2_2_norm, int_eff_des_2_2, robust=TRUE)
y.genes.filt.H1andH2_2_2 <- estimateDisp(list_2_2_norm, main_eff_des_2_2, robust=TRUE)

#17hrs
y.genes.filt.H1xH2_2_17 <- estimateDisp(list_2_17_norm, int_eff_des_2_17, robust=TRUE)
y.genes.filt.H1andH2_2_17 <- estimateDisp(list_2_17_norm, main_eff_des_2_17, robust=TRUE)
```


## Quasilikelihood fit object
```{r QL-Fit}
#Switch 1
#...........

#2hrs
fit.genes.filt.H1xH2_1_2 <- glmQLFit(y.genes.filt.H1xH2_1_2,int_eff_des_1_2, robust=TRUE)
fit.genes.filt.H1andH2_1_2 <- glmQLFit(y.genes.filt.H1andH2_1_2,main_eff_des_1_2, robust=TRUE)

#17hrs
fit.genes.filt.H1xH2_1_17 <- glmQLFit(y.genes.filt.H1xH2_1_17,int_eff_des_1_17, robust=TRUE)
fit.genes.filt.H1andH2_1_17 <- glmQLFit(y.genes.filt.H1andH2_1_17,main_eff_des_1_17, robust=TRUE)

#Switch 2
#...........

#2hrs
fit.genes.filt.H1xH2_2_2 <- glmQLFit(y.genes.filt.H1xH2_2_2,int_eff_des_2_2, robust=TRUE)
fit.genes.filt.H1andH2_2_2 <- glmQLFit(y.genes.filt.H1andH2_2_2,main_eff_des_2_2, robust=TRUE)

#17hrs
fit.genes.filt.H1xH2_2_17 <- glmQLFit(y.genes.filt.H1xH2_2_17,int_eff_des_2_17, robust=TRUE)
fit.genes.filt.H1andH2_2_17 <- glmQLFit(y.genes.filt.H1andH2_2_17,main_eff_des_2_17, robust=TRUE)
```

## glmQLFTest

```{r FTest}
#Switch 1 
#............

#2hrs
qlf.genes.filt.H1xH2_interaction_1_2 <- glmQLFTest(fit.genes.filt.H1xH2_1_2, coef = 7:10 )
qlf.genes.filt.H1andH2_H1_1_2 <- glmQLFTest(fit.genes.filt.H1andH2_1_2, coef = 3:4)
qlf.genes.filt.H1andH2_H2_1_2 <- glmQLFTest(fit.genes.filt.H1andH2_1_2, coef = 5:6)

qlf.genes_1_2 <- list()

qlf.genes_1_2$interaction <- (topTags(qlf.genes.filt.H1xH2_interaction_1_2, Inf))$table %>% mutate(effect = "Int")
qlf.genes_1_2$H1 <- (topTags(qlf.genes.filt.H1andH2_H1_1_2, Inf))$table %>% mutate(effect = "H1")
qlf.genes_1_2$H2 <- (topTags(qlf.genes.filt.H1andH2_H2_1_2, Inf))$table %>% mutate(effect = "H2")

length(qlf.genes_1_2$H1$FDR[qlf.genes_1_2$H1$FDR< 0.05])#507
length(qlf.genes_1_2$H2$FDR[qlf.genes_1_2$H2$FDR< 0.05])#2
length(qlf.genes_1_2$interaction$FDR[qlf.genes_1_2$interaction$FDR< 0.05])#41


#17hrs
qlf.genes.filt.H1xH2_interaction_1_17 <- glmQLFTest(fit.genes.filt.H1xH2_1_17, coef = 7:10 )
qlf.genes.filt.H1andH2_H1_1_17 <- glmQLFTest(fit.genes.filt.H1andH2_1_17, coef = 3:4)
qlf.genes.filt.H1andH2_H2_1_17 <- glmQLFTest(fit.genes.filt.H1andH2_1_17, coef = 5:6)

qlf.genes_1_17 <- list()

qlf.genes_1_17$interaction <- (topTags(qlf.genes.filt.H1xH2_interaction_1_17, Inf, adjust.method="BH"))$table %>% mutate(effect = "Int")
qlf.genes_1_17$H1 <- (topTags(qlf.genes.filt.H1andH2_H1_1_17, Inf, adjust.method="BH"))$table %>% mutate(effect = "H1")
qlf.genes_1_17$H2 <- (topTags(qlf.genes.filt.H1andH2_H2_1_17, Inf, adjust.method="BH"))$table %>% mutate(effect = "H2")

length(qlf.genes_1_17$H1$FDR[qlf.genes_1_17$H1$FDR< 0.05])#234
length(qlf.genes_1_17$H2$FDR[qlf.genes_1_17$H2$FDR< 0.05])#15
length(qlf.genes_1_17$interaction$FDR[qlf.genes_1_17$interaction$FDR< 0.05])#0


#Switch 2 
#............

#2hrs
qlf.genes.filt.H1xH2_interaction_2_2 <- glmQLFTest(fit.genes.filt.H1xH2_2_2, coef = 7:10 )
qlf.genes.filt.H1andH2_H1_2_2 <- glmQLFTest(fit.genes.filt.H1andH2_2_2, coef = 2:4)
qlf.genes.filt.H1andH2_H2_2_2 <- glmQLFTest(fit.genes.filt.H1andH2_2_2, coef = 5:6)

qlf.genes_2_2 <- list()

qlf.genes_2_2$interaction <- (topTags(qlf.genes.filt.H1xH2_interaction_2_2, Inf))$table %>% mutate(effect = "Int")
qlf.genes_2_2$H1 <- (topTags(qlf.genes.filt.H1andH2_H1_2_2, Inf))$table %>% mutate(effect = "H1")
qlf.genes_2_2$H2 <- (topTags(qlf.genes.filt.H1andH2_H2_2_2, Inf))$table %>% mutate(effect = "H2")

length(qlf.genes_2_2$H1$FDR[qlf.genes_2_2$H1$FDR< 0.05])
length(qlf.genes_2_2$H2$FDR[qlf.genes_2_2$H2$FDR< 0.05])
length(qlf.genes_2_2$interaction$FDR[qlf.genes_2_2$interaction$FDR< 0.05])

#17hrs
qlf.genes.filt.H1xH2_interaction_2_17 <- glmQLFTest(fit.genes.filt.H1xH2_2_17, coef = 7:10)
qlf.genes.filt.H1andH2_H1_2_17 <- glmQLFTest(fit.genes.filt.H1andH2_2_17, coef = 3:4)
qlf.genes.filt.H1andH2_H2_2_17 <- glmQLFTest(fit.genes.filt.H1andH2_2_17, coef = 5:6)

qlf.genes_2_17 <- list()

qlf.genes_2_17$interaction <- (topTags(qlf.genes.filt.H1xH2_interaction_2_17, Inf))$table %>% mutate(effect = "Int")
qlf.genes_2_17$H1 <- (topTags(qlf.genes.filt.H1andH2_H1_2_17, Inf))$table %>% mutate(effect = "H1")
qlf.genes_2_17$H2 <- (topTags(qlf.genes.filt.H1andH2_H2_2_17, Inf))$table %>% mutate(effect = "H2")

length(qlf.genes_2_17$H1$FDR[qlf.genes_2_17$H1$FDR< 0.05])#365
length(qlf.genes_2_17$H2$FDR[qlf.genes_2_17$H2$FDR< 0.05])#3783
length(qlf.genes_2_17$interaction$FDR[qlf.genes_2_17$interaction$FDR< 0.05])#68

```

## Visualize effect of H1, H2, H1xH2
```{r Venn_diagram H1, H2, H1xH2}
pval_threshold=0.05

#Switch 1

interaction.degs_1_2=row.names(qlf.genes_1_2$interaction[which(qlf.genes_1_2$interaction$FDR<=pval_threshold),])
H1.degs_1_2=row.names(qlf.genes_1_2$H1[which(qlf.genes_1_2$H1$FDR<=pval_threshold),])
H2.degs_1_2=row.names(qlf.genes_1_2$H2[which(qlf.genes_1_2$H2$FDR<=pval_threshold),])

interaction.degs_1_17=row.names(qlf.genes_1_17$interaction[which(qlf.genes_1_17$interaction$FDR<=pval_threshold),])
H1.degs_1_17=row.names(qlf.genes_1_17$H1[which(qlf.genes_1_17$H1$FDR<=pval_threshold),])
H2.degs_1_17=row.names(qlf.genes_1_17$H2[which(qlf.genes_1_17$H2$FDR<=pval_threshold),])

S1_2=list(H1.degs_1_2, H2.degs_1_2, interaction.degs_1_2 )
S1_17=list(H1.degs_1_17, H2.degs_1_17, interaction.degs_1_17)

venn_1_2=ggVennDiagram(S1_2,
                       label_alpha=0,
                       category.names = c("H1", "H2", "H1xH2"), set_size=8,
                       label_size = 6)+
  #geom_sf_text(size=10)+
  scale_color_manual(values = c("black","black","black"))+
  ggplot2::scale_fill_gradient(low = "white", high = "grey")+
  ggplot2::theme(legend.position="none")+
  ggplot2::ggtitle("2 hrs")+
  ggplot2::theme(plot.title=element_text(size=20, face="bold"))



venn_1_17=ggVennDiagram(S1_17,
                        label_alpha=0,
                        category.names = c("H1", "H2", "H1xH2"), set_size=8,
                        label_size = 6)+
  #geom_sf_text(size=10)+
  scale_color_manual(values = c("black","black","black"))+
  ggplot2::scale_fill_gradient(low = "white", high = "grey")+
  ggplot2::theme(legend.position="none")+
  ggplot2::ggtitle(" hrs")+
  ggplot2::theme(plot.title=element_text(size=20, face="bold"))


#Switch 2

interaction.degs_2_2=row.names(qlf.genes_2_2$interaction[which(qlf.genes_2_2$interaction$FDR<=pval_threshold),])
H1.degs_2_2=row.names(qlf.genes_2_2$H1[which(qlf.genes_2_2$H1$FDR<=pval_threshold),])
H2.degs_2_2=row.names(qlf.genes_2_2$H2[which(qlf.genes_2_2$H2$FDR<=pval_threshold),])

interaction.degs_2_17=row.names(qlf.genes_2_17$interaction[which(qlf.genes_2_17$interaction$FDR<=pval_threshold),])
H1.degs_2_17=row.names(qlf.genes_2_17$H1[which(qlf.genes_2_17$H1$FDR<=pval_threshold),])
H2.degs_2_17=row.names(qlf.genes_2_17$H2[which(qlf.genes_2_17$H2$FDR<=pval_threshold),])

S2_2=list(H1.degs_2_2, H2.degs_2_2, interaction.degs_2_2)
S2_17=list(H1.degs_2_17, H2.degs_2_17, interaction.degs_2_17)

venn_2_2=ggVennDiagram(S2_2,
                       label_alpha=0,
                       category.names = c("H1", "H2", "H1xH2"), set_size=8,
                       label_size = 6)+
  #geom_sf_text(size=10)+
  scale_color_manual(values = c("black","black","black"))+
  ggplot2::scale_fill_gradient(low = "white", high = "grey")+
  ggplot2::theme(legend.position="none")+
  ggplot2::ggtitle("2 hrs")+
  ggplot2::theme(plot.title=element_text(size=20, face="bold"))


venn_2_17=ggVennDiagram(S2_17,
                        label_alpha=0,
                        category.names = c("H1", "H2", "H1xH2"),set_size=8,
                        label_size = 6)+
  #geom_sf_text(size=10)+
  scale_color_manual(values = c("black","black","black"))+
  ggplot2::scale_fill_gradient(low = "white", high = "grey")+
  ggplot2::theme(legend.position="none")+
  ggplot2::ggtitle("17 hrs")+
  ggplot2::theme(plot.title=element_text(size=20, face="bold"))

ggarrange(venn_1_2, venn_1_17, venn_2_2, venn_2_17)

# coloration was added with Affinity Designer


#

```


# 7. Effect of treatment and host switches on the number of differentieally expressed genes
## design model matrix

```{r model matrix}
# https://support.bioconductor.org/p/56568 + edgeR User guide

#Switch 1
#............
#2hrs
Family_1_2 <-list_1_2$samples$BioRep
Group_1_2<-list_1_2_norm$samples$group
group_des_1_2<-model.matrix(~0+Family_1_2+Group_1_2)

#17hrs
Family_1_17 <-list_1_17$samples$BioRep
Group_1_17<-list_1_17_norm$samples$group
group_des_1_17<-model.matrix(~0+Family_1_17+Group_1_17)


#Switch 2
#............
#2hrs
Family_2_2 <-list_2_2$samples$Family
Group_2_2<-list_2_2_norm$samples$group
group_des_2_2<-model.matrix(~0+Family_2_2+Group_2_2)

#17hrs
Family_2_17 <-list_2_17$samples$Family
Group_2_17<-list_2_17_norm$samples$group
group_des_2_17<-model.matrix(~0+Family_2_17+Group_2_17)

```

```{r Estimate Dispersion}
#Switch 1
#..........
#2hrs
y.genes.filt.group_1_2<-estimateDisp(list_1_2_norm, group_des_1_2, robust=TRUE)
#17hrs
y.genes.filt.group_1_17 <- estimateDisp(list_1_17_norm,group_des_1_17, robust=TRUE)

#Switch 2
#..........
#2hrs
y.genes.filt.group_2_2 <- estimateDisp(list_2_2_norm, group_des_2_2, robust=TRUE)
#17hrs
y.genes.filt.group_2_17 <- estimateDisp(list_2_17_norm, group_des_2_17, robust=TRUE)
```

```{r Quasilikelihood fit object}

#Switch 1
#...........
#2hrs
fit.genes.filt.group_1_2=glmQLFit(y.genes.filt.group_1_2, group_des_1_2, robust=TRUE)
#17hrs
fit.genes.filt.group_1_17=glmQLFit(y.genes.filt.group_1_17, group_des_1_17, robust=TRUE)


#Switch 2
#...........
#2hrs
fit.genes.filt.group_2_2=glmQLFit(y.genes.filt.group_2_2, group_des_2_2, robust=TRUE)

#17hrs
fit.genes.filt.group_2_17=glmQLFit(y.genes.filt.group_2_17, group_des_2_17, robust=TRUE)

```

```{r make contrasts}
# to compare the different host switch treatments, the contrasts have to be given as a list

#Switch 1
#.............
#2hrs

colnames(group_des_1_2)
# "Family_1_2"     "Group_1_2E.E.2" "Group_1_2E.N.2" "Group_1_2E.S.2" "Group_1_2N.E.2" "Group_1_2N.N.2" "Group_1_2N.S.2" "Group_1_2S.E.2" "Group_1_2S.N.2" "Group_1_2S.S.2"

contrasts.S1_2 <-  makeContrasts(
  Group_1_2N.N.2-Group_1_2S.S.2,
  Group_1_2E.E.2-Group_1_2S.S.2,
  Group_1_2N.N.2-Group_1_2E.E.2,

  Group_1_2E.E.2-Group_1_2N.N.2,
  Group_1_2S.S.2-Group_1_2E.E.2,
  Group_1_2S.S.2-Group_1_2N.N.2,

  Group_1_2N.N.2-Group_1_2S.N.2,
  Group_1_2N.N.2-Group_1_2N.S.2,
  Group_1_2N.N.2-Group_1_2N.E.2,
  Group_1_2N.N.2-Group_1_2E.N.2,

  Group_1_2S.S.2-Group_1_2N.S.2,
  Group_1_2S.S.2-Group_1_2S.N.2,
  Group_1_2S.S.2-Group_1_2E.S.2,
  Group_1_2S.S.2-Group_1_2S.E.2,

  Group_1_2E.E.2-Group_1_2E.S.2,
  Group_1_2E.E.2-Group_1_2S.E.2,
  Group_1_2E.E.2-Group_1_2N.E.2,
  Group_1_2E.E.2-Group_1_2E.N.2,

  levels=group_des_1_2)

colnames(contrasts.S1_2) <- c("NNvSS", "EEvSS", "NNvEE", "EEvNN", "SSvEE", "SSvNN",
                            "NNvSN", "NNvNS", "NNvNE","NNvEN",
                             "SSvNS", "SSvSN", "SSvES", "SSvSE",
                            "EEvES","EEvSE", "EEvNE", "EEvEN")

#17hrs

colnames(group_des_1_17)
# "Family_1_17"      "Group_1_17E.E.17" "Group_1_17E.N.17" "Group_1_17E.S.17" "Group_1_17N.E.17" "Group_1_17N.N.17" "Group_1_17N.S.17" "Group_1_17S.E.17" "Group_1_17S.N.17" "Group_1_17S.S.17"

contrasts.S1_17 <-  makeContrasts(
  Group_1_17N.N.17-Group_1_17S.S.17,
  Group_1_17E.E.17-Group_1_17S.S.17,
  Group_1_17N.N.17-Group_1_17E.E.17,

  Group_1_17E.E.17-Group_1_17N.N.17,
  Group_1_17S.S.17-Group_1_17E.E.17,
  Group_1_17S.S.17-Group_1_17N.N.17,

  Group_1_17N.N.17-Group_1_17S.N.17,
  Group_1_17N.N.17-Group_1_17N.S.17,
  Group_1_17N.N.17-Group_1_17N.E.17,
  Group_1_17N.N.17-Group_1_17E.N.17,

  Group_1_17S.S.17-Group_1_17N.S.17,
  Group_1_17S.S.17-Group_1_17S.N.17,
  Group_1_17S.S.17-Group_1_17E.S.17,
  Group_1_17S.S.17-Group_1_17S.E.17,

  Group_1_17E.E.17-Group_1_17E.S.17,
  Group_1_17E.E.17-Group_1_17S.E.17,
  Group_1_17E.E.17-Group_1_17N.E.17,
  Group_1_17E.E.17-Group_1_17E.N.17,

  levels=group_des_1_17)

colnames(contrasts.S1_17) <- c("NNvSS", "EEvSS", "NNvEE", "EEvNN", "SSvEE", "SSvNN",
                              "NNvSN", "NNvNS", "NNvNE","NNvEN",
                              "SSvNS", "SSvSN", "SSvES", "SSvSE",
                              "EEvES","EEvSE", "EEvNE", "EEvEN")


#Switch 2
#.............
#2hrs

colnames(group_des_2_2)
#"Family_2_2"     "Group_2_2G.G.2" "Group_2_2G.N.2" "Group_2_2G.S.2" "Group_2_2N.G.2" "Group_2_2N.N.2" "Group_2_2N.S.2" "Group_2_2S.G.2" "Group_2_2S.N.2" "Group_2_2S.S.2"

contrasts.S2_2 <-  makeContrasts(
  Group_2_2N.N.2-Group_2_2S.S.2,
  Group_2_2G.G.2-Group_2_2S.S.2,
  Group_2_2N.N.2-Group_2_2G.G.2,

  Group_2_2G.G.2-Group_2_2N.N.2,
  Group_2_2S.S.2-Group_2_2G.G.2,
  Group_2_2S.S.2-Group_2_2N.N.2,

  Group_2_2N.N.2-Group_2_2S.N.2,
  Group_2_2N.N.2-Group_2_2N.S.2,
  Group_2_2N.N.2-Group_2_2N.G.2,
  Group_2_2N.N.2-Group_2_2G.N.2,

  Group_2_2S.S.2-Group_2_2N.S.2,
  Group_2_2S.S.2-Group_2_2S.N.2,
  Group_2_2S.S.2-Group_2_2G.S.2,
  Group_2_2S.S.2-Group_2_2S.G.2,

  Group_2_2G.G.2-Group_2_2G.S.2,
  Group_2_2G.G.2-Group_2_2S.G.2,
  Group_2_2G.G.2-Group_2_2N.G.2,
  Group_2_2G.G.2-Group_2_2G.N.2,


  levels=group_des_2_2)

colnames(contrasts.S2_2) <- c("NNvSS", "GGvSS", "NNvGG", "GGvNN", "SSvGG", "SSvNN",
                              "NNvSN", "NNvNS", "NNvNG","NNvGN",
                              "SSvNS", "SSvSN", "SSvGS", "SSvSG",
                              "GGvGS","GGvSG", "GGvNG", "GGvGN")


#17hrs

colnames(group_des_2_17)
#"Family_2_17"      "Group_2_17G.G.17" "Group_2_17G.N.17" "Group_2_17G.S.17" "Group_2_17N.G.17" "Group_2_17N.N.17" "Group_2_17N.S.17" "Group_2_17S.G.17""Group_2_17S.N.17" "Group_2_17S.S.17"

contrasts.S2_17 <-  makeContrasts(
  Group_2_17N.N.17-Group_2_17S.S.17,
  Group_2_17G.G.17-Group_2_17S.S.17,
  Group_2_17N.N.17-Group_2_17G.G.17,

  Group_2_17G.G.17-Group_2_17N.N.17,
  Group_2_17S.S.17-Group_2_17G.G.17,
  Group_2_17S.S.17-Group_2_17N.N.17,

  Group_2_17N.N.17-Group_2_17S.N.17,
  Group_2_17N.N.17-Group_2_17N.S.17,
  Group_2_17N.N.17-Group_2_17N.G.17,
  Group_2_17N.N.17-Group_2_17G.N.17,

  Group_2_17S.S.17-Group_2_17N.S.17,
  Group_2_17S.S.17-Group_2_17S.N.17,
  Group_2_17S.S.17-Group_2_17G.S.17,
  Group_2_17S.S.17-Group_2_17S.G.17,

  Group_2_17G.G.17-Group_2_17G.S.17,
  Group_2_17G.G.17-Group_2_17S.G.17,
  Group_2_17G.G.17-Group_2_17N.G.17,
  Group_2_17G.G.17-Group_2_17G.N.17,

  levels=group_des_2_17)

colnames(contrasts.S2_17) <- c("NNvSS", "GGvSS", "NNvGG", "GGvNN", "SSvGG", "SSvNN",
                              "NNvSN", "NNvNS", "NNvNG","NNvGN",
                              "SSvNS", "SSvSN", "SSvGS", "SSvSG",
                              "GGvGS","GGvSG", "GGvNG", "GGvGN")
```


```{r glmQLFTest}

#SWITCH 1
#Switch 1- 2hrs####
DGE_1_2 <- list()
DGE_1_2_gene <- list()

for (i in 1:dim(contrasts.S1_2)[2]){
  DGE_1_2[[i]]<- glmQLFTest(fit.genes.filt.group_1_2, contrast = contrasts.S1_2[,i])
  names(DGE_1_2)[i] <- colnames(contrasts.S1_2)[i]
  ngenes <- length(DGE_1_2[[i]]$genes$geneID)
  DGE_1_2_gene[[i]] <- topTags(DGE_1_2[[i]], n = Inf, adjust.method = "BH")
  names(DGE_1_2_gene)[i] <- colnames(contrasts.S1_2)[i]
  DGE_1_2_gene[[i]]$table$contrast <- colnames(contrasts.S1_2)[i]
  DGE_1_2_gene[[i]]$adjPvalBH <- p.adjust(p = DGE_1_2_gene[[i]]$P.Value, n = (ngenes*dim(contrasts.S1_2)[2]), method = "BH")
}

for(i in 1:length(DGE_1_2)){
  print(summary(decideTestsDGE(DGE_1_2[[i]])))
}

#Switch 1- 17hrs
DGE_1_17 <- list()
DGE_1_17_gene <- list()

for (i in 1:dim(contrasts.S1_17)[2]){
  DGE_1_17[[i]]<- glmQLFTest(fit.genes.filt.group_1_17, contrast = contrasts.S1_17[,i])
  names(DGE_1_17)[i] <- colnames(contrasts.S1_17)[i]
  ngenes <- length(DGE_1_17[[i]]$genes$geneID)
  DGE_1_17_gene[[i]] <- topTags(DGE_1_17[[i]], n = Inf, adjust.method = "BH")
  names(DGE_1_17_gene)[i] <- colnames(contrasts.S1_17)[i]
  DGE_1_17_gene[[i]]$table$contrast <- colnames(contrasts.S1_17)[i]
}

for(i in 1:length(DGE_1_17)){
   print(summary(decideTestsDGE(DGE_1_17[[i]])))
}


#Switch 2- 2hrs
 DGE_2_2 <- list()
 DGE_2_2_gene <- list()

 for (i in 1:dim(contrasts.S2_2)[2]){
   DGE_2_2[[i]]<- glmQLFTest(fit.genes.filt.group_2_2, contrast = contrasts.S2_2[,i])
   names(DGE_2_2)[i] <- colnames(contrasts.S2_2)[i]
   ngenes <- length(DGE_2_2[[i]]$genes$geneID)
   DGE_2_2_gene[[i]] <- topTags(DGE_2_2[[i]], n = Inf, adjust.method = "BH")
   names(DGE_2_2_gene)[i] <- colnames(contrasts.S2_2)[i]
   DGE_2_2_gene[[i]]$table$contrast <- colnames(contrasts.S2_2)[i]
}


 for(i in 1:length(DGE_2_2)){
   print(summary(decideTestsDGE(DGE_2_2[[i]])))
  }

#Switch 2- 17hrs
 DGE_2_17 <- list()
  DGE_2_17_gene <- list()

  for (i in 1:dim(contrasts.S2_17)[2]){
    DGE_2_17[[i]]<- glmQLFTest(fit.genes.filt.group_2_17, contrast = contrasts.S2_17[,i])
    names(DGE_2_17)[i] <- colnames(contrasts.S2_17)[i]
    ngenes <- length(DGE_2_17[[i]]$genes$geneID)
    DGE_2_17_gene[[i]] <- topTags(DGE_2_17[[i]], n = Inf, adjust.method = "BH")
    names(DGE_2_17_gene)[i] <- colnames(contrasts.S2_17)[i]
    DGE_2_17_gene[[i]]$table$contrast <- colnames(contrasts.S2_17)[i]
  }


  for(i in 1:length(DGE_2_17)){
    print(summary(decideTestsDGE(DGE_2_17[[i]])))
}
```

## Visualize differentially expressed genes
```{r UpSetR Plot for significatly differentially expressed genes (FDR <0.05)}

#note: N= Urtica dioica, S=Salix caprea, E=Ulmus glabra, G=Ribes uva crisp

#Switch 1
#join all data


df12_1=dplyr::select(DGE_1_2_gene$NNvSS$table, genes, logFC_NNSS=logFC, FDR_NNSS=FDR)
df12_2=dplyr::select(DGE_1_2_gene$EEvSS$table, genes, logFC_GGSS=logFC, FDR_EESS=FDR)
df12_3=dplyr::select(DGE_1_2_gene$NNvEE$table, genes, logFC_NNGG=logFC, FDR_NNEE=FDR)
df12_4=dplyr::select(DGE_1_2_gene$NNvSN$table, genes, logFC_NNSN=logFC, FDR_NNSN=FDR)
df12_5=dplyr::select(DGE_1_2_gene$NNvNS$table, genes, logFC_NNNS=logFC, FDR_NNNS=FDR)
df12_6=dplyr::select(DGE_1_2_gene$NNvNE$table, genes, logFC_NNNG=logFC, FDR_NNNE=FDR)
df12_7=dplyr::select(DGE_1_2_gene$NNvEN$table, genes, logFC_NNGN=logFC, FDR_NNEN=FDR)
df12_8=dplyr::select(DGE_1_2_gene$SSvNS$table, genes, logFC_SSNS=logFC, FDR_SSNS=FDR)
df12_9=dplyr::select(DGE_1_2_gene$SSvSN$table, genes, logFC_SSSN=logFC, FDR_SSSN=FDR)
df12_10=dplyr::select(DGE_1_2_gene$SSvES$table, genes, logFC_SSGS=logFC, FDR_SSES=FDR)
df12_11=dplyr::select(DGE_1_2_gene$SSvSE$table, genes, logFC_SSSG=logFC, FDR_SSSE=FDR)
df12_12=dplyr::select(DGE_1_2_gene$EEvES$table, genes, logFC_GGGS=logFC, FDR_EEES=FDR)
df12_13=dplyr::select(DGE_1_2_gene$EEvSE$table, genes, logFC_GGSG=logFC, FDR_EESE=FDR)
df12_14=dplyr::select(DGE_1_2_gene$EEvNE$table, genes, logFC_GGNG=logFC, FDR_EENE=FDR)
df12_15=dplyr::select(DGE_1_2_gene$EEvEN$table, genes, logFC_GGGN=logFC, FDR_EEEN=FDR)
df12_16=dplyr::select(DGE_1_2_gene$EEvNN$table, genes, logFC_GGNN=logFC, FDR_EENN=FDR)
df12_17=dplyr::select(DGE_1_2_gene$SSvEE$table, genes, logFC_SSGG=logFC, FDR_SSEE=FDR)
df12_18=dplyr::select(DGE_1_2_gene$SSvNN$table, genes, logFC_SSNN=logFC, FDR_SSNN=FDR)




df12_list <- list(df12_1, df12_2, df12_3, df12_4, df12_5, df12_6, df12_7,
                  df12_8, df12_9, df12_10, df12_11, df12_12, df12_13, df12_14, df12_15, df12_16, df12_17, df12_18)

#merge all data frames together
merge_12=Reduce(function(x, y) merge(x, y, all=TRUE), df12_list)

DEG_FDR_12=list(
  UdUdvScSc=merge_12[merge_12$FDR_NNSS<0.05,],
  UgUgvScSc=merge_12[merge_12$FDR_EESS<0.05,],
  UdUdvUgUg=merge_12[merge_12$FDR_NNEE<0.05,],
  UdUdvScUd=merge_12[merge_12$FDR_NNSN<0.05,],
  UdUdvUdSc=merge_12[merge_12$FDR_NNNS<0.05,],
  UdUdvUdUg=merge_12[merge_12$FDR_NNNE<0.05,],
  UdUdvUgUd=merge_12[merge_12$FDR_NNEN<0.05,],
  ScScvUdSc=merge_12[merge_12$FDR_SSNS<0.05,],
  ScScvScUd=merge_12[merge_12$FDR_SSSN<0.05,],
  ScScvUgSc=merge_12[merge_12$FDR_SSES<0.05,],
  ScScvScUg=merge_12[merge_12$FDR_SSSE<0.05,],
  UgUgvUgSc=merge_12[merge_12$FDR_EEES<0.05,],
  UgUgvScUg=merge_12[merge_12$FDR_EESE<0.05,],
  UgUgvUdUg=merge_12[merge_12$FDR_EENE<0.05,],
  UgUgvUgUd=merge_12[merge_12$FDR_EEEN<0.05,],
  UgUgvUdUd=merge_12[merge_12$FDR_EENN<0.05,],
  ScScvUgUg=merge_12[merge_12$FDR_SSEE<0.05,],
  ScScvUdUd=merge_12[merge_12$FDR_SSNN<0.05,]
)



DEG_sig_12=list(
  UdUd.v.ScSc=DEG_FDR_12$UdUdvScSc$genes,
  UgUg.v.ScSc=DEG_FDR_12$UgUgvScSc$genes,
  UdUd.v.UgUg=DEG_FDR_12$UdUdvUgUg$genes,
  UdUd.v.ScUd=DEG_FDR_12$UdUdvScUd$genes,
  UdUd.v.UdSc=DEG_FDR_12$UdUdvUdSc$genes,
  UdUd.v.UdUg=DEG_FDR_12$UdUdvUdUg$genes,
  UdUd.v.UgUd=DEG_FDR_12$UdUdvUgUd$genes,
  ScSc.v.UdSc=DEG_FDR_12$ScScvUdSc$genes,
  ScSc.v.ScUd=DEG_FDR_12$ScScvScUd$genes,
  ScSc.v.UgSc=DEG_FDR_12$ScScvUgSc$genes,
  ScSc.v.ScUg=DEG_FDR_12$ScScvScUg$genes,
  UgUg.v.UgSc=DEG_FDR_12$UgUgvUgSc$genes,
  UgUg.v.ScUg=DEG_FDR_12$UgUgvScUg$genes,
  UgUg.v.UdUg=DEG_FDR_12$UgUgvUdUg$genes,
  UgUg.v.UgUd=DEG_FDR_12$UgUgvUgUd$genes,
  UgUg.v.UdUd=DEG_FDR_12$UgUgvUdUd$genes,
  ScSc.v.UgUg=DEG_FDR_12$ScScvUgUg$genes,
  ScSc.v.UdUd=DEG_FDR_12$ScScvUdUd$genes
)


#make upset plot
#--------------------------
N_1_12=upset(fromList(DEG_sig_12), sets = c( "UgUg.v.UgUd", "UgUg.v.UdUd", "ScSc.v.ScUd","ScSc.v.UdUd"),keep.order=TRUE, mb.ratio = c(0.50, 0.50),
             order.by = "freq", text.scale=1.5)

N_cow_1_12=cowplot::plot_grid(NULL, N_1_12$Main_bar, N_1_12$Sizes, N_1_12$Matrix,
                              nrow=2, align="hv", rel_widths = c(0.4, 1), rel_heights = c(0.5,0.5))

#Salix
S_1_12=upset(fromList(DEG_sig_12), sets = c("UgUg.v.UgSc", "UgUg.v.ScSc","UdUd.v.UdSc","UdUd.v.ScSc"),keep.order=TRUE, mb.ratio = c(0.50, 0.50),
             order.by = "freq" ,text.scale=1.5)

S_cow_1_12=cowplot::plot_grid(NULL, S_1_12$Main_bar, S_1_12$Sizes, S_1_12$Matrix,
                              nrow=2, align="hv", rel_widths = c(0.4, 1), rel_heights = c(0.5,0.5))

#Ulmus
E_1_12=upset(fromList(DEG_sig_12), sets = c("ScSc.v.ScUg", "ScSc.v.UgUg","UdUd.v.UdUg","UdUd.v.UgUg"),keep.order=TRUE, mb.ratio = c(0.50, 0.50),
             order.by = "freq",text.scale=1.5)

E_cow_1_12=cowplot::plot_grid(NULL, E_1_12$Main_bar, E_1_12$Sizes,E_1_12$Matrix,
                              nrow=2, align="hv", rel_widths = c(0.4, 1), rel_heights = c(0.5,0.5))

#Switch 1_17

df117_1=dplyr::select(DGE_1_17_gene$NNvSS$table, genes, logFC_NNSS=logFC, FDR_NNSS=FDR)
df117_2=dplyr::select(DGE_1_17_gene$EEvSS$table, genes, logFC_GGSS=logFC, FDR_EESS=FDR)
df117_3=dplyr::select(DGE_1_17_gene$NNvEE$table, genes, logFC_NNGG=logFC, FDR_NNEE=FDR)
df117_4=dplyr::select(DGE_1_17_gene$NNvSN$table, genes, logFC_NNSN=logFC, FDR_NNSN=FDR)
df117_5=dplyr::select(DGE_1_17_gene$NNvNS$table, genes, logFC_NNNS=logFC, FDR_NNNS=FDR)
df117_6=dplyr::select(DGE_1_17_gene$NNvNE$table, genes, logFC_NNNG=logFC, FDR_NNNE=FDR)
df117_7=dplyr::select(DGE_1_17_gene$NNvEN$table, genes, logFC_NNGN=logFC, FDR_NNEN=FDR)
df117_8=dplyr::select(DGE_1_17_gene$SSvNS$table, genes, logFC_SSNS=logFC, FDR_SSNS=FDR)
df117_9=dplyr::select(DGE_1_17_gene$SSvSN$table, genes, logFC_SSSN=logFC, FDR_SSSN=FDR)
df117_10=dplyr::select(DGE_1_17_gene$SSvES$table, genes, logFC_SSGS=logFC, FDR_SSES=FDR)
df117_11=dplyr::select(DGE_1_17_gene$SSvSE$table, genes, logFC_SSSG=logFC, FDR_SSSE=FDR)
df117_117=dplyr::select(DGE_1_17_gene$EEvES$table, genes, logFC_GGGS=logFC, FDR_EEES=FDR)
df117_13=dplyr::select(DGE_1_17_gene$EEvSE$table, genes, logFC_GGSG=logFC, FDR_EESE=FDR)
df117_14=dplyr::select(DGE_1_17_gene$EEvNE$table, genes, logFC_GGNG=logFC, FDR_EENE=FDR)
df117_15=dplyr::select(DGE_1_17_gene$EEvEN$table, genes, logFC_GGGN=logFC, FDR_EEEN=FDR)
df117_16=dplyr::select(DGE_1_17_gene$EEvNN$table, genes, logFC_GGNN=logFC, FDR_EENN=FDR)
df117_17=dplyr::select(DGE_1_17_gene$SSvEE$table, genes, logFC_SSGG=logFC, FDR_SSEE=FDR)
df117_18=dplyr::select(DGE_1_17_gene$SSvNN$table, genes, logFC_SSNN=logFC, FDR_SSNN=FDR)




df117_list <- list(df117_1, df117_2, df117_3, df117_4, df117_5, df117_6, df117_7,
                   df117_8, df117_9, df117_10, df117_11, df117_117, df117_13, df117_14, df117_15, df117_16, df117_17, df117_18)

#merge all data frames together
merge_117=Reduce(function(x, y) merge(x, y, all=TRUE), df117_list)

DEG_FDR_117=list(
  UdUdvScSc=merge_117[merge_117$FDR_NNSS<0.05,],
  UgUgvScSc=merge_117[merge_117$FDR_EESS<0.05,],
  UdUdvUgUg=merge_117[merge_117$FDR_NNEE<0.05,],
  UdUdvScUd=merge_117[merge_117$FDR_NNSN<0.05,],
  UdUdvUdSc=merge_117[merge_117$FDR_NNNS<0.05,],
  UdUdvUdUg=merge_117[merge_117$FDR_NNNE<0.05,],
  UdUdvUgUd=merge_117[merge_117$FDR_NNEN<0.05,],
  ScScvUdSc=merge_117[merge_117$FDR_SSNS<0.05,],
  ScScvScUd=merge_117[merge_117$FDR_SSSN<0.05,],
  ScScvUgSc=merge_117[merge_117$FDR_SSES<0.05,],
  ScScvScUg=merge_117[merge_117$FDR_SSSE<0.05,],
  UgUgvUgSc=merge_117[merge_117$FDR_EEES<0.05,],
  UgUgvScUg=merge_117[merge_117$FDR_EESE<0.05,],
  UgUgvUdUg=merge_117[merge_117$FDR_EENE<0.05,],
  UgUgvUgUd=merge_117[merge_117$FDR_EEEN<0.05,],
  UgUgvUdUd=merge_117[merge_117$FDR_EENN<0.05,],
  ScScvUgUg=merge_117[merge_117$FDR_SSEE<0.05,],
  ScScvUdUd=merge_117[merge_117$FDR_SSNN<0.05,]
)



DEG_sig_117=list(
  UdUd.v.ScSc=DEG_FDR_117$UdUdvScSc$genes,
  UgUg.v.ScSc=DEG_FDR_117$UgUgvScSc$genes,
  UdUd.v.UgUg=DEG_FDR_117$UdUdvUgUg$genes,
  UdUd.v.ScUd=DEG_FDR_117$UdUdvScUd$genes,
  UdUd.v.UdSc=DEG_FDR_117$UdUdvUdSc$genes,
  UdUd.v.UdUg=DEG_FDR_117$UdUdvUdUg$genes,
  UdUd.v.UgUd=DEG_FDR_117$UdUdvUgUd$genes,
  ScSc.v.UdSc=DEG_FDR_117$ScScvUdSc$genes,
  ScSc.v.ScUd=DEG_FDR_117$ScScvScUd$genes,
  ScSc.v.UgSc=DEG_FDR_117$ScScvUgSc$genes,
  ScSc.v.ScUg=DEG_FDR_117$ScScvScUg$genes,
  UgUg.v.UgSc=DEG_FDR_117$UgUgvUgSc$genes,
  UgUg.v.ScUg=DEG_FDR_117$UgUgvScUg$genes,
  UgUg.v.UdUg=DEG_FDR_117$UgUgvUdUg$genes,
  UgUg.v.UgUd=DEG_FDR_117$UgUgvUgUd$genes,
  UgUg.v.UdUd=DEG_FDR_117$UgUgvUdUd$genes,
  ScSc.v.UgUg=DEG_FDR_117$ScScvUgUg$genes,
  ScSc.v.UdUd=DEG_FDR_117$ScScvUdUd$genes
)

#make upset plot
N_1_17=upset(fromList(DEG_sig_117), sets = c( "UgUg.v.UgUd", "UgUg.v.UdUd", "ScSc.v.ScUd","ScSc.v.UdUd"),keep.order=TRUE, mb.ratio = c(0.50, 0.50),
             order.by = "freq", text.scale=1.5)

N_cow_1_17=cowplot::plot_grid(NULL, N_1_17$Main_bar, N_1_17$Sizes, N_1_17$Matrix,
                              nrow=2, align="hv", rel_widths = c(0.4, 1), rel_heights = c(0.5,0.5))

#Salix
S_1_17=upset(fromList(DEG_sig_117), sets = c("UgUg.v.UgSc", "UgUg.v.ScSc","UdUd.v.UdSc","UdUd.v.ScSc"),keep.order=TRUE, mb.ratio = c(0.50, 0.50),
             order.by = "freq" ,text.scale=1.5)

S_cow_1_17=cowplot::plot_grid(NULL, S_1_17$Main_bar, S_1_17$Sizes, S_1_17$Matrix,
                              nrow=2, align="hv", rel_widths = c(0.4, 1), rel_heights = c(0.5,0.5))

#Ulmus
E_1_17=upset(fromList(DEG_sig_117), sets = c("ScSc.v.ScUg", "ScSc.v.UgUg","UdUd.v.UdUg","UdUd.v.UgUg"),keep.order=TRUE, mb.ratio = c(0.50, 0.50),
             order.by = "freq",text.scale=1.5)

E_cow_1_17=cowplot::plot_grid(NULL, E_1_17$Main_bar, E_1_17$Sizes, E_1_17$Matrix,
                              nrow=2, align="hv", rel_widths = c(0.4, 1), rel_heights = c(0.5,0.5))


upset_1_17=grid.arrange(N_cow_1_17, S_cow_1_17, E_cow_1_17, ncol=1)

#Switch 2_2

df22_1=dplyr::select(DGE_2_2_gene$NNvSS$table, genes, logFC_NNSS=logFC, FDR_NNSS=FDR)
df22_2=dplyr::select(DGE_2_2_gene$GGvSS$table, genes, logFC_GGSS=logFC, FDR_GGSS=FDR)
df22_3=dplyr::select(DGE_2_2_gene$NNvGG$table, genes, logFC_NNGG=logFC, FDR_NNGG=FDR)
df22_4=dplyr::select(DGE_2_2_gene$NNvSN$table, genes, logFC_NNSN=logFC, FDR_NNSN=FDR)
df22_5=dplyr::select(DGE_2_2_gene$NNvNS$table, genes, logFC_NNNS=logFC, FDR_NNNS=FDR)
df22_6=dplyr::select(DGE_2_2_gene$NNvNG$table, genes, logFC_NNNG=logFC, FDR_NNNG=FDR)
df22_7=dplyr::select(DGE_2_2_gene$NNvGN$table, genes, logFC_NNGN=logFC, FDR_NNGN=FDR)
df22_8=dplyr::select(DGE_2_2_gene$SSvNS$table, genes, logFC_SSNS=logFC, FDR_SSNS=FDR)
df22_9=dplyr::select(DGE_2_2_gene$SSvSN$table, genes, logFC_SSSN=logFC, FDR_SSSN=FDR)
df22_10=dplyr::select(DGE_2_2_gene$SSvGS$table, genes, logFC_SSGS=logFC, FDR_SSGS=FDR)
df22_11=dplyr::select(DGE_2_2_gene$SSvSG$table, genes, logFC_SSSG=logFC, FDR_SSSG=FDR)
df22_12=dplyr::select(DGE_2_2_gene$GGvGS$table, genes, logFC_GGGS=logFC, FDR_GGGS=FDR)
df22_13=dplyr::select(DGE_2_2_gene$GGvSG$table, genes, logFC_GGSG=logFC, FDR_GGSG=FDR)
df22_14=dplyr::select(DGE_2_2_gene$GGvNG$table, genes, logFC_GGNG=logFC, FDR_GGNG=FDR)
df22_15=dplyr::select(DGE_2_2_gene$GGvGN$table, genes, logFC_GGGN=logFC, FDR_GGGN=FDR)
df22_16=dplyr::select(DGE_2_2_gene$GGvNN$table, genes, logFC_GGNN=logFC, FDR_GGNN=FDR)
df22_17=dplyr::select(DGE_2_2_gene$SSvGG$table, genes, logFC_SSGG=logFC, FDR_SSGG=FDR)
df22_18=dplyr::select(DGE_2_2_gene$SSvNN$table, genes, logFC_SSNN=logFC, FDR_SSNN=FDR)




df22_list <- list(df22_1, df22_2, df22_3, df22_4, df22_5, df22_6, df22_7,
                  df22_8, df22_9, df22_10, df22_11, df22_12, df22_13, df22_14, df22_15, df22_16, df22_17, df22_18)

#merge all data frames together
merge_22=Reduce(function(x, y) merge(x, y, all=TRUE), df22_list)

DEG_FDR_22=list(
  UdUdvScSc=merge_22[merge_22$FDR_NNSS<0.05,],
  RuRuvScSc=merge_22[merge_22$FDR_GGSS<0.05,],
  UdUdvRuRu=merge_22[merge_22$FDR_NNGG<0.05,],
  UdUdvScUd=merge_22[merge_22$FDR_NNSN<0.05,],
  UdUdvUdSc=merge_22[merge_22$FDR_NNNS<0.05,],
  UdUdvUdRu=merge_22[merge_22$FDR_NNNG<0.05,],
  UdUdvRuUd=merge_22[merge_22$FDR_NNGN<0.05,],
  ScScvUdSc=merge_22[merge_22$FDR_SSNS<0.05,],
  ScScvScUd=merge_22[merge_22$FDR_SSSN<0.05,],
  ScScvRuSc=merge_22[merge_22$FDR_SSGS<0.05,],
  ScScvScRu=merge_22[merge_22$FDR_SSSG<0.05,],
  RuRuvRuSc=merge_22[merge_22$FDR_GGGS<0.05,],
  RuRuvScRu=merge_22[merge_22$FDR_GGSG<0.05,],
  RuRuvUdRu=merge_22[merge_22$FDR_GGNG<0.05,],
  RuRuvRuUd=merge_22[merge_22$FDR_GGGN<0.05,],
  RuRuvUdUd=merge_22[merge_22$FDR_GGNN<0.05,],
  ScScvRuRu=merge_22[merge_22$FDR_SSGG<0.05,],
  ScScvUdUd=merge_22[merge_22$FDR_SSNN<0.05,]
)



DEG_sig_22=list(
  UdUd.v.ScSc=DEG_FDR_22$UdUdvScSc$genes,
  RuRu.v.ScSc=DEG_FDR_22$RuRuvScSc$genes,
  UdUd.v.RuRu=DEG_FDR_22$UdUdvRuRu$genes,
  UdUd.v.ScUd=DEG_FDR_22$UdUdvScUd$genes,
  UdUd.v.UdSc=DEG_FDR_22$UdUdvUdSc$genes,
  UdUd.v.UdRu=DEG_FDR_22$UdUdvUdRu$genes,
  UdUd.v.RuUd=DEG_FDR_22$UdUdvRuUd$genes,
  ScSc.v.UdSc=DEG_FDR_22$ScScvUdSc$genes,
  ScSc.v.ScUd=DEG_FDR_22$ScScvScUd$genes,
  ScSc.v.RuSc=DEG_FDR_22$ScScvRuSc$genes,
  ScSc.v.ScRu=DEG_FDR_22$ScScvScRu$genes,
  RuRu.v.RuSc=DEG_FDR_22$RuRuvRuSc$genes,
  RuRu.v.ScRu=DEG_FDR_22$RuRuvScRu$genes,
  RuRu.v.UdRu=DEG_FDR_22$RuRuvUdRu$genes,
  RuRu.v.RuUd=DEG_FDR_22$RuRuvRuUd$genes,
  RuRu.v.UdUd=DEG_FDR_22$RuRuvUdUd$genes,
  ScSc.v.RuRu=DEG_FDR_22$ScScvRuRu$genes,
  ScSc.v.UdUd=DEG_FDR_22$ScScvUdUd$genes
)

#make upset plots

#Urtica
N_2_2=upset(fromList(DEG_sig_22), sets = c( "RuRu.v.RuUd", "RuRu.v.UdUd", "ScSc.v.ScUd","ScSc.v.UdUd"),keep.order=TRUE, mb.ratio = c(0.50, 0.50),
             order.by = "freq", text.scale=1.5)

N_cow_2_2=cowplot::plot_grid(NULL, N_2_2$Main_bar, N_2_2$Sizes, N_2_2$Matrix,
                              nrow=2, align="hv", rel_widths = c(0.4, 1), rel_heights = c(0.5,0.5))

#Salix
S_2_2=upset(fromList(DEG_sig_22), sets = c("RuRu.v.RuSc", "RuRu.v.ScSc","UdUd.v.UdSc","UdUd.v.ScSc"),keep.order=TRUE, mb.ratio = c(0.50, 0.50),
             order.by = "freq" ,text.scale=1.5)

S_cow_2_2=cowplot::plot_grid(NULL, S_2_2$Main_bar, S_2_2$Sizes, S_2_2$Matrix,
                              nrow=2, align="hv", rel_widths = c(0.4, 1), rel_heights = c(0.5,0.5))

#Ribes
G_2_2=upset(fromList(DEG_sig_22), sets = c("ScSc.v.ScRu", "ScSc.v.RuRu","UdUd.v.UdRu","UdUd.v.RuRu"),keep.order=TRUE, mb.ratio = c(0.50, 0.50),
             order.by = "freq",text.scale=1.5)
G_cow_2_2=cowplot::plot_grid(NULL, G_2_2$Main_bar, G_2_2$Sizes, G_2_2$Matrix,
                              nrow=2, align="hv", rel_widths = c(0.4, 1), rel_heights = c(0.5,0.5))


upset_2_2=grid.arrange(N_cow_2_2, S_cow_2_2, G_cow_2_2, ncol=1)

#Switch 2_17
df217_1=dplyr::select(DGE_2_17_gene$NNvSS$table, genes, logFC_NNSS=logFC, FDR_NNSS=FDR)
df217_2=dplyr::select(DGE_2_17_gene$GGvSS$table, genes, logFC_GGSS=logFC, FDR_GGSS=FDR)
df217_3=dplyr::select(DGE_2_17_gene$NNvGG$table, genes, logFC_NNGG=logFC, FDR_NNGG=FDR)
df217_4=dplyr::select(DGE_2_17_gene$NNvSN$table, genes, logFC_NNSN=logFC, FDR_NNSN=FDR)
df217_5=dplyr::select(DGE_2_17_gene$NNvNS$table, genes, logFC_NNNS=logFC, FDR_NNNS=FDR)
df217_6=dplyr::select(DGE_2_17_gene$NNvNG$table, genes, logFC_NNNG=logFC, FDR_NNNG=FDR)
df217_7=dplyr::select(DGE_2_17_gene$NNvGN$table, genes, logFC_NNGN=logFC, FDR_NNGN=FDR)
df217_8=dplyr::select(DGE_2_17_gene$SSvNS$table, genes, logFC_SSNS=logFC, FDR_SSNS=FDR)
df217_9=dplyr::select(DGE_2_17_gene$SSvSN$table, genes, logFC_SSSN=logFC, FDR_SSSN=FDR)
df217_10=dplyr::select(DGE_2_17_gene$SSvGS$table, genes, logFC_SSGS=logFC, FDR_SSGS=FDR)
df217_11=dplyr::select(DGE_2_17_gene$SSvSG$table, genes, logFC_SSSG=logFC, FDR_SSSG=FDR)
df217_12=dplyr::select(DGE_2_17_gene$GGvGS$table, genes, logFC_GGGS=logFC, FDR_GGGS=FDR)
df217_13=dplyr::select(DGE_2_17_gene$GGvSG$table, genes, logFC_GGSG=logFC, FDR_GGSG=FDR)
df217_14=dplyr::select(DGE_2_17_gene$GGvNG$table, genes, logFC_GGNG=logFC, FDR_GGNG=FDR)
df217_15=dplyr::select(DGE_2_17_gene$GGvGN$table, genes, logFC_GGGN=logFC, FDR_GGGN=FDR)
df217_16=dplyr::select(DGE_2_17_gene$GGvNN$table, genes, logFC_GGNN=logFC, FDR_GGNN=FDR)
df217_17=dplyr::select(DGE_2_17_gene$SSvGG$table, genes, logFC_SSGG=logFC, FDR_SSGG=FDR)
df217_18=dplyr::select(DGE_2_17_gene$SSvNN$table, genes, logFC_SSNN=logFC, FDR_SSNN=FDR)

df217_list <- list(df217_1, df217_2, df217_3, df217_4, df217_5, df217_6, df217_7,
                  df217_8, df217_9, df217_10, df217_11, df217_12, df217_13, df217_14, df217_15, df217_16, df217_17, df217_18)

#merge all data frames together
merge_217=Reduce(function(x, y) merge(x, y, all=TRUE), df217_list)

DEG_FDR_217=list(
  UdUdvScSc=merge_217[merge_217$FDR_NNSS<0.05,],
  RuRuvScSc=merge_217[merge_217$FDR_GGSS<0.05,],
  UdUdvRuRu=merge_217[merge_217$FDR_NNGG<0.05,],
  UdUdvScUd=merge_217[merge_217$FDR_NNSN<0.05,],
  UdUdvUdSc=merge_217[merge_217$FDR_NNNS<0.05,],
  UdUdvUdRu=merge_217[merge_217$FDR_NNNG<0.05,],
  UdUdvRuUd=merge_217[merge_217$FDR_NNGN<0.05,],
  ScScvUdSc=merge_217[merge_217$FDR_SSNS<0.05,],
  ScScvScUd=merge_217[merge_217$FDR_SSSN<0.05,],
  ScScvRuSc=merge_217[merge_217$FDR_SSGS<0.05,],
  ScScvScRu=merge_217[merge_217$FDR_SSSG<0.05,],
  RuRuvRuSc=merge_217[merge_217$FDR_GGGS<0.05,],
  RuRuvScRu=merge_217[merge_217$FDR_GGSG<0.05,],
  RuRuvUdRu=merge_217[merge_217$FDR_GGNG<0.05,],
  RuRuvRuUd=merge_217[merge_217$FDR_GGGN<0.05,],
  RuRuvUdUd=merge_217[merge_217$FDR_GGNN<0.05,],
  ScScvRuRu=merge_217[merge_217$FDR_SSGG<0.05,],
  ScScvUdud=merge_217[merge_217$FDR_SSNN<0.05,]
)



DEG_sig_217=list(
  UdUd.v.ScSc=DEG_FDR_217$UdUdvScSc$genes,
  RuRu.v.ScSc=DEG_FDR_217$RuRuvScSc$genes,
  UdUd.v.RuRu=DEG_FDR_217$UdUdvRuRu$genes,
  UdUd.v.ScUd=DEG_FDR_217$UdUdvScUd$genes,
  UdUd.v.UdSc=DEG_FDR_217$UdUdvUdSc$genes,
  UdUd.v.UdRu=DEG_FDR_217$UdUdvUdRu$genes,
  UdUd.v.RuUd=DEG_FDR_217$UdUdvRuUd$genes,
  ScSc.v.UdSc=DEG_FDR_217$ScScvUdSc$genes,
  ScSc.v.ScUd=DEG_FDR_217$ScScvScUd$genes,
  ScSc.v.RuSc=DEG_FDR_217$ScScvRuSc$genes,
  ScSc.v.ScRu=DEG_FDR_217$ScScvScRu$genes,
  RuRu.v.RuSc=DEG_FDR_217$RuRuvRuSc$genes,
  RuRu.v.ScRu=DEG_FDR_217$RuRuvScRu$genes,
  RuRu.v.UdRu=DEG_FDR_217$RuRuvUdRu$genes,
  RuRu.v.RuUd=DEG_FDR_217$RuRuvRuUd$genes,
  RuRu.v.UdUd=DEG_FDR_217$RuRuvUdUd$genes,
  ScSc.v.RuRu=DEG_FDR_217$ScScvRuRu$genes,
  ScSc.v.UdUd=DEG_FDR_217$ScScvUdUd$genes
)


#make upset plot
#Urtica
N_2_17=upset(fromList(DEG_sig_217), sets = c( "RuRu.v.RuUd", "RuRu.v.UdUd", "ScSc.v.ScUd","ScSc.v.UdUd"),keep.order=TRUE, mb.ratio = c(0.50, 0.50),
            order.by = "freq", text.scale=1.5)

N_cow_2_17=cowplot::plot_grid(NULL, N_2_17$Main_bar, N_2_17$Sizes, N_2_17$Matrix,
             nrow=2, align="hv", rel_widths = c(0.4, 1), rel_heights = c(0.5,0.5))


#Salix
S_2_17=upset(fromList(DEG_sig_217), sets = c("RuRu.v.RuSc", "RuRu.v.ScSc","UdUd.v.UdSc","UdUd.v.ScSc"),keep.order=TRUE, mb.ratio = c(0.50, 0.50),
            order.by = "freq" ,text.scale=1.5)

S_cow_2_17=cowplot::plot_grid(NULL, S_2_17$Main_bar, S_2_17$Sizes, S_2_17$Matrix,
            nrow=2, align="hv", rel_widths = c(0.4, 1), rel_heights = c(0.5,0.5))

#Ribes
G_2_17=upset(fromList(DEG_sig_217), sets = c("ScSc.v.ScRu", "ScSc.v.RuRu","UdUd.v.UdRu","UdUd.v.RuRu"),keep.order=TRUE, mb.ratio = c(0.50, 0.50),
            order.by = "freq",text.scale=1.5)

G_cow_2_17=cowplot::plot_grid(NULL, G_2_17$Main_bar, G_2_17$Sizes, G_2_17$Matrix,
            nrow=2, align="hv", rel_widths = c(0.4, 1), rel_heights = c(0.5,0.5))

upset_2_17=grid.arrange(N_cow_2_17, S_cow_2_17, G_cow_2_17, ncol=1)
```


#8. Evaluation of differentially expressed genes

plot genes differentially expressed in the comparison of Control vs Switch (e.g. UdUd-UdSc)
Differentially expressed genes could either be:
1. Differences in the gene expression after the host switch correspond to differences found between the controls, indicating an adjustment to the new host (green)
2. No differences in gene expression after the switch despite differences between the controls, corresponding to the gene expression profiles on Host 1 (yellow)
3. Differences in gene expression after the switch but no differences between the controls (blue)
4. Differences in gene expression after the switch that are opposite to differences between controls (orange)
5. not significantly differently expressed (grey)

reference plot for expected differences was created with Inkscape

Note: N=Urtica dioica, S=Salix caprea, E=Ulmus glabra, G=Ribes uva-crispa

```{r scatterplots and GO-enrichtment analysis - Switch 1}

####NNEE-NNNE####
# 2hrs
NNNE2=full_join(DGE_1_2_gene$NNvEE$table %>% dplyr::select( genes, logFC_NNvEE=logFC, FDR_NNvEE=FDR),
                DGE_1_2_gene$NNvNE$table %>% dplyr::select( genes, logFC_NNvNE=logFC, FDR_NNvNE=FDR ))

plotNNNE1_2=
  ggplot()+
  geom_point(data=NNNE2, aes(x=logFC_NNvEE, y=logFC_NNvNE),color="#999999", alpha=0.30)+
  geom_point(data=NNNE2[NNNE2$FDR_NNvEE<0.05,], aes(x=logFC_NNvEE, y=logFC_NNvNE), color="#E6AB02", size=1)+ # no diff although expected --> "lag of plasticity", adjustment to host 1
  geom_point(data=NNNE2[NNNE2$FDR_NNvNE<0.05,], aes(x=logFC_NNvEE, y=logFC_NNvNE), color="#3a538b", size=1)+  # diff that were not expected --> "stress" or transient plasticity
  geom_point(data=NNNE2[NNNE2$FDR_NNvEE<0.05& NNNE2$FDR_NNvNE<0.05 & #diff as expected by Hx and Hy --> "host 2 specific"
                          ((NNNE2$logFC_NNvEE < 0 & NNNE2$logFC_NNvNE < 0) |
                             (NNNE2$logFC_NNvEE > 0 & NNNE2$logFC_NNvNE > 0)),],
             aes(x=logFC_NNvEE, y=logFC_NNvNE), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=NNNE2[NNNE2$FDR_NNvEE<0.05 & NNNE2$FDR_NNvNE<0.05 &
                          ((NNNE2$logFC_NNvEE < 0 & NNNE2$logFC_NNvNE > 0) |
                             (NNNE2$logFC_NNvEE > 0 & NNNE2$logFC_NNvNE < 0)),],
             aes(x=logFC_NNvEE, y=logFC_NNvNE ), color="#ef6c23", size=1)+   # slope = -1 line
  labs(title="2hrs", x=(expression(paste("log FC Ud",bold("Ud"),"-Ug",bold("Ug")))), y=(expression(paste("log FC Ud",bold("Ud"),"-Ud",bold("Ug"))))) +
  theme_bw()+
  xlim(-15,15)+
  ylim(-15, 15)+
  theme(aspect.ratio = 1)

#topGO
yellow=NNNE2[NNNE2$FDR_NNvEE<0.05,] #7
blue=NNNE2[NNNE2$FDR_NNvNE<0.05,] #0
green=NNNE2[NNNE2$FDR_NNvEE<0.05& NNNE2$FDR_NNvNE<0.05,]#0  --> this code would also give differentially expressed genes that fall in category 4 (i.e.orange); to distinguish green from orange--> see line below
orange=NNNE2[NNNE2$FDR_NNvEE<0.05 & NNNE2$FDR_NNvNE<0.05 & ((NNNE2$logFC_NNvEE < 0 & NNNE2$logFC_NNvNE > 0) |(NNNE2$logFC_NNvEE > 0 & NNNE2$logFC_NNvNE < 0)),] #0

NNNE2_list=list(C2=yellow$genes)

allResNNNE2 <- list()
for (i in 1:1) {
  candidate_list<- NNNE2_list[[i]]
  annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"

  
  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResNNNE2[[i]] <- GenTable(myGOdata.bv,
                               parentchildFisher = resultParentchild,
                               orderBy = "parentchildFisher",
                               topNodes = sum(resultParentchild@score < 1)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResNNNE2) <- names(NNNE2_list)


write.xlsx(allResNNNE2, file="NNNE_GO_S1.xlsx", sheetName = "NNNE2_C2", col.names=TRUE, row.names=TRUE, append=FALSE) # the names of these files might need to be changed to Latin abbreviations


allRes_plotsNNNE2 <-list()
for(i in 1:1) {
  allRes_plotsNNNE2[[i]] <- ggplot(head(allResNNNE2[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allRes_plotsNNNE2)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_NNNE2=ggarrange(plotlist = allRes_plotsNNNE2, ncol = 3, nrow = 1,labels=c("missing", "unexpected", "expected"), vjust=1 )
annotate_figure(GO_NNNE2, top=text_grob("NNNE2"))


#17 hrs

NNNE17=NNNE17=full_join(DGE_1_17_gene$NNvEE$table %>% dplyr::select( genes, logFC_NNvEE=logFC, FDR_NNvEE=FDR),
                        DGE_1_17_gene$NNvNE$table %>% dplyr::select( genes, logFC_NNvNE=logFC, FDR_NNvNE=FDR ))

plotNNNE1_17=ggplot()+
  geom_point(data=NNNE17, aes(x=logFC_NNvEE, y=logFC_NNvNE),color="#999999", alpha=0.30)+
  geom_point(data=NNNE17[NNNE17$FDR_NNvEE<0.05,], aes(x=logFC_NNvEE, y=logFC_NNvNE), color="#E6AB02", size=1)+
  geom_point(data=NNNE17[NNNE17$FDR_NNvNE<0.05,], aes(x=logFC_NNvEE, y=logFC_NNvNE), color="#3a538b", size=1)+
  #  geom_point(data=NNNE17[NNNE17$FDR_NNvEE<0.05& NNNE17$FDR_NNvNE<0.05,], aes(x=logFC_NNvEE, y=logFC_NNvNE), color="#21a585")+
  geom_point(data=NNNE17[NNNE17$FDR_NNvEE<0.05& NNNE17$FDR_NNvNE<0.05 & #diff as expected by Hx and Hy --> "host specific"
                           ((NNNE17$logFC_NNvEE < 0 & NNNE17$logFC_NNvNE < 0) |
                              (NNNE17$logFC_NNvEE > 0 & NNNE17$logFC_NNvNE > 0)),],
             aes(x=logFC_NNvEE, y=logFC_NNvNE), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=NNNE17[NNNE17$FDR_NNvEE<0.05 & NNNE17$FDR_NNvNE<0.05 &
                           ((NNNE17$logFC_NNvEE < 0 & NNNE17$logFC_NNvNE > 0) |
                              (NNNE17$logFC_NNvEE > 0 & NNNE17$logFC_NNvNE < 0)),],
             aes(x=logFC_NNvEE, y=logFC_NNvNE ), color="#ef6c23", size=1)+
  labs(title="17hrs", x=(expression(paste("log FC Ud",bold("Ud"),"-Ug",bold("Ug")))), y=(expression(paste("log FC Ud",bold("Ud"),"-Ud",bold("Ug"))))) +
  theme_bw()+
  xlim(-15,15)+
  ylim(-15, 15)+
  theme(aspect.ratio = 1)

  yellow=NNNE17[NNNE17$FDR_NNvEE<0.05,] #0
  blue=NNNE17[NNNE17$FDR_NNvNE<0.05,] #3
  green=NNNE17[NNNE17$FDR_NNvEE<0.05& NNNE17$FDR_NNvNE<0.05,]#0
  orange=NNNE17[NNNE17$FDR_NNvEE<0.05 & NNNE17$FDR_NNvNE<0.05 & ((NNNE17$logFC_NNvEE < 0 & NNNE17$logFC_NNvNE > 0) |(NNNE17$logFC_NNvEE > 0 & NNNE17$logFC_NNvNE < 0)),] #0

  NNNE17_list=list(C3=blue$genes)

  allResNNNE17 <- list()
  for (i in 1:1) {
    candidate_list<- NNNE17_list[[i]]
    annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

    # here I will be only analyzing GO terms with at least 5 members,
    # as this yield more stable results.
    node_size=5

    # We will focus on Biological prmocess terms (BP)
    GO_category="BP"
    geneID2GO <- readMappings(annotation)
    names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
    geneUniverse <- names(geneID2GO)

    genesOfInterest.bv <- candidate_list

    genesOfInterest.bv <- as.character(genesOfInterest.bv)
    geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
    names(geneList.bv) <- geneUniverse

    myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

    # STATS#
    # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
    # and so reduces false positives due to the inheritance problem
    resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

    allResNNNE17[[i]] <- GenTable(myGOdata.bv,
                                  parentchildFisher = resultParentchild,
                                  orderBy = "parentchildFisher",
                                  topNodes = sum(resultParentchild@score < 1)) %>%
      mutate(parentchildFisher = as.numeric(parentchildFisher))
  }

  names(allResNNNE17) <- names(NNNE17_list)


  write.xlsx(allResNNNE17, file="NNNE_GO_S1.xlsx", sheetName = "NNNE17_C3", col.names=TRUE, row.names=TRUE, append=TRUE)

  allRes_plotsNNNE17 <-list()
  for(i in 1:1) {
    allRes_plotsNNNE17[[i]] <- ggplot(head(allResNNNE17[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
      geom_point(aes(size = 100*Significant/Annotated))+
      labs(title = names(allRes_plotsNNNE17)[i], x =expression('-log'[10]~'(p-value)'))+
      theme(legend.position = "none")
  }

  GO_NNNE17=ggarrange(plotlist = allRes_plotsNNNE17, ncol = 3, nrow = 1,labels=c("missing", "unexpected", "expected"), vjust=1 )
  annotate_figure(GO_NNNE17, top=text_grob("NNNE17"))


#===========================================================================================
#### EENN- EEEN####
#2hrs
EEEN12=full_join(DGE_1_2_gene$EEvNN$table %>% dplyr::select( genes, logFC_EEvNN=logFC, FDR_EEvNN=FDR),
                 DGE_1_2_gene$EEvEN$table %>% dplyr::select( genes, logFC_EEvEN=logFC, FDR_EEvEN=FDR ))

plotEEEN1_2=ggplot()+
  geom_point(data=EEEN12, aes(x=logFC_EEvNN, y=logFC_EEvEN),color="#999999", alpha=0.30)+
  geom_point(data=EEEN12[EEEN12$FDR_EEvNN<0.05,], aes(x=logFC_EEvNN, y=logFC_EEvEN), color="#E6AB02", size=1)+ #=#F0E442 -> no diff although expected --> "lag of plasticity"
  geom_point(data=EEEN12[EEEN12$FDR_EEvEN<0.05,], aes(x=logFC_EEvNN, y=logFC_EEvEN), color="#3a538b", size=1)+  # #3a538b -> not expected differences --> "stress"
  geom_point(data=EEEN12[EEEN12$FDR_EEvNN<0.05 & EEEN12$FDR_EEvEN<0.05 &
                           ((EEEN12$logFC_EEvNN < 0 & EEEN12$logFC_EEvEN < 0) |
                              (EEEN12$logFC_EEvNN > 0 & EEEN12$logFC_EEvEN > 0)),],
             aes(x=logFC_EEvNN, y=logFC_EEvEN), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=EEEN12[EEEN12$FDR_EEvNN<0.05 & EEEN12$FDR_EEvEN<0.05 &
                           ((EEEN12$logFC_EEvNN < 0 & EEEN12$logFC_EEvEN > 0) |
                              (EEEN12$logFC_EEvNN > 0 & EEEN12$logFC_EEvEN < 0)),],
             aes(x=logFC_EEvNN, y=logFC_EEvEN ), color="#ef6c23", size=1)+
  labs(title="2hrs", x=(expression(paste("log FC Ug",bold("Ug"),"-Ud",bold("Ud")))), y=(expression(paste("log FC Ug",bold("Ug"),"-Ug",bold("Ud"))))) +
  theme_bw()+
  xlim(-15,15)+
  ylim(-15, 15)+
  theme(aspect.ratio = 1)

#topGO
yellow=EEEN12[EEEN12$FDR_EEvNN<0.05,] #7
blue=EEEN12[EEEN12$FDR_EEvEN<0.05,] #3304
green=EEEN12[EEEN12$FDR_EEvNN<0.05& EEEN12$FDR_EEvEN<0.05,]#2
green=EEEN12[EEEN12$FDR_EEvNN<0.05& EEEN12$FDR_EEvEN<0.05 & ((EEEN12$logFC_EEvNN < 0 & EEEN12$logFC_EEvEN < 0) |
                           (EEEN12$logFC_EEvNN > 0 & EEEN12$logFC_EEvEN > 0)),] #0
orange=EEEN12[EEEN12$FDR_EEvNN<0.05 & EEEN12$FDR_EEvEN<0.05 & ((EEEN12$logFC_EEvNN < 0 & EEEN12$logFC_EEvEN > 0) |(EEEN12$logFC_EEvNN > 0 & EEEN12$logFC_EEvEN < 0)),] #0


EEEN12_list=list(C2=yellow$genes,
                 C3=blue$genes)



allResEEEN12 <- list()
for (i in 1:2) {   
  candidate_list<- EEEN12_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResEEEN12[[i]] <- GenTable(myGOdata.bv,
                                parentchildFisher = resultParentchild ,
                                orderBy = "parentchildFisher",
                                topNodes = sum(resultParentchild@score < 1)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResEEEN12) <- names(EEEN12_list)

write.xlsx(allResEEEN12$C2, file="EEEN_GO_A.xlsx", sheetName = "EEEN2_C2", col.names=TRUE, row.names=TRUE, append=FALSE)
write.xlsx(allResEEEN12$C3, file="EEEN_GO_A.xlsx", sheetName = "EEEN2_C3", col.names=TRUE, row.names=TRUE, append=TRUE)



allRes_plotsEEEN12 <-list()
for(i in 1:2) {
  allRes_plotsEEEN12[[i]] <- ggplot(head(allResEEEN12[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResEEEN12)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_EEEN12=ggarrange(plotlist = allRes_plotsEEEN12, ncol = 1, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_EEEN12, top=text_grob("EEEN12"))

#17hrs
EEEN117=full_join(DGE_1_17_gene$EEvNN$table %>% dplyr::select( genes, logFC_EEvNN=logFC, FDR_EEvNN=FDR),
                  DGE_1_17_gene$EEvEN$table %>% dplyr::select( genes, logFC_EEvEN=logFC, FDR_EEvEN=FDR ))

plotEEEN1_17=ggplot()+
  geom_point(data=EEEN117, aes(x=logFC_EEvNN, y=logFC_EEvEN),color="#999999", alpha=0.30)+
  geom_point(data=EEEN117[EEEN117$FDR_EEvNN<0.05,], aes(x=logFC_EEvNN, y=logFC_EEvEN), color="#E6AB02", size=1)+
  geom_point(data=EEEN117[EEEN117$FDR_EEvEN<0.05,], aes(x=logFC_EEvNN, y=logFC_EEvEN), color="#3a538b", size=1)+
  geom_point(data=EEEN117[EEEN117$FDR_EEvNN<0.05& EEEN117$FDR_EEvEN<0.05 &
                            ((EEEN117$logFC_EEvNN < 0 & EEEN117$logFC_EEvEN < 0) |
                               (EEEN117$logFC_EEvNN > 0 & EEEN117$logFC_EEvEN > 0)),],
             aes(x=logFC_EEvNN, y=logFC_EEvEN), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=EEEN117[EEEN117$FDR_EEvNN<0.05 & EEEN117$FDR_EEvEN<0.05 &
                            ((EEEN117$logFC_EEvNN < 0 & EEEN117$logFC_EEvEN > 0) |
                               (EEEN117$logFC_EEvNN > 0 & EEEN117$logFC_EEvEN < 0)),],
             aes(x=logFC_EEvNN, y=logFC_EEvEN ), color="#ef6c23", size=1)+
  labs(title="17hrs", x=(expression(paste("log FC Ug",bold("Ug"),"-Ud",bold("Ud")))), y=(expression(paste("log FC Ug",bold("Ug"),"-Ug",bold("Ud"))))) +
  theme_bw()+
  xlim(-15,15)+
  ylim(-15, 15)+
  theme(aspect.ratio = 1)


  #17

  yellow=EEEN117[EEEN117$FDR_EEvNN<0.05,] #0
  blue=EEEN117[EEEN117$FDR_EEvEN<0.05,] #0
  green=EEEN117[EEEN117$FDR_EEvNN<0.05& EEEN117$FDR_EEvEN<0.05,]#0

  ### no significantly differentially expressed genes  --> so no Gene enrichment analysis possible/necessary


#===========================================================================================
####EESS-EEES####
#2hrs

EEES12=full_join(DGE_1_2_gene$EEvSS$table %>% dplyr::select( genes, logFC_EEvSS=logFC, FDR_EEvSS=FDR),
                 DGE_1_2_gene$EEvES$table %>% dplyr::select( genes, logFC_EEvES=logFC, FDR_EEvES=FDR ))

plotEEES1_2=ggplot()+
  geom_point(data=EEES12, aes(x=logFC_EEvSS, y=logFC_EEvES),color="#999999", alpha=0.30)+
  geom_point(data=EEES12[EEES12$FDR_EEvSS<0.05,], aes(x=logFC_EEvSS, y=logFC_EEvES), color="#E6AB02", size=1)+ #=#F0E442 -> no diff although expected --> "lag of plasticity"
  geom_point(data=EEES12[EEES12$FDR_EEvES<0.05,], aes(x=logFC_EEvSS, y=logFC_EEvES), color="#3a538b", size=1)+  # #3a538b -> not expected differences --> "stress"
  geom_point(data=EEES12[EEES12$FDR_EEvSS<0.05& EEES12$FDR_EEvES<0.05 &
                           ((EEES12$logFC_EEvSS < 0 & EEES12$logFC_EEvES < 0) |
                              (EEES12$logFC_EEvSS > 0 & EEES12$logFC_EEvES > 0)),],
             aes(x=logFC_EEvSS, y=logFC_EEvES), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=EEES12[EEES12$FDR_EEvSS<0.05 & EEES12$FDR_EEvES<0.05 &
                           ((EEES12$logFC_EEvSS < 0 & EEES12$logFC_EEvES > 0) |
                              (EEES12$logFC_EEvSS > 0 & EEES12$logFC_EEvES < 0)),],
             aes(x=logFC_EEvSS, y=logFC_EEvES ), color="#ef6c23", size=1)+
  labs(title="2hrs", x=(expression(paste("log FC Ug",bold("Ug"),"-Sc",bold("Sc")))), y=(expression(paste("log FC Ug",bold("Ug"),"-Ug",bold("Sc"))))) +
  theme_bw()+
  xlim(-15,15)+
  ylim(-15, 15)+
  theme(aspect.ratio = 1)

#topGO

yellow=EEES12[EEES12$FDR_EEvSS<0.05,] #5
blue=EEES12[EEES12$FDR_EEvES<0.05,] #0
green=EEES12[EEES12$FDR_EEvSS<0.05& EEES12$FDR_EEvES<0.05,]#0
orange=EEES12[EEES12$FDR_EEvSS<0.05 & EEES12$FDR_EEvES<0.05 &
                         ((EEES12$logFC_EEvSS < 0 & EEES12$logFC_EEvES > 0) |
                            (EEES12$logFC_EEvSS > 0 & EEES12$logFC_EEvES < 0)),] #0
EEES12_list=list(c2=yellow$genes)

allResEEES12 <- list()
for (i in 1:1) {
  candidate_list<- EEES12_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    


  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResEEES12[[i]] <- GenTable(myGOdata.bv,
                                parentchildFisher = resultParentchild ,
                                orderBy = "parentchildFisher",
                                topNodes = sum(resultParentchild@score < 0.05)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResEEES12) <- names(EEES12_list)

write.xlsx(allResEEEN12$C2, file="EEES_GO_A.xlsx", sheetName = "EEES2_C2", col.names=TRUE, row.names=TRUE, append=FALSE)



allRes_plotsEEES12 <-list()
for(i in 1:1) {
  allRes_plotsEEES12[[i]] <- ggplot(head(allResEEES12[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResEEES12)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_EEES12=ggarrange(plotlist = allRes_plotsEEES12, ncol = 1, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_EEES12, top=text_grob("EEES12"))


#17hrs
EEES117=full_join(DGE_1_17_gene$EEvSS$table %>% dplyr::select( genes, logFC_EEvSS=logFC, FDR_EEvSS=FDR),
                  DGE_1_17_gene$EEvES$table %>% dplyr::select( genes, logFC_EEvES=logFC, FDR_EEvES=FDR ))

plotEEES1_17=ggplot()+
  geom_point(data=EEES117, aes(x=logFC_EEvSS, y=logFC_EEvES),color="#999999", alpha=0.30)+
  geom_point(data=EEES117[EEES117$FDR_EEvSS<0.05,], aes(x=logFC_EEvSS, y=logFC_EEvES), color="#E6AB02", size=1)+
  geom_point(data=EEES117[EEES117$FDR_EEvES<0.05,], aes(x=logFC_EEvSS, y=logFC_EEvES), color="#3a538b", size=1)+
  geom_point(data=EEES117[EEES117$FDR_EEvSS<0.05& EEES117$FDR_EEvES<0.05 &
                            ((EEES117$logFC_EEvSS < 0 & EEES117$logFC_EEvES < 0) |
                               (EEES117$logFC_EEvSS > 0 & EEES117$logFC_EEvES > 0)),],
             aes(x=logFC_EEvSS, y=logFC_EEvES), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=EEES117[EEES117$FDR_EEvSS<0.05 & EEES117$FDR_EEvES<0.05 &
                            ((EEES117$logFC_EEvSS < 0 & EEES117$logFC_EEvES > 0) |
                               (EEES117$logFC_EEvSS > 0 & EEES117$logFC_EEvES < 0)),],
             aes(x=logFC_EEvSS, y=logFC_EEvES ), color="#ef6c23", size=1)+
  labs(title="17hrs", x=(expression(paste("log FC Ug",bold("Ug"),"-Sc",bold("Sc")))), y=(expression(paste("log FC Ug",bold("Ug"),"-Ug",bold("Sc"))))) +
  theme_bw()+
  xlim(-15,15)+
  ylim(-15, 15)+
  theme(aspect.ratio = 1)

#topGO


yellow=EEES117[EEES117$FDR_EEvSS<0.05,] #3
blue=EEES117[EEES117$FDR_EEvES<0.05,] #0
green=EEES117[EEES117$FDR_EEvSS<0.05& EEES117$FDR_EEvES<0.05,]#0
orange=EEES117[EEES117$FDR_EEvSS<0.05 & EEES117$FDR_EEvES<0.05 &
                 ((EEES117$logFC_EEvSS < 0 & EEES117$logFC_EEvES > 0) |
                    (EEES117$logFC_EEvSS > 0 & EEES117$logFC_EEvES < 0)),]#0

EEES117_list=list(c2=yellow$genes)


allResEEES117 <- list()
for (i in 1:1) {
  candidate_list<- EEES117_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResEEES117[[i]] <- GenTable(myGOdata.bv,
                                 parentchildFisher = resultParentchild ,
                                 orderBy = "parentchildFisher",
                                 topNodes = sum(resultParentchild@score < 0.05)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResEEES117) <- names(EEES117_list)

write.xlsx(allResEEES117$c2, file="EEES_GO_A.xlsx", sheetName = "EEES17_c2", col.names = TRUE, row.names = TRUE, append = TRUE)


allRes_plotsEEES117 <-list()
for(i in 1:1) {
  allRes_plotsEEES117[[i]] <- ggplot(head(allResEEES117[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResEEES117)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_EEES117=ggarrange(plotlist = allRes_plotsEEES117, ncol = 1, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_EEES117, top=text_grob("EEES117"))



#===========================================================================================
#### SSEE - SSSE ####
#2hr

SSSE2=full_join(DGE_1_2_gene$SSvEE$table %>% dplyr::select( genes, logFC_SSvEE=logFC, FDR_SSvEE=FDR),
                DGE_1_2_gene$SSvSE$table %>% dplyr::select( genes, logFC_SSvSE=logFC, FDR_SSvSE=FDR ))

plotSSSE1_2=ggplot()+
  geom_point(data=SSSE2, aes(x=logFC_SSvEE, y=logFC_SSvSE),color="#999999", alpha=0.30)+
  geom_point(data=SSSE2[SSSE2$FDR_SSvEE<0.05,], aes(x=logFC_SSvEE, y=logFC_SSvSE), color="#E6AB02", size=1)+ #=#F0E442 -> no diff although expected --> "lag of plasticity"
  geom_point(data=SSSE2[SSSE2$FDR_SSvSE<0.05,], aes(x=logFC_SSvEE, y=logFC_SSvSE), color="#3a538b", size=1)+  # #3a538b -> not expected differences --> "stress"
  geom_point(data=SSSE2[SSSE2$FDR_SSvEE<0.05& SSSE2$FDR_SSvSE<0.05 &
                          ((SSSE2$logFC_SSvEE < 0 & SSSE2$logFC_SSvSE < 0) |
                             (SSSE2$logFC_SSvEE > 0 & SSSE2$logFC_SSvSE > 0)),],
             aes(x=logFC_SSvEE, y=logFC_SSvSE), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=SSSE2[SSSE2$FDR_SSvEE<0.05 & SSSE2$FDR_SSvSE<0.05 &
                          ((SSSE2$logFC_SSvEE < 0 & SSSE2$logFC_SSvSE > 0) |
                             (SSSE2$logFC_SSvEE > 0 & SSSE2$logFC_SSvSE < 0)),],
             aes(x=logFC_SSvEE, y=logFC_SSvSE ), color="#ef6c23", size=1)+
  labs(title="2hrs", x=(expression(paste("log FC Sc",bold("Sc"),"-Ug",bold("Ug")))), y=(expression(paste("log FC Sc",bold("Sc"),"-Sc",bold("Ug"))))) +
  theme_bw()+
  xlim(-15,15)+
  ylim(-15, 15)+
  theme(aspect.ratio = 1)

#topGO
yellow=SSSE2[SSSE2$FDR_SSvEE<0.05,] #5
blue=SSSE2[SSSE2$FDR_SSvSE<0.05,] #0
green=SSSE2[SSSE2$FDR_SSvEE<0.05& SSSE2$FDR_SSvSE<0.05,]#0
orange=SSSE2[SSSE2$FDR_SSvEE<0.05 & SSSE2$FDR_SSvSE<0.05 &
               ((SSSE2$logFC_SSvEE < 0 & SSSE2$logFC_SSvSE > 0) |
                  (SSSE2$logFC_SSvEE > 0 & SSSE2$logFC_SSvSE < 0)),] #0

SSSE2_list=list(c2=yellow$genes)


allResSSSE2 <- list()
for (i in 1:1) {
  candidate_list<- SSSE2_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResSSSE2[[i]] <- GenTable(myGOdata.bv,
                               parentchildFisher = resultParentchild ,
                               orderBy = "parentchildFisher",
                               topNodes = sum(resultParentchild@score < 0.05)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResSSSE2) <- names(SSSE2_list)

write.xlsx(allResSSSE2$c2, file="SSSE_GO_A.xlsx", sheetName = "SSSE2_C2", col.names = TRUE, row.names = TRUE, append = FALSE)


allRes_plotsSSSE2 <-list()
for(i in 1:1) {
  allRes_plotsSSSE2[[i]] <- ggplot(head(allResSSSE2[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResSSSE2)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_SSSE2=ggarrange(plotlist = allRes_plotsSSSE2, ncol = 1, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_SSSE2, top=text_grob("SSSE2"))


#17
yellow=SSSE17[SSSE17$FDR_SSvEE<0.05,] #3
blue=SSSE17[SSSE17$FDR_SSvSE<0.05,] #0
green=SSSE17[SSSE17$FDR_SSvEE<0.05& SSSE17$FDR_SSvSE<0.05,]#0

SSSE17_list=list(c2=yellow$genes)


#17hrs
SSSE17=full_join(DGE_1_17_gene$SSvEE$table %>% dplyr::select( genes, logFC_SSvEE=logFC, FDR_SSvEE=FDR),
                 DGE_1_17_gene$SSvSE$table %>% dplyr::select( genes, logFC_SSvSE=logFC, FDR_SSvSE=FDR ))

plotSSSE1_17=ggplot()+
  geom_point(data=SSSE17, aes(x=logFC_SSvEE, y=logFC_SSvSE),color="#999999", alpha=0.30)+
  geom_point(data=SSSE17[SSSE17$FDR_SSvEE<0.05,], aes(x=logFC_SSvEE, y=logFC_SSvSE), color="#E6AB02", size=1)+
  geom_point(data=SSSE17[SSSE17$FDR_SSvSE<0.05,], aes(x=logFC_SSvEE, y=logFC_SSvSE), color="#3a538b", size=1)+
  geom_point(data=SSSE17[SSSE17$FDR_SSvEE<0.05& SSSE17$FDR_SSvSE<0.05&
                           ((SSSE17$logFC_SSvEE < 0 & SSSE17$logFC_SSvSE < 0) |
                              (SSSE17$logFC_SSvEE > 0 & SSSE17$logFC_SSvSE > 0)),],
             aes(x=logFC_SSvEE, y=logFC_SSvSE), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=SSSE17[SSSE17$FDR_SSvEE<0.05 & SSSE17$FDR_SSvSE<0.05 &
                           ((SSSE17$logFC_SSvEE < 0 & SSSE17$logFC_SSvSE > 0) |
                              (SSSE17$logFC_SSvEE > 0 & SSSE17$logFC_SSvSE < 0)),],
             aes(x=logFC_SSvEE, y=logFC_SSvSE ), color="#ef6c23", size=1)+
  labs(title="17hrs", x=(expression(paste("log FC Sc",bold("Sc"),"-Ug",bold("Ug")))), y=(expression(paste("log FC Sc",bold("Sc"),"-Sc",bold("Ug"))))) +
  theme_bw()+
  xlim(-15,15)+
  ylim(-15, 15)+
  theme(aspect.ratio = 1)

#topGO
yellow=SSSE17[SSSE17$FDR_SSvEE<0.05,] #3
blue=SSSE17[SSSE17$FDR_SSvSE<0.05,] #0
green=SSSE17[SSSE17$FDR_SSvEE<0.05& SSSE17$FDR_SSvSE<0.05,]#0
orange=SSSE17[SSSE17$FDR_SSvEE<0.05 & SSSE17$FDR_SSvSE<0.05 &
                ((SSSE17$logFC_SSvEE < 0 & SSSE17$logFC_SSvSE > 0) |
                   (SSSE17$logFC_SSvEE > 0 & SSSE17$logFC_SSvSE < 0)),]#0


SSSE17_list=list(c2=yellow$genes)


allResSSSE17 <- list()
for (i in 1:1) {
  candidate_list<- SSSE17_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResSSSE17[[i]] <- GenTable(myGOdata.bv,
                                parentchildFisher = resultParentchild ,
                                orderBy = "parentchildFisher",
                                topNodes = sum(resultParentchild@score < 0.05)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResSSSE17) <- names(SSSE17_list)

write.xlsx(allResSSSE17$c2, file="SSSE_GO_A.xlsx", sheetName = "SSSE17_c2", col.names = TRUE, row.names = TRUE, append=TRUE)



allRes_plotsSSSE17 <-list()
for(i in 1:1) {
  allRes_plotsSSSE17[[i]] <- ggplot(head(allResSSSE17[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResSSSE17)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_SSSE17=ggarrange(plotlist = allRes_plotsSSSE17, ncol = 1, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_SSSE17, top=text_grob("SSSE17"))

#===========================================================================================
####SSNN-SSSN 12 ####
#2hrs

SSSN12=full_join(DGE_1_2_gene$SSvNN$table %>% dplyr::select( genes, logFC_SSvNN=logFC, FDR_SSvNN=FDR),
                 DGE_1_2_gene$SSvSN$table %>% dplyr::select( genes, logFC_SSvSN=logFC, FDR_SSvSN=FDR ))

plotSSSN1_2=ggplot()+
  geom_point(data=SSSN12, aes(x=logFC_SSvNN, y=logFC_SSvSN),color="#999999", alpha=0.30)+
  geom_point(data=SSSN12[SSSN12$FDR_SSvNN<0.05,], aes(x=logFC_SSvNN, y=logFC_SSvSN), color="#E6AB02", size=1)+ #=#F0E442 -> no diff although expected --> "lag of plasticity"
  geom_point(data=SSSN12[SSSN12$FDR_SSvSN<0.05,], aes(x=logFC_SSvNN, y=logFC_SSvSN), color="#3a538b", size=1)+  # #3a538b -> not expected differences --> "stress"
  geom_point(data=SSSN12[SSSN12$FDR_SSvNN<0.05& SSSN12$FDR_SSvSN<0.05&
                           ((SSSN12$logFC_SSvNN < 0 & SSSN12$logFC_SSvSN < 0) |
                              (SSSN12$logFC_SSvNN > 0 & SSSN12$logFC_SSvSN > 0)),],
             aes(x=logFC_SSvNN, y=logFC_SSvSN), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=SSSN12[SSSN12$FDR_SSvNN<0.05 & SSSN12$FDR_SSvSN<0.05 &
                           ((SSSN12$logFC_SSvNN < 0 & SSSN12$logFC_SSvSN > 0) |
                              (SSSN12$logFC_SSvNN > 0 & SSSN12$logFC_SSvSN < 0)),],
             aes(x=logFC_SSvNN, y=logFC_SSvSN ), color="#ef6c23", size=1)+
  labs(title="2hrs", x=(expression(paste("log FC Sc",bold("Sc"),"-Ud",bold("Ud")))), y=(expression(paste("log FC Sc",bold("Sc"),"-Sc",bold("Ud"))))) +
  theme_bw()+
  xlim(-15,15)+
  ylim(-15, 15)+
  theme(aspect.ratio = 1)

#topGO
yellow=SSSN12[SSSN12$FDR_SSvNN<0.05,] #13
blue=SSSN12[SSSN12$FDR_SSvSN<0.05,] #0
green=SSSN12[SSSN12$FDR_SSvNN<0.05& SSSN12$FDR_SSvSN<0.05,]#0
orange=SSSN12[SSSN12$FDR_SSvNN<0.05& SSSN12$FDR_SSvSN<0.05&
                ((SSSN12$logFC_SSvNN < 0 & SSSN12$logFC_SSvSN < 0) |
                   (SSSN12$logFC_SSvNN > 0 & SSSN12$logFC_SSvSN > 0)),]#0

SSSN12_list=list(c2=yellow$genes)


allResSSSN12 <- list()
for (i in 1:1) {
  candidate_list<- SSSN12_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResSSSN12[[i]] <- GenTable(myGOdata.bv,
                                parentchildFisher = resultParentchild ,
                                orderBy = "parentchildFisher",
                                topNodes = sum(resultParentchild@score < 0.05)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResSSSN12) <- names(SSSN12_list)

write.xlsx(allResSSSN12$c2, file="SSSN_GO_A.xlsx", sheetName = "SSSN2_c2", col.names = TRUE, row.names = TRUE, append=FALSE)



allRes_plotsSSSN12 <-list()
for(i in 1:1) {
  allRes_plotsSSSN12[[i]] <- ggplot(head(allResSSSN12[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResSSSN12)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_SSSN12=ggarrange(plotlist = allRes_plotsSSSN12, ncol = 1, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_SSSN12, top=text_grob("SSSN12"))

#17hrs
SSSN117=full_join(DGE_1_17_gene$SSvNN$table %>% dplyr::select( genes, logFC_SSvNN=logFC, FDR_SSvNN=FDR),
                  DGE_1_17_gene$SSvSN$table %>% dplyr::select( genes, logFC_SSvSN=logFC, FDR_SSvSN=FDR ))

plotSSSN1_17=ggplot()+
  geom_point(data=SSSN117, aes(x=logFC_SSvNN, y=logFC_SSvSN),color="#999999", alpha=0.30)+
  geom_point(data=SSSN117[SSSN117$FDR_SSvNN<0.05,], aes(x=logFC_SSvNN, y=logFC_SSvSN), color="#E6AB02", size=1)+
  geom_point(data=SSSN117[SSSN117$FDR_SSvSN<0.05,], aes(x=logFC_SSvNN, y=logFC_SSvSN), color="#3a538b", size=1)+
  geom_point(data=SSSN117[SSSN117$FDR_SSvNN<0.05& SSSN117$FDR_SSvSN<0.05    &
                            ((SSSN117$logFC_SSvNN < 0 & SSSN117$logFC_SSvSN < 0) |
                               (SSSN117$logFC_SSvNN > 0 & SSSN117$logFC_SSvSN > 0)),],
             aes(x=logFC_SSvNN, y=logFC_SSvSN), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=SSSN117[SSSN117$FDR_SSvNN<0.05 & SSSN117$FDR_SSvSN<0.05 &
                            ((SSSN117$logFC_SSvNN < 0 & SSSN117$logFC_SSvSN > 0) |
                               (SSSN117$logFC_SSvNN > 0 & SSSN117$logFC_SSvSN < 0)),],
             aes(x=logFC_SSvNN, y=logFC_SSvSN ), color="#ef6c23", size=1)+
  labs(title="17hrs", x=(expression(paste("log FC Sc",bold("Sc"),"-Ud",bold("Ud")))), y=(expression(paste("log FC Sc",bold("Sc"),"-Sc",bold("Ud"))))) +
  theme_bw()+
  xlim(-15,15)+
  ylim(-15, 15)+
  theme(aspect.ratio = 1)

#topGO
yellow=SSSN117[SSSN117$FDR_SSvNN<0.05,] #3
blue=SSSN117[SSSN117$FDR_SSvSN<0.05,] #0
green=SSSN117[SSSN117$FDR_SSvNN<0.05& SSSN117$FDR_SSvSN<0.05,]#0
orange=SSSN117[SSSN117$FDR_SSvNN<0.05 & SSSN117$FDR_SSvSN<0.05 &
                 ((SSSN117$logFC_SSvNN < 0 & SSSN117$logFC_SSvSN > 0) |
                    (SSSN117$logFC_SSvNN > 0 & SSSN117$logFC_SSvSN < 0)),]#0

SSSN117_list=list(c2=yellow$genes)


allResSSSN117 <- list()
for (i in 1:1) {
  candidate_list<- SSSN117_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResSSSN117[[i]] <- GenTable(myGOdata.bv,
                                 parentchildFisher = resultParentchild ,
                                 orderBy = "parentchildFisher",
                                 topNodes = sum(resultParentchild@score < 0.05)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResSSSN117) <- names(SSSN117_list)

write.xlsx(allResSSSN117$c2, file="SSSN_GO_A.xlsx", sheetName = "SSSN17_c2", col.names = TRUE, row.names = TRUE, append=TRUE)


allRes_plotsSSSN117 <-list()
for(i in 1:1) {
  allRes_plotsSSSN117[[i]] <- ggplot(head(allResSSSN117[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResSSSN117)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_SSSN117=ggarrange(plotlist = allRes_plotsSSSN117, ncol = 1, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_SSSN117, top=text_grob("SSSN117"))

#===========================================================================================
####NNSS-NNNS####
#2hrs

NNNS12=full_join(DGE_1_2_gene$NNvSS$table %>% dplyr::select( genes, logFC_NNvSS=logFC, FDR_NNvSS=FDR),
                 DGE_1_2_gene$NNvNS$table %>% dplyr::select( genes, logFC_NNvNS=logFC, FDR_NNvNS=FDR ))

plotNNNS1_2=ggplot()+
  geom_point(data=NNNS12, aes(x=logFC_NNvSS, y=logFC_NNvNS),color="#999999", alpha=0.30)+
  geom_point(data=NNNS12[NNNS12$FDR_NNvSS<0.05,], aes(x=logFC_NNvSS, y=logFC_NNvNS), color="#E6AB02", size=1)+ #=#F0E442 -> no diff although expected --> "lag of plasticity"
  geom_point(data=NNNS12[NNNS12$FDR_NNvNS<0.05,], aes(x=logFC_NNvSS, y=logFC_NNvNS), color="#3a538b", size=1)+  # #3a538b -> not expected differences --> "stress"
  geom_point(data=NNNS12[NNNS12$FDR_NNvSS<0.05& NNNS12$FDR_NNvNS<0.05  &
                           ((NNNS12$logFC_NNvSS < 0 & NNNS12$logFC_NNvNS < 0) |
                              (NNNS12$logFC_NNvSS > 0 & NNNS12$logFC_NNvNS > 0)),],
             aes(x=logFC_NNvSS, y=logFC_NNvNS), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=NNNS12[NNNS12$FDR_NNvSS<0.05 & NNNS12$FDR_NNvNS<0.05 &
                           ((NNNS12$logFC_NNvSS < 0 & NNNS12$logFC_NNvNS > 0) |
                              (NNNS12$logFC_NNvSS > 0 & NNNS12$logFC_NNvNS < 0)),],
             aes(x=logFC_NNvSS, y=logFC_NNvNS ), color="#ef6c23", size=1)+
  labs(title="2hrs", x=(expression(paste("log FC Ud",bold("Ud"),"-Sc",bold("Sc")))), y=(expression(paste("log FC Ud",bold("Ud"),"-Ud",bold("Sc"))))) +
  theme_bw()+
  xlim(-15,15)+
  ylim(-15, 15)+
  theme(aspect.ratio = 1)

#topGO
yellow=NNNS12[NNNS12$FDR_NNvSS<0.05,] #13
blue=NNNS12[NNNS12$FDR_NNvNS<0.05,] #0
green=NNNS12[NNNS12$FDR_NNvSS<0.05& NNNS12$FDR_NNvNS<0.05,]#0
orange=NNNS12[NNNS12$FDR_NNvSS<0.05 & NNNS12$FDR_NNvNS<0.05 &
                ((NNNS12$logFC_NNvSS < 0 & NNNS12$logFC_NNvNS > 0) |
                   (NNNS12$logFC_NNvSS > 0 & NNNS12$logFC_NNvNS < 0)),]#0

NNNS12_list=list(c2=yellow$genes)


allResNNNS12 <- list()
for (i in 1:1) {
  candidate_list<- NNNS12_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResNNNS12[[i]] <- GenTable(myGOdata.bv,
                                parentchildFisher = resultParentchild ,
                                orderBy = "parentchildFisher",
                                topNodes = sum(resultParentchild@score < 0.05)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResNNNS12) <- names(NNNS12_list)

write.xlsx(allResNNNS12$c2, file="NNNS_GO_A.xlsx", sheetName = "NNNS2_c2", col.names = TRUE, row.names = TRUE, append=FALSE)


allRes_plotsNNNS12 <-list()
for(i in 1:1) {
  allRes_plotsNNNS12[[i]] <- ggplot(head(allResNNNS12[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResNNNS12)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_NNNS12=ggarrange(plotlist = allRes_plotsNNNS12, ncol = 1, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_NNNS12, top=text_grob("NNNS12"))


#17hrs
NNNS117=full_join(DGE_1_17_gene$NNvSS$table %>% dplyr::select( genes, logFC_NNvSS=logFC, FDR_NNvSS=FDR),
                  DGE_1_17_gene$NNvNS$table %>% dplyr::select( genes, logFC_NNvNS=logFC, FDR_NNvNS=FDR ))


plotNNNS1_17=ggplot()+
  geom_point(data=NNNS117, aes(x=logFC_NNvSS, y=logFC_NNvNS),color="#999999", alpha=0.30)+
  geom_point(data=NNNS117[NNNS117$FDR_NNvSS<0.05,], aes(x=logFC_NNvSS, y=logFC_NNvNS), color="#E6AB02", size=1)+
  geom_point(data=NNNS117[NNNS117$FDR_NNvNS<0.05,], aes(x=logFC_NNvSS, y=logFC_NNvNS), color="#3a538b", size=1)+
  geom_point(data=NNNS117[NNNS117$FDR_NNvSS<0.05& NNNS117$FDR_NNvNS<0.05 &
                            ((NNNS117$logFC_NNvSS < 0 & NNNS117$logFC_NNvNS < 0) |
                               (NNNS117$logFC_NNvSS > 0 & NNNS117$logFC_NNvNS > 0)),],
             aes(x=logFC_NNvSS, y=logFC_NNvNS), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=NNNS117[NNNS117$FDR_NNvSS<0.05 & NNNS117$FDR_NNvNS<0.05 &
                            ((NNNS117$logFC_NNvSS < 0 & NNNS117$logFC_NNvNS > 0) |
                               (NNNS117$logFC_NNvSS > 0 & NNNS117$logFC_NNvNS < 0)),],
             aes(x=logFC_NNvSS, y=logFC_NNvNS ), color="#ef6c23", size=1)+
  labs(title="17hrs", x=(expression(paste("log FC Ud",bold("Ud"),"-Sc",bold("Sc")))), y=(expression(paste("log FC Ud",bold("Ud"),"-Ud",bold("Sc"))))) +
  theme_bw()+
  xlim(-15,15)+
  ylim(-15, 15)+
  theme(aspect.ratio = 1)

#topGo
yellow=NNNS117[NNNS117$FDR_NNvSS<0.05,] #3
blue=NNNS117[NNNS117$FDR_NNvNS<0.05,] #0
green=NNNS117[NNNS117$FDR_NNvSS<0.05& NNNS117$FDR_NNvNS<0.05,]#0
orange=NNNS117[NNNS117$FDR_NNvSS<0.05 & NNNS117$FDR_NNvNS<0.05 &
                 ((NNNS117$logFC_NNvSS < 0 & NNNS117$logFC_NNvNS > 0) |
                    (NNNS117$logFC_NNvSS > 0 & NNNS117$logFC_NNvNS < 0)),]#0

NNNS117_list=list(c2=yellow$genes)


allResNNNS117 <- list()
for (i in 1:1) {
  candidate_list<- NNNS117_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResNNNS117[[i]] <- GenTable(myGOdata.bv,
                                 parentchildFisher = resultParentchild ,
                                 orderBy = "parentchildFisher",
                                 topNodes = sum(resultParentchild@score < 0.05)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResNNNS117) <- names(NNNS117_list)

write.xlsx(allResNNNS117$c2, file="NNNS_GO_A.xlsx", sheetName = "NNNS17_c2", col.names = TRUE, row.names = TRUE, append=TRUE)


allRes_plotsNNNS117 <-list()
for(i in 1:1) {
  allRes_plotsNNNS117[[i]] <- ggplot(head(allResNNNS117[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResNNNS117)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_NNNS117=ggarrange(plotlist = allRes_plotsNNNS117, ncol = 1, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_NNNS117, top=text_grob("NNNS117"))

#plot Switch 1

plot_S1_NE=ggarrange(plotNNNE1_2, plotNNNE1_17, plotEEEN1_2, plotEEEN1_17,  ncol = 2, nrow=2, labels=c("a", "b", "c", "d"))
plot_S1_SN=ggarrange(plotNNNS1_2, plotNNNS1_17, plotSSSN1_2, plotSSSN1_17,  ncol = 2, nrow=2,labels=c("e", "f", "g", "h" ))
plot_S1_SE=ggarrange(plotSSSE1_2, plotSSSE1_17, plotEEES1_2, plotEEES1_17,  ncol = 2, nrow=2, labels=c("i", "j", "k", "l" ))

ggarrange(plot_S1_SN, plot_S1_SE, plot_S1_NE, ncol=3, nrow=1)
```

```{r scatterplots and GO enrichment analysis -Switch 2}

####NNGG- NNNG ####
#2hrs

NNNG2=full_join(DGE_2_2_gene$NNvGG$table %>% dplyr::select( genes, logFC_NNvGG=logFC, FDR_NNvGG=FDR),
                DGE_2_2_gene$NNvNG$table %>% dplyr::select( genes, logFC_NNvNG=logFC, FDR_NNvNG=FDR ))

plotNNNG2=
  ggplot()+
  geom_point(data=NNNG2, aes(x=logFC_NNvGG, y=logFC_NNvNG),color="#999999", alpha=0.30)+
  geom_point(data=NNNG2[NNNG2$FDR_NNvGG<0.05,], aes(x=logFC_NNvGG, y=logFC_NNvNG), color="#E6AB02", size=1)+ #=#F0E442 -> no diff although expected --> "lag of plasticity"
  geom_point(data=NNNG2[NNNG2$FDR_NNvNG<0.05,], aes(x=logFC_NNvGG, y=logFC_NNvNG), color="#3a538b", size=1)+  # #3a538b -> not expected differences --> "stress"
  geom_point(data=NNNG2[NNNG2$FDR_NNvGG<0.05& NNNG2$FDR_NNvNG<0.05 & #diff as expected by Hx and Hy --> "host specific"
                      ((NNNG2$logFC_NNvGG < 0 & NNNG2$logFC_NNvNG < 0) |
                         (NNNG2$logFC_NNvGG > 0 & NNNG2$logFC_NNvNG > 0)),],
                    aes(x=logFC_NNvGG, y=logFC_NNvNG), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=NNNG2[NNNG2$FDR_NNvGG<0.05 & NNNG2$FDR_NNvNG<0.05 &
                      ((NNNG2$logFC_NNvGG < 0 & NNNG2$logFC_NNvNG > 0) |
                          (NNNG2$logFC_NNvGG > 0 & NNNG2$logFC_NNvNG < 0)),],
                      aes(x=logFC_NNvGG, y=logFC_NNvNG ), color="#ef6c23", size=1)+   # slope = -1 line
  labs(title="2hrs", x=(expression(paste("log FC U",bold("U"),"-R",bold("R")))), y=(expression(paste("log FC U",bold("U"),"-U",bold("R"))))) +
  theme_bw()+
  xlim(-12,12)+
  ylim(-12, 12)+
  theme(aspect.ratio = 1)

#topGO
yellow=NNNG2[NNNG2$FDR_NNvGG<0.05,] #136
blue=NNNG2[NNNG2$FDR_NNvNG<0.05,]
green=NNNG2[NNNG2$FDR_NNvGG<0.05& NNNG2$FDR_NNvNG<0.05,]
orange=NNNG2[NNNG2$FDR_NNvGG<0.05 & NNNG2$FDR_NNvNG<0.05 &
               ((NNNG2$logFC_NNvGG < 0 & NNNG2$logFC_NNvNG > 0) |
                  (NNNG2$logFC_NNvGG > 0 & NNNG2$logFC_NNvNG < 0)),] #0

NNNG2_list=list(c2=yellow$genes)



allResNNNG2 <- list()
for (i in 1:1) {
  candidate_list<- NNNG2_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResNNNG2[[i]] <- GenTable(myGOdata.bv,
                          parentchildFisher = resultParentchild,
                          orderBy = "parentchildFisher",
                          topNodes = sum(resultParentchild@score < 0.01)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResNNNG2) <- names(NNNG2_list)

write.xlsx(allResNNNG2$c2, file="NNNG_GO_B.xlsx", sheetName = "NNNG2_c2", col.names = TRUE, row.names = TRUE, append=FALSE)


allRes_plotsNNNG2 <-list()
for(i in 1:1) {
  allRes_plotsNNNG2[[i]] <- ggplot(head(allResNNNG2[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResNNNG2)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_NNNG2=ggarrange(plotlist = allRes_plotsNNNG2, ncol = 1, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_NNNG2, top=text_grob("NNNG2"))

#17hrs
NNNG17=NNNG17=full_join(DGE_2_17_gene$NNvGG$table %>% dplyr::select( genes, logFC_NNvGG=logFC, FDR_NNvGG=FDR),
                        DGE_2_17_gene$NNvNG$table %>% dplyr::select( genes, logFC_NNvNG=logFC, FDR_NNvNG=FDR ))

plotNNNG17=ggplot()+
  geom_point(data=NNNG17, aes(x=logFC_NNvGG, y=logFC_NNvNG),color="#999999", alpha=0.30)+
  geom_point(data=NNNG17[NNNG17$FDR_NNvGG<0.05,], aes(x=logFC_NNvGG, y=logFC_NNvNG), color="#E6AB02", size=1)+
  geom_point(data=NNNG17[NNNG17$FDR_NNvNG<0.05,], aes(x=logFC_NNvGG, y=logFC_NNvNG), color="#3a538b", size=1)+
#  geom_point(data=NNNG17[NNNG17$FDR_NNvGG<0.05& NNNG17$FDR_NNvNG<0.05,], aes(x=logFC_NNvGG, y=logFC_NNvNG), color="#21a585")+
  geom_point(data=NNNG17[NNNG17$FDR_NNvGG<0.05& NNNG17$FDR_NNvNG<0.05 & #diff as expected by Hx and Hy --> "host specific"
                          ((NNNG17$logFC_NNvGG < 0 & NNNG17$logFC_NNvNG < 0) |
                             (NNNG17$logFC_NNvGG > 0 & NNNG17$logFC_NNvNG > 0)),],
             aes(x=logFC_NNvGG, y=logFC_NNvNG), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=NNNG17[NNNG17$FDR_NNvGG<0.05 & NNNG17$FDR_NNvNG<0.05 &
                          ((NNNG17$logFC_NNvGG < 0 & NNNG17$logFC_NNvNG > 0) |
                             (NNNG17$logFC_NNvGG > 0 & NNNG17$logFC_NNvNG < 0)),],
             aes(x=logFC_NNvGG, y=logFC_NNvNG ), color="#ef6c23", size=1)+
  labs(title="17hrs", x=(expression(paste("log FC U",bold("U"),"-R",bold("R")))), y=(expression(paste("log FC U",bold("U"),"-U",bold("R"))))) +
  theme_bw()+
  xlim(-12,12)+
  ylim(-12,12)+
  theme(aspect.ratio = 1)

plot_NNNG=ggarrange(plotNNNG2, NULL, plotNNNG17, nrow=1, widths=c(1,0.01, 1))

#topGo
yellow=NNNG17[NNNG17$FDR_NNvGG<0.05,] #199
blue=NNNG17[NNNG17$FDR_NNvNG<0.05,] #1219
green=NNNG17[NNNG17$FDR_NNvGG<0.05& NNNG17$FDR_NNvNG<0.05,]#123
orange= NNNG17[NNNG17$FDR_NNvGG<0.05 & NNNG17$FDR_NNvNG<0.05 &
                 ((NNNG17$logFC_NNvGG < 0 & NNNG17$logFC_NNvNG > 0) |
                    (NNNG17$logFC_NNvGG > 0 & NNNG17$logFC_NNvNG < 0)),]#0


NNNG17_list=list(c1=green$genes,
                c2=yellow$genes,
                c3=blue$genes)

allResNNNG17 <- list()
for (i in 1:3) {
  candidate_list<- NNNG17_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResNNNG17[[i]] <- GenTable(myGOdata.bv,
                          parentchildFisher = resultParentchild,
                          orderBy = "parentchildFisher",
                          topNodes = sum(resultParentchild@score < 0.01)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResNNNG17) <- names(NNNG17_list)

write.xlsx(allResNNNG17$c1, file="NNNG_GO_B.xlsx", sheetName = "NNNG17_c1", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResNNNG17$c2, file="NNNG_GO_B.xlsx", sheetName = "NNNG17_c2", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResNNNG17$c3, file="NNNG_GO_B.xlsx", sheetName = "NNNG17_c3", col.names = TRUE, row.names = TRUE, append=TRUE)

allRes_plotsNNNG17 <-list()
for(i in 1:3) {
  allRes_plotsNNNG17[[i]] <- ggplot(head(allResNNNG17[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allRes_plotsNNNG17)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_NNNG17=ggarrange(plotlist = allRes_plotsNNNG17, ncol = 3, nrow = 1,labels=c("missing", "unexpected", "expected"), vjust=1 )
annotate_figure(GO_NNNG17, top=text_grob("NNNG17"))

#===========================================================================================
####GGNN - GGGN ####
#2hrs

GGGN2=full_join(DGE_2_2_gene$GGvNN$table %>% dplyr::select( genes, logFC_GGvNN=logFC, FDR_GGvNN=FDR),
                DGE_2_2_gene$GGvGN$table %>% dplyr::select( genes, logFC_GGvGN=logFC, FDR_GGvGN=FDR ))

plotGGGN2=ggplot()+
  geom_point(data=GGGN2, aes(x=logFC_GGvNN, y=logFC_GGvGN),color="#999999", alpha=0.30)+
  geom_point(data=GGGN2[GGGN2$FDR_GGvNN<0.05,], aes(x=logFC_GGvNN, y=logFC_GGvGN), color="#E6AB02", size=1)+ #=#F0E442 -> no diff although expected --> "lag of plasticity"
  geom_point(data=GGGN2[GGGN2$FDR_GGvGN<0.05,], aes(x=logFC_GGvNN, y=logFC_GGvGN), color="#3a538b", size=1)+  # #3a538b -> not expected differences --> "stress"
  geom_point(data=GGGN2[GGGN2$FDR_GGvNN<0.05 & GGGN2$FDR_GGvGN<0.05 &
                           ((GGGN2$logFC_GGvNN < 0 & GGGN2$logFC_GGvGN < 0) |
                             (GGGN2$logFC_GGvNN > 0 & GGGN2$logFC_GGvGN > 0)),],
             aes(x=logFC_GGvNN, y=logFC_GGvGN), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=GGGN2[GGGN2$FDR_GGvNN<0.05 & GGGN2$FDR_GGvGN<0.05 &
                          ((GGGN2$logFC_GGvNN < 0 & GGGN2$logFC_GGvGN > 0) |
                             (GGGN2$logFC_GGvNN > 0 & GGGN2$logFC_GGvGN < 0)),],
             aes(x=logFC_GGvNN, y=logFC_GGvGN ), color="#ef6c23", size=1)+
  labs(title="2hrs", x=(expression(paste("log FC R",bold("R"),"-U",bold("U")))), y=(expression(paste("log FC R",bold("R"),"-R",bold("U"))))) +
  theme_bw()+
  xlim(-12,12)+
  ylim(-12,12)+
  theme(aspect.ratio = 1)

#topGo
yellow=GGGN2[GGGN2$FDR_GGvNN<0.05,]#136
blue=GGGN2[GGGN2$FDR_GGvGN<0.05,]
green=GGGN2[GGGN2$FDR_GGvNN<0.05& GGGN2$FDR_GGvGN<0.05,]
orange=GGGN17[GGGN17$FDR_GGvNN<0.05 & GGGN17$FDR_GGvGN<0.05 &
                ((GGGN17$logFC_GGvNN < 0 & GGGN17$logFC_GGvGN > 0) |
                   (GGGN17$logFC_GGvNN > 0 & GGGN17$logFC_GGvGN < 0)),]#0

GGGN2_list=list(c2=yellow$genes)



allResGGGN2 <- list()
for (i in 1:1) {
  candidate_list<- GGGN2_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResGGGN2[[i]] <- GenTable(myGOdata.bv,
                               parentchildFisher = resultParentchild,
                               orderBy = "parentchildFisher",
                               topNodes = sum(resultParentchild@score < 0.01)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResGGGN2) <- names(GGGN2_list)

write.xlsx(allResGGGN2$c2, file="GGGN_GO_B.xlsx", sheetName = "GGGN2_c2", col.names = TRUE, row.names = TRUE, append=FALSE)

allRes_plotsGGGN2 <-list()
for(i in 1:1) {
  allRes_plotsGGGN2[[i]] <- ggplot(head(allResGGGN2[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResGGGN2)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_GGGN2=ggarrange(plotlist = allRes_plotsGGGN2, ncol = 2, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_GGGN2, top=text_grob("GGGN2"))


#17hrs
GGGN17=GGGN17=full_join(DGE_2_17_gene$GGvNN$table %>% dplyr::select( genes, logFC_GGvNN=logFC, FDR_GGvNN=FDR),
                        DGE_2_17_gene$GGvGN$table %>% dplyr::select( genes, logFC_GGvGN=logFC, FDR_GGvGN=FDR ))

plotGGGN17=ggplot()+
  geom_point(data=GGGN17, aes(x=logFC_GGvNN, y=logFC_GGvGN),color="#999999", alpha=0.30)+
  geom_point(data=GGGN17[GGGN17$FDR_GGvNN<0.05,], aes(x=logFC_GGvNN, y=logFC_GGvGN), color="#E6AB02", size=1)+
  geom_point(data=GGGN17[GGGN17$FDR_GGvGN<0.05,], aes(x=logFC_GGvNN, y=logFC_GGvGN), color="#3a538b", size=1)+
  geom_point(data=GGGN17[GGGN17$FDR_GGvNN<0.05& GGGN17$FDR_GGvGN<0.05 &
             ((GGGN17$logFC_GGvNN < 0 & GGGN17$logFC_GGvGN < 0) |
                (GGGN17$logFC_GGvNN > 0 & GGGN17$logFC_GGvGN > 0)),],
aes(x=logFC_GGvNN, y=logFC_GGvGN), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=GGGN17[GGGN17$FDR_GGvNN<0.05 & GGGN17$FDR_GGvGN<0.05 &
                          ((GGGN17$logFC_GGvNN < 0 & GGGN17$logFC_GGvGN > 0) |
                             (GGGN17$logFC_GGvNN > 0 & GGGN17$logFC_GGvGN < 0)),],
             aes(x=logFC_GGvNN, y=logFC_GGvGN ), color="#ef6c23", size=1)+
  labs(title="17hrs", x=(expression(paste("log FC R",bold("R"),"-U",bold("U")))), y=(expression(paste("log FC R",bold("R"),"-R",bold("U"))))) +
  theme_bw()+
  xlim(-12,12)+
  ylim(-12,12)+
  theme(aspect.ratio = 1)

#topG0
yellow=GGGN17[GGGN17$FDR_GGvNN<0.05,]#199
blue=GGGN17[GGGN17$FDR_GGvGN<0.05,]#394
green=GGGN17[GGGN17$FDR_GGvNN<0.05& GGGN17$FDR_GGvGN<0.05,]#90
orange=GGGN17[GGGN17$FDR_GGvNN<0.05 & GGGN17$FDR_GGvGN<0.05 &
                ((GGGN17$logFC_GGvNN < 0 & GGGN17$logFC_GGvGN > 0) |
                   (GGGN17$logFC_GGvNN > 0 & GGGN17$logFC_GGvGN < 0)),]#0

GGGN17_list=list(c1=green$genes,
                 c2=yellow$genes,
                 c3=blue$genes)

allResGGGN17 <- list()
for (i in 1:3) {
  candidate_list<- GGGN17_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResGGGN17[[i]] <- GenTable(myGOdata.bv,
                                parentchildFisher = resultParentchild,
                                orderBy = "parentchildFisher",
                                topNodes = sum(resultParentchild@score < 0.01)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResGGGN17) <- names(GGGN17_list)

write.xlsx(allResGGGN17$c1, file="GGGN_GO_B.xlsx", sheetName = "GGGN17_c1", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResGGGN17$c2, file="GGGN_GO_B.xlsx", sheetName = "GGGN17_c2", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResGGGN17$c3, file="GGGN_GO_B.xlsx", sheetName = "GGGN17_c3", col.names = TRUE, row.names = TRUE, append=TRUE)

allRes_plotsGGGN17 <-list()
for(i in 1:3) {
  allRes_plotsGGGN17[[i]] <- ggplot(head(allResGGGN17[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allRes)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_GGGN17=ggarrange(plotlist = allRes_plotsGGGN17, ncol = 3, nrow = 1,labels=c("missing", "unexpected", "expected"), vjust=1 )
annotate_figure(GO_GGGN17, top=text_grob("GGGN17"))

#===========================================================================================
####GGSS-GGGS####
#2hrs

GGGS2=full_join(DGE_2_2_gene$GGvSS$table %>% dplyr::select( genes, logFC_GGvSS=logFC, FDR_GGvSS=FDR),
                DGE_2_2_gene$GGvGS$table %>% dplyr::select( genes, logFC_GGvGS=logFC, FDR_GGvGS=FDR ))

plotGGGS2=ggplot()+
  geom_point(data=GGGS2, aes(x=logFC_GGvSS, y=logFC_GGvGS),color="#999999", alpha=0.30)+
  geom_point(data=GGGS2[GGGS2$FDR_GGvSS<0.05,], aes(x=logFC_GGvSS, y=logFC_GGvGS), color="#E6AB02", size=1)+ #=#F0E442 -> no diff although expected --> "lag of plasticity"
  geom_point(data=GGGS2[GGGS2$FDR_GGvGS<0.05,], aes(x=logFC_GGvSS, y=logFC_GGvGS), color="#56B4E9", size=1)+  # #3a538b -> not expected differences --> "stress"
  geom_point(data=GGGS2[GGGS2$FDR_GGvSS<0.05& GGGS2$FDR_GGvGS<0.05 &
                          ((GGGS2$logFC_GGvSS < 0 & GGGS2$logFC_GGvGS < 0) |
                             (GGGS2$logFC_GGvSS > 0 & GGGS2$logFC_GGvGS > 0)),],
             aes(x=logFC_GGvSS, y=logFC_GGvGS), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=GGGS2[GGGS2$FDR_GGvSS<0.05 & GGGS2$FDR_GGvGS<0.05 &
                           ((GGGS2$logFC_GGvSS < 0 & GGGS2$logFC_GGvGS > 0) |
                              (GGGS2$logFC_GGvSS > 0 & GGGS2$logFC_GGvGS < 0)),],
             aes(x=logFC_GGvSS, y=logFC_GGvGS ), color="#ef6c23", size=1)+
  labs(title="2hrs", x=(expression(paste("log FC R",bold("R"),"-S",bold("S")))), y=(expression(paste("log FC R",bold("R"),"-R",bold("S"))))) +
  theme_bw()+
  xlim(-12,12)+
  ylim(-12,12)+
  theme(aspect.ratio = 1)

#topGO
yellow=GGGS2[GGGS2$FDR_GGvSS<0.05,] #313
blue=GGGS2[GGGS2$FDR_GGvGS<0.05,] #0
green=GGGS2[GGGS2$FDR_GGvSS<0.05& GGGS2$FDR_GGvGS<0.05,] #0
orange=GGGS2[GGGS2$FDR_GGvSS<0.05 & GGGS2$FDR_GGvGS<0.05 &
               ((GGGS2$logFC_GGvSS < 0 & GGGS2$logFC_GGvGS > 0) |
                  (GGGS2$logFC_GGvSS > 0 & GGGS2$logFC_GGvGS < 0)),] #0

GGGS2_list=list(c2=yellow$genes)


allResGGGS2 <- list()
for (i in 1:1) {
  candidate_list<- GGGS2_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResGGGS2[[i]] <- GenTable(myGOdata.bv,
                               parentchildFisher = resultParentchild,
                               orderBy = "parentchildFisher",
                               topNodes = sum(resultParentchild@score < 0.01)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResGGGS2) <- names(GGGS2_list)

write.xlsx(allResGGGS2$c2, file="GGGS_GO_B.xlsx", sheetName = "GGGS2_c2", col.names = TRUE, row.names = TRUE, append=FALSE)


allRes_plotsGGGS2 <-list()
for(i in 1:1) {
  allRes_plotsGGGS2[[i]] <- ggplot(head(allResGGGS2[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResGGGS2)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_GGGS2=ggarrange(plotlist = allRes_plotsGGGS2, ncol = 1, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_GGGS2, top=text_grob("GGGS2"))

#17hrs
GGGS17=GGGS17=full_join(DGE_2_17_gene$GGvSS$table %>% dplyr::select( genes, logFC_GGvSS=logFC, FDR_GGvSS=FDR),
                        DGE_2_17_gene$GGvGS$table %>% dplyr::select( genes, logFC_GGvGS=logFC, FDR_GGvGS=FDR ))

plotGGGS17=ggplot()+
  geom_point(data=GGGS17, aes(x=logFC_GGvSS, y=logFC_GGvGS),color="#999999", alpha=0.30)+
  geom_point(data=GGGS17[GGGS17$FDR_GGvSS<0.05,], aes(x=logFC_GGvSS, y=logFC_GGvGS), color="#E6AB02", size=1)+
  geom_point(data=GGGS17[GGGS17$FDR_GGvGS<0.05,], aes(x=logFC_GGvSS, y=logFC_GGvGS), color="#414487FF", size=1)+
  geom_point(data=GGGS17[GGGS17$FDR_GGvSS<0.05& GGGS17$FDR_GGvGS<0.05 &
                      ((GGGS17$logFC_GGvSS < 0 & GGGS17$logFC_GGvGS < 0) |
                              (GGGS17$logFC_GGvSS > 0 & GGGS17$logFC_GGvGS > 0)),],
             aes(x=logFC_GGvSS, y=logFC_GGvGS), color="#21a585", size=1)+  # slope = 1 line
    geom_point(data=GGGS17[GGGS17$FDR_GGvSS<0.05 & GGGS17$FDR_GGvGS<0.05 &
                          ((GGGS17$logFC_GGvSS < 0 & GGGS17$logFC_GGvGS > 0) |
                             (GGGS17$logFC_GGvSS > 0 & GGGS17$logFC_GGvGS < 0)),],
             aes(x=logFC_GGvSS, y=logFC_GGvGS ), color="#ef6c23", size=1)+
  labs(title="17hrs", x=(expression(paste("log FC R",bold("R"),"-S",bold("S")))), y=(expression(paste("log FC R",bold("R"),"-R",bold("S"))))) +
  theme_bw()+
  xlim(-12,12)+
  ylim(-12,12)+
  theme(aspect.ratio = 1)

#topGO

yellow=GGGS17[GGGS17$FDR_GGvSS<0.05,] #215
blue=GGGS17[GGGS17$FDR_GGvGS<0.05,] #32
green=GGGS17[GGGS17$FDR_GGvSS<0.05& GGGS17$FDR_GGvGS<0.05,]#21
orange=GGGS17[GGGS17$FDR_GGvSS<0.05 & GGGS17$FDR_GGvGS<0.05 &
                ((GGGS17$logFC_GGvSS < 0 & GGGS17$logFC_GGvGS > 0) |
                   (GGGS17$logFC_GGvSS > 0 & GGGS17$logFC_GGvGS < 0)),]#0

GGGS17_list=list(c1=green$genes,
                 c2=yellow$genes,
                 c3=blue$genes)

allResGGGS17 <- list()
for (i in 1:3) {
  candidate_list<- GGGS17_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResGGGS17[[i]] <- GenTable(myGOdata.bv,
                                parentchildFisher = resultParentchild,
                                orderBy = "parentchildFisher",
                                topNodes = sum(resultParentchild@score < 0.01)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResGGGS17) <- names(GGGS17_list)

write.xlsx(allResGGGS17$c1, file="GGGS_GO_B.xlsx", sheetName = "GGGS17_c1", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResGGGS17$c2, file="GGGS_GO_B.xlsx", sheetName = "GGGS17_c2", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResGGGS17$c3, file="GGGS_GO_B.xlsx", sheetName = "GGGS17_c3", col.names = TRUE, row.names = TRUE, append=TRUE)

allRes_plotsGGGS17 <-list()
for(i in 1:3) {
  allRes_plotsGGGS17[[i]] <- ggplot(head(allResGGGS17[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allRes_plotsGGGS17)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_GGGS17=ggarrange(plotlist = allRes_plotsGGGS17, ncol = 3, nrow = 1,labels=c("missing", "unexpected", "expected"), vjust=1 )
annotate_figure(GO_GGGS17, top=text_grob("GGGS17"))

#===========================================================================================
####SSGG-SSSG####
#2hrs

SSSG2=full_join(DGE_2_2_gene$SSvGG$table %>% dplyr::select( genes, logFC_SSvGG=logFC, FDR_SSvGG=FDR),
                DGE_2_2_gene$SSvSG$table %>% dplyr::select( genes, logFC_SSvSG=logFC, FDR_SSvSG=FDR ))

plotSSSG2=ggplot()+
  geom_point(data=SSSG2, aes(x=logFC_SSvGG, y=logFC_SSvSG),color="#999999", alpha=0.30)+
  geom_point(data=SSSG2[SSSG2$FDR_SSvGG<0.05,], aes(x=logFC_SSvGG, y=logFC_SSvSG), color="#E6AB02", size=1)+ #=#F0E442 -> no diff although expected --> "lag of plasticity"
  geom_point(data=SSSG2[SSSG2$FDR_SSvSG<0.05,], aes(x=logFC_SSvGG, y=logFC_SSvSG), color="#414487FF", size=1)+  # #3a538b -> not expected differences --> "stress"
  geom_point(data=SSSG2[SSSG2$FDR_SSvGG<0.05& SSSG2$FDR_SSvSG<0.05 &
                          ((SSSG2$logFC_SSvGG < 0 & SSSG2$logFC_SSvSG < 0) |
                             (SSSG2$logFC_SSvGG > 0 & SSSG2$logFC_SSvSG > 0)),],
             aes(x=logFC_SSvGG, y=logFC_SSvSG), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=SSSG2[SSSG2$FDR_SSvGG<0.05 & SSSG2$FDR_SSvSG<0.05 &
                           ((SSSG2$logFC_SSvGG < 0 & SSSG2$logFC_SSvSG > 0) |
                              (SSSG2$logFC_SSvGG > 0 & SSSG2$logFC_SSvSG < 0)),],
             aes(x=logFC_SSvGG, y=logFC_SSvSG ), color="#ef6c23", size=1)+
  labs(title="2hrs", x=(expression(paste("log FC S",bold("S"),"-R",bold("R")))), y=(expression(paste("log FC S",bold("S"),"-S",bold("R"))))) +
  theme_bw()+
  xlim(-12,12)+
  ylim(-12,12)+
  theme(aspect.ratio = 1)

#topGO
yellow=SSSG2[SSSG2$FDR_SSvGG<0.05,] #313
blue=SSSG2[SSSG2$FDR_SSvSG<0.05,] #1
green=SSSG2[SSSG2$FDR_SSvGG<0.05& SSSG2$FDR_SSvSG<0.05,]#1 # same coordinate as green one --> makes sense
orange=SSSG2[SSSG2$FDR_SSvGG<0.05 & SSSG2$FDR_SSvSG<0.05 &
        ((SSSG2$logFC_SSvGG < 0 & SSSG2$logFC_SSvSG > 0) |
           (SSSG2$logFC_SSvGG > 0 & SSSG2$logFC_SSvSG < 0)),] #0

SSSG2_list=list(c1=green$genes,
                c2=yellow$genes,
                c3=blue$genes)


allResSSSG2 <- list()
for (i in 1:3) {
  candidate_list<- SSSG2_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResSSSG2[[i]] <- GenTable(myGOdata.bv,
                               parentchildFisher = resultParentchild,
                               orderBy = "parentchildFisher",
                               topNodes = sum(resultParentchild@score < 0.01)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResSSSG2) <- names(SSSG2_list)

write.xlsx(allResSSSG2$c1, file="SSSG_GO_B.xlsx", sheetName = "SSSG2_c1", col.names = TRUE, row.names = TRUE, append=FALSE)
write.xlsx(allResSSSG2$c2, file="SSSG_GO_B.xlsx", sheetName = "SSSG2_c2", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResSSSG2$c3, file="SSSG_GO_B.xlsx", sheetName = "SSSG2_c3", col.names = TRUE, row.names = TRUE, append=TRUE)

allRes_plotsSSSG2 <-list()
for(i in 1:3) {
  allRes_plotsSSSG2[[i]] <- ggplot(head(allResSSSG2[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResSSSG2)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_SSSG2=ggarrange(plotlist = allRes_plotsSSSG2, ncol = 2, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_SSSG2, top=text_grob("SSSG2"))

#17hrs
SSSG17=SSSG17=full_join(DGE_2_17_gene$SSvGG$table %>% dplyr::select( genes, logFC_SSvGG=logFC, FDR_SSvGG=FDR),
                        DGE_2_17_gene$SSvSG$table %>% dplyr::select( genes, logFC_SSvSG=logFC, FDR_SSvSG=FDR ))

plotSSSG17=ggplot()+
  geom_point(data=SSSG17, aes(x=logFC_SSvGG, y=logFC_SSvSG),color="#999999", alpha=0.30)+
  geom_point(data=SSSG17[SSSG17$FDR_SSvGG<0.05,], aes(x=logFC_SSvGG, y=logFC_SSvSG), color="#E6AB02", size=1)+
  geom_point(data=SSSG17[SSSG17$FDR_SSvSG<0.05,], aes(x=logFC_SSvGG, y=logFC_SSvSG), color="#414487FF", size=1)+
  geom_point(data=SSSG17[SSSG17$FDR_SSvGG<0.05& SSSG17$FDR_SSvSG<0.05&
                           ((SSSG17$logFC_SSvGG < 0 & SSSG17$logFC_SSvSG < 0) |
                              (SSSG17$logFC_SSvGG > 0 & SSSG17$logFC_SSvSG > 0)),],
             aes(x=logFC_SSvGG, y=logFC_SSvSG), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=SSSG17[SSSG17$FDR_SSvGG<0.05 & SSSG17$FDR_SSvSG<0.05 &
                          ((SSSG17$logFC_SSvGG < 0 & SSSG17$logFC_SSvSG > 0) |
                             (SSSG17$logFC_SSvGG > 0 & SSSG17$logFC_SSvSG < 0)),],
             aes(x=logFC_SSvGG, y=logFC_SSvSG ), color="#ef6c23", size=1)+
  labs(title="17hrs", x=(expression(paste("log FC S",bold("S"),"-R",bold("R")))), y=(expression(paste("log FC S",bold("S"),"-S",bold("R"))))) +
  theme_bw()+
  xlim(-12,12)+
  ylim(-12,12)+
  theme(aspect.ratio = 1)

#topGO

yellow=SSSG17[SSSG17$FDR_SSvGG<0.05,]#215
blue=SSSG17[SSSG17$FDR_SSvSG<0.05,] #698
#there was one red gene, so we need to differentiate green and red:
green=SSSG17[SSSG17$FDR_SSvGG<0.05& SSSG17$FDR_SSvSG<0.05 & ((SSSG17$logFC_SSvGG < 0 & SSSG17$logFC_SSvSG < 0) | (SSSG17$logFC_SSvGG > 0 & SSSG17$logFC_SSvSG > 0)),] #117
red=SSSG17[SSSG17$FDR_SSvGG<0.05 & SSSG17$FDR_SSvSG<0.05 & ((SSSG17$logFC_SSvGG < 0 & SSSG17$logFC_SSvSG > 0) | (SSSG17$logFC_SSvGG > 0 & SSSG17$logFC_SSvSG < 0)),] #1

#the "red" gene was however not annotated. Also the Heliconius ortholog wasn't annotated.
# genes logFC_SSvGG  FDR_SSvGG logFC_SSvSG  FDR_SSvSG
# 137 file_1_file_1_g8891   0.6844723 0.02574927  -0.6128706 0.02216444 --> HMEL014858-PA --> not annotated

SSSG17_list=list(c1=green$genes,
                 c2=yellow$genes,
                 c3=blue$genes)
                  #c4=red$genes) #not annotated -> remove from list otherwise error in topGo loop




allResSSSG17 <- list()
for (i in 1:3) {          # only 3, because "red" wasn't annotated
  candidate_list<- SSSG17_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResSSSG17[[i]] <- GenTable(myGOdata.bv,
                               parentchildFisher = resultParentchild,
                               orderBy = "parentchildFisher",
                               topNodes = sum(resultParentchild@score < 0.01)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}



names(allResSSSG17) <- names(SSSG17_list)

write.xlsx(allResSSSG17$c1, file="SSSG_GO_B.xlsx", sheetName = "SSSG17_c1", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResSSSG17$c2, file="SSSG_GO_B.xlsx", sheetName = "SSSG17_c2", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResSSSG17$c3, file="SSSG_GO_B.xlsx", sheetName = "SSSG17_c3", col.names = TRUE, row.names = TRUE, append=TRUE)


allRes_plotsSSSG17 <-list()
for(i in 1:3) {
  allRes_plotsSSSG17[[i]] <- ggplot(head(allResSSSG17[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allRes_plotsSSSG17)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_SSSG17=ggarrange(plotlist = allRes_plotsSSSG17, ncol = 3, nrow = 1,labels=c("missing", "unexpected", "expected"), vjust=1 )
annotate_figure(GO_SSSG17, top=text_grob("SSSG17"))


#===========================================================================================
####SSNN-SSSN -Switch 2!####
#2hrs

SSSN2=full_join(DGE_2_2_gene$SSvNN$table %>% dplyr::select( genes, logFC_SSvNN=logFC, FDR_SSvNN=FDR),
                DGE_2_2_gene$SSvSN$table %>% dplyr::select( genes, logFC_SSvSN=logFC, FDR_SSvSN=FDR ))

plotSSSN2=ggplot()+
  geom_point(data=SSSN2, aes(x=logFC_SSvNN, y=logFC_SSvSN),color="#999999", alpha=0.30)+
  geom_point(data=SSSN2[SSSN2$FDR_SSvNN<0.05,], aes(x=logFC_SSvNN, y=logFC_SSvSN), color="#E6AB02", size=1)+ #=#F0E442 -> no diff although expected --> "lag of plasticity"
  geom_point(data=SSSN2[SSSN2$FDR_SSvSN<0.05,], aes(x=logFC_SSvNN, y=logFC_SSvSN), color="#3a538b", size=1)+  # #3a538b -> not expected differences --> "stress"
  geom_point(data=SSSN2[SSSN2$FDR_SSvNN<0.05& SSSN2$FDR_SSvSN<0.05&
                          ((SSSN2$logFC_SSvNN < 0 & SSSN2$logFC_SSvSN < 0) |
                             (SSSN2$logFC_SSvNN > 0 & SSSN2$logFC_SSvSN > 0)),],
             aes(x=logFC_SSvNN, y=logFC_SSvSN), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=SSSN2[SSSN2$FDR_SSvNN<0.05 & SSSN2$FDR_SSvSN<0.05 &
                          ((SSSN2$logFC_SSvNN < 0 & SSSN2$logFC_SSvSN > 0) |
                             (SSSN2$logFC_SSvNN > 0 & SSSN2$logFC_SSvSN < 0)),],
             aes(x=logFC_SSvNN, y=logFC_SSvSN ), color="#ef6c23", size=1)+
  labs(title="2hrs", x=(expression(paste("log FC S",bold("S"),"-U",bold("U")))), y=(expression(paste("log FC S",bold("S"),"-S",bold("U"))))) +
  theme_bw()+
  xlim(-12,12)+
  ylim(-12,12)+
  theme(aspect.ratio = 1)

#topGO
yellow=SSSN2[SSSN2$FDR_SSvNN<0.05,]#639
blue=SSSN2[SSSN2$FDR_SSvSN<0.05,] #7
green=SSSN2[SSSN2$FDR_SSvNN<0.05& SSSN2$FDR_SSvSN<0.05,]#6
orange=SSSN2[SSSN2$FDR_SSvNN<0.05 & SSSN2$FDR_SSvSN<0.05 &
               ((SSSN2$logFC_SSvNN < 0 & SSSN2$logFC_SSvSN > 0) |
                  (SSSN2$logFC_SSvNN > 0 & SSSN2$logFC_SSvSN < 0)),]#0

SSSN2_list=list(c1=green$genes,
                c2=yellow$genes,
                c3=blue$genes)


allResSSSN2 <- list()
for (i in 1:3) {
  candidate_list<- SSSN2_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResSSSN2[[i]] <- GenTable(myGOdata.bv,
                               parentchildFisher = resultParentchild,
                               orderBy = "parentchildFisher",
                               topNodes = sum(resultParentchild@score < 0.01)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResSSSN2) <- names(SSSN2_list)

write.xlsx(allResSSSN2$c1, file="SSSN_GO_B.xlsx", sheetName = "SSSN2_c1", col.names = TRUE, row.names = TRUE, append=FALSE)
write.xlsx(allResSSSN2$c2, file="SSSN_GO_B.xlsx", sheetName = "SSSN2_c2", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResSSSN2$c3, file="SSSN_GO_B.xlsx", sheetName = "SSSN2_c3", col.names = TRUE, row.names = TRUE, append=TRUE)


allRes_plotsSSSN2 <-list()
for(i in 1:3) {
  allRes_plotsSSSN2[[i]] <- ggplot(head(allResSSSN2[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResSSSN2)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_SSSN2=ggarrange(plotlist = allRes_plotsSSSN2, ncol = 3, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_SSSN2, top=text_grob("SSSN2"))

#17hrs
SSSN17=SSSN17=full_join(DGE_2_17_gene$SSvNN$table %>% dplyr::select( genes, logFC_SSvNN=logFC, FDR_SSvNN=FDR),
                        DGE_2_17_gene$SSvSN$table %>% dplyr::select( genes, logFC_SSvSN=logFC, FDR_SSvSN=FDR ))

plotSSSN17=ggplot()+
  geom_point(data=SSSN17, aes(x=logFC_SSvNN, y=logFC_SSvSN),color="#999999", alpha=0.30)+
  geom_point(data=SSSN17[SSSN17$FDR_SSvNN<0.05,], aes(x=logFC_SSvNN, y=logFC_SSvSN), color="#E6AB02", size=1)+
  geom_point(data=SSSN17[SSSN17$FDR_SSvSN<0.05,], aes(x=logFC_SSvNN, y=logFC_SSvSN), color="#3a538b", size=1)+
  geom_point(data=SSSN17[SSSN17$FDR_SSvNN<0.05& SSSN17$FDR_SSvSN<0.05    &
                           ((SSSN17$logFC_SSvNN < 0 & SSSN17$logFC_SSvSN < 0) |
                              (SSSN17$logFC_SSvNN > 0 & SSSN17$logFC_SSvSN > 0)),],
             aes(x=logFC_SSvNN, y=logFC_SSvSN), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=SSSN17[SSSN17$FDR_SSvNN<0.05 & SSSN17$FDR_SSvSN<0.05 &
                          ((SSSN17$logFC_SSvNN < 0 & SSSN17$logFC_SSvSN > 0) |
                             (SSSN17$logFC_SSvNN > 0 & SSSN17$logFC_SSvSN < 0)),],
             aes(x=logFC_SSvNN, y=logFC_SSvSN ), color="#ef6c23", size=1)+
  labs(title="17hrs", x=(expression(paste("log FC S",bold("S"),"-U",bold("U")))), y=(expression(paste("log FC S",bold("S"),"-S",bold("U"))))) +
  theme_bw()+
  xlim(-12,12)+
  ylim(-12,12)+
  theme(aspect.ratio = 1)

#topGO
yellow=SSSN17[SSSN17$FDR_SSvNN<0.05,] #36
blue=SSSN17[SSSN17$FDR_SSvSN<0.05,] #20
green=SSSN17[SSSN17$FDR_SSvNN<0.05& SSSN17$FDR_SSvSN<0.05,] #6
orange=SSSN17[SSSN17$FDR_SSvNN<0.05 & SSSN17$FDR_SSvSN<0.05 &
                ((SSSN17$logFC_SSvNN < 0 & SSSN17$logFC_SSvSN > 0) |
                   (SSSN17$logFC_SSvNN > 0 & SSSN17$logFC_SSvSN < 0)),]#0

SSSN17_list=list(c1=green$genes,
                 c2=yellow$genes,
                 c3=blue$genes)

allResSSSN17 <- list()
for (i in 1:3) {
  candidate_list<- SSSN17_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological process terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResSSSN17[[i]] <- GenTable(myGOdata.bv,
                                parentchildFisher = resultParentchild,
                                orderBy = "parentchildFisher",
                                topNodes = sum(resultParentchild@score < 0.01)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResSSSN17) <- names(SSSN17_list)


write.xlsx(allResSSSN17$c1, file="SSSN_GO_B.xlsx", sheetName = "SSSN17_c1", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResSSSN17$c2, file="SSSN_GO_B.xlsx", sheetName = "SSSN17_c2", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResSSSN17$c3, file="SSSN_GO_B.xlsx", sheetName = "SSSN17_c3", col.names = TRUE, row.names = TRUE, append=TRUE)

allRes_plotsSSSN17 <-list()
for(i in 1:3) {
  allRes_plotsSSSN17[[i]] <- ggplot(head(allResSSSN17[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allRes_plotsSSSN17)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_SSSN17=ggarrange(plotlist = allRes_plotsSSSN17, ncol = 3, nrow = 1,labels=c("missing", "unexpected", "expected"), vjust=1 )
annotate_figure(GO_SSSN17, top=text_grob("SSSN17"))


#===========================================================================================
####NNSS-NNNS - Switch 2!####

NNNS2=full_join(DGE_2_2_gene$NNvSS$table %>% dplyr::select( genes, logFC_NNvSS=logFC, FDR_NNvSS=FDR),
                DGE_2_2_gene$NNvNS$table %>% dplyr::select( genes, logFC_NNvNS=logFC, FDR_NNvNS=FDR ))

plotNNNS2=ggplot()+
  geom_point(data=NNNS2, aes(x=logFC_NNvSS, y=logFC_NNvNS),color="#999999", alpha=0.30)+
  geom_point(data=NNNS2[NNNS2$FDR_NNvSS<0.05,], aes(x=logFC_NNvSS, y=logFC_NNvNS), color="#E6AB02", size=1)+ #=#F0E442 -> no diff although expected --> "lag of plasticity"
  geom_point(data=NNNS2[NNNS2$FDR_NNvNS<0.05,], aes(x=logFC_NNvSS, y=logFC_NNvNS), color="#3a538b", size=1)+  # #3a538b -> not expected differences --> "stress"
  geom_point(data=NNNS2[NNNS2$FDR_NNvSS<0.05& NNNS2$FDR_NNvNS<0.05  &
                          ((NNNS2$logFC_NNvSS < 0 & NNNS2$logFC_NNvNS < 0) |
                             (NNNS2$logFC_NNvSS > 0 & NNNS2$logFC_NNvNS > 0)),],
             aes(x=logFC_NNvSS, y=logFC_NNvNS), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=NNNS2[NNNS2$FDR_NNvSS<0.05 & NNNS2$FDR_NNvNS<0.05 &
                           ((NNNS2$logFC_NNvSS < 0 & NNNS2$logFC_NNvNS > 0) |
                              (NNNS2$logFC_NNvSS > 0 & NNNS2$logFC_NNvNS < 0)),],
             aes(x=logFC_NNvSS, y=logFC_NNvNS ), color="#ef6c23", size=1)+
  labs(title="2hrs", x=(expression(paste("log FC U",bold("U"),"-S",bold("S")))), y=(expression(paste("log FC U",bold("U"),"-U",bold("S"))))) +
  theme_bw()+
  xlim(-12,12)+
  ylim(-12,12)+
  theme(aspect.ratio = 1)

#topGO
yellow=NNNS2[NNNS2$FDR_NNvSS<0.05,] #639
blue=NNNS2[NNNS2$FDR_NNvNS<0.05,] #0
green=NNNS2[NNNS2$FDR_NNvSS<0.05& NNNS2$FDR_NNvNS<0.05,]#0
orange=NNNS2[NNNS2$FDR_NNvSS<0.05 & NNNS2$FDR_NNvNS<0.05 &
               ((NNNS2$logFC_NNvSS < 0 & NNNS2$logFC_NNvNS > 0) |
                  (NNNS2$logFC_NNvSS > 0 & NNNS2$logFC_NNvNS < 0)),]#0

NNNS2_list=list(c2=yellow$genes)


allResNNNS2 <- list()
for (i in 1:1) {
  candidate_list<- NNNS2_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResNNNS2[[i]] <- GenTable(myGOdata.bv,
                               parentchildFisher = resultParentchild,
                               orderBy = "parentchildFisher",
                               topNodes = sum(resultParentchild@score < 0.01)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResNNNS2) <- names(NNNS2_list)

write.xlsx(allResNNNS2$c2, file="NNNS_GO_B.xlsx", sheetName = "NNNS2_c2", col.names = TRUE, row.names = TRUE, append=FALSE)

allRes_plotsNNNS2 <-list()
for(i in 1:1) {
  allRes_plotsNNNS2[[i]] <- ggplot(head(allResNNNS2[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResNNNS2)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_NNNS2=ggarrange(plotlist = allRes_plotsNNNS2, ncol = 1, nrow = 1,labels=c("missing", "unexpected", "host specific"), vjust=1 )
annotate_figure(GO_NNNS2, top=text_grob("NNNS2"))

#17hrs

NNNS17=NNNS17=full_join(DGE_2_17_gene$NNvSS$table %>% dplyr::select( genes, logFC_NNvSS=logFC, FDR_NNvSS=FDR),
                        DGE_2_17_gene$NNvNS$table %>% dplyr::select( genes, logFC_NNvNS=logFC, FDR_NNvNS=FDR ))


plotNNNS17=ggplot()+
  geom_point(data=NNNS17, aes(x=logFC_NNvSS, y=logFC_NNvNS),color="#999999", alpha=0.30)+
  geom_point(data=NNNS17[NNNS17$FDR_NNvSS<0.05,], aes(x=logFC_NNvSS, y=logFC_NNvNS), color="#E6AB02", size=1)+
  geom_point(data=NNNS17[NNNS17$FDR_NNvNS<0.05,], aes(x=logFC_NNvSS, y=logFC_NNvNS), color="#3a538b", size=1)+
  geom_point(data=NNNS17[NNNS17$FDR_NNvSS<0.05& NNNS17$FDR_NNvNS<0.05 &
                           ((NNNS17$logFC_NNvSS < 0 & NNNS17$logFC_NNvNS < 0) |
                              (NNNS17$logFC_NNvSS > 0 & NNNS17$logFC_NNvNS > 0)),],
             aes(x=logFC_NNvSS, y=logFC_NNvNS), color="#21a585", size=1)+  # slope = 1 line
  geom_point(data=NNNS17[NNNS17$FDR_NNvSS<0.05 & NNNS17$FDR_NNvNS<0.05 &
                          ((NNNS17$logFC_NNvSS < 0 & NNNS17$logFC_NNvNS > 0) |
                             (NNNS17$logFC_NNvSS > 0 & NNNS17$logFC_NNvNS < 0)),],
             aes(x=logFC_NNvSS, y=logFC_NNvNS ), color="#ef6c23", size=1)+
    labs(title="17hrs", x=(expression(paste("log FC U",bold("U"),"-S",bold("S")))), y=(expression(paste("log FC U",bold("U"),"-U",bold("S"))))) +
  theme_bw()+
  xlim(-12,12)+
  ylim(-12,12)+
  theme(aspect.ratio = 1)

#topGO
yellow=NNNS17[NNNS17$FDR_NNvSS<0.05,] #36
blue=NNNS17[NNNS17$FDR_NNvNS<0.05,] #54
green=NNNS17[NNNS17$FDR_NNvSS<0.05& NNNS17$FDR_NNvNS<0.05,] #16
orange=NNNS17[NNNS17$FDR_NNvSS<0.05 & NNNS17$FDR_NNvNS<0.05 &
                ((NNNS17$logFC_NNvSS < 0 & NNNS17$logFC_NNvNS > 0) |
                   (NNNS17$logFC_NNvSS > 0 & NNNS17$logFC_NNvNS < 0)),]#0

NNNS17_list=list(c1=green$genes,
                 c2=yellow$genes,
                 c3=blue$genes)

allResNNNS17 <- list()
for (i in 1:3) {
  candidate_list<- NNNS17_list[[i]]
  annotation <-annotation<-"/Users/Kathi Schneider/Desktop/Host switch in Polygonia c-album/Manuscript/Submission/Revision/ReSubmission/SupplementaryMaterial/Annotation/braker_Pca_proteindatabase_agatLongestIsoform_goodTranscripts_AA.faa_2line_eggNOGdefaultparams.GO.eggNOG.tsv"    

  # here I will be only analyzing GO terms with at least 5 members,
  # as this yield more stable results.
  node_size=5

  # We will focus on Biological prmocess terms (BP)
  GO_category="BP"
  geneID2GO <- readMappings(annotation)
  names(geneID2GO) <- gsub("\\..*","",names(geneID2GO))
  geneUniverse <- names(geneID2GO)

  genesOfInterest.bv <- candidate_list

  genesOfInterest.bv <- as.character(genesOfInterest.bv)
  geneList.bv <- factor(as.integer(geneUniverse %in% genesOfInterest.bv))
  names(geneList.bv) <- geneUniverse

  myGOdata.bv <- new("topGOdata", description="Candidate genes", ontology=GO_category, allGenes=geneList.bv,  annot = annFUN.gene2GO, gene2GO = geneID2GO, nodeSize = node_size)

  # STATS#
  # when assessing a GO term, it takes into accoount the annotation of terms to the current term's parents,
  # and so reduces false positives due to the inheritance problem
  resultParentchild <- runTest(myGOdata.bv, algorithm="parentchild", statistic="fisher")

  allResNNNS17[[i]] <- GenTable(myGOdata.bv,
                                parentchildFisher = resultParentchild,
                                orderBy = "parentchildFisher",
                                topNodes = sum(resultParentchild@score < 0.01)) %>%
    mutate(parentchildFisher = as.numeric(parentchildFisher))
}

names(allResNNNS17) <- names(NNNS17_list)

write.xlsx(allResNNNS17$c1, file="NNNS_GO_B.xlsx", sheetName = "NNNS17_c1", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResNNNS17$c2, file="NNNS_GO_B.xlsx", sheetName = "NNNS17_c2", col.names = TRUE, row.names = TRUE, append=TRUE)
write.xlsx(allResNNNS17$c3, file="NNNS_GO_B.xlsx", sheetName = "NNNS17_c3", col.names = TRUE, row.names = TRUE, append=TRUE)


allRes_plotsNNNS17 <-list()
for(i in 1:3) {
  allRes_plotsNNNS17[[i]] <- ggplot(head(allResNNNS17[[i]],10), aes(x = -log10(as.numeric(parentchildFisher)), y = Term))+
    geom_point(aes(size = 100*Significant/Annotated))+
    labs(title = names(allResNNNS17)[i], x =expression('-log'[10]~'(p-value)'))+
    theme(legend.position = "none")
}

GO_NNNS17=ggarrange(plotlist = allRes_plotsNNNS17, ncol = 3, nrow = 1,labels=c("missing", "unexpected", "expected"), vjust=1 )
annotate_figure(GO_NNNS17, top=text_grob("NNNS17"))


#plot Switch 2

plot_S2_NG=ggarrange(plotNNNG2, plotNNNG17, plotGGGN2, plotGGGN17,  ncol = 2, nrow=2, labels=c("a", "b", "c", "d"))
plot_S2_SN=ggarrange(plotNNNS2, plotNNNS17, plotSSSN2, plotSSSN17,  ncol = 2, nrow=2,labels=c("a", "b", "c", "d" ))
plot_S2_SG=ggarrange(plotSSSG2, plotSSSG17, plotGGGS2, plotGGGS17,  ncol = 2, nrow=2, labels=c("e", "f", "g", "h" ))

ggarrange(plot_S2_SN, plot_S2_SG, ncol=1, nrow=2)
ggarrange(plot_S2_SN, plot_S2_SG, plot_S2_NG, ncol=3, nrow=1)

```


